\chapter{A Day of Writing}
 
\begin{quotation} 
  Exercise indeed we do, but that very forebackwardly,
  for where we should exercise to know, we exercise as having known.
\sourceatright{--- Sir PHILIP SIDNEY, Defence of Poesy}
\end{quotation}
\vspace{0.2cm}
 
 
  IT WAS TOO EARLY to rise. But Darconville, long before dawn, couldn't refrain
from literally jumping out of bed, the excitement in the sense of well-being he
felt serving as the best premonition 
\mynote{premonition}{(n.): The action of premonishing or forewarning; a previous notification or
  warning of subsequent events; a forewarning.}
possible for a full, uninterrupted day of writing, and when the sun bowled up 
over Quinsyburg---which Spellvexit, somersaulting, 
\mynote{somersault}{(v.): A leap or spring in which a person turns
heels over head in the air and alights on his feet; esp. such a feat as
performed by acrobats or tumblers; a pitchpoll. Hence, a turning over in this
fashion; a complete overturn, upset, etc. 
}
always greeted with a glossoepiglottic 
\mynote{glossoepiglottic}{(adj.): pertaining to the tongue and epiglottis.}
gurgle 
\mynote{gurgle}{(n.): A guttural sound such as is produced by irregular emission of air
  from the throat, gargling, or the like. }
of joy, something
like: "Gleep!"---he welcomed its appearance with three finished pages and an
emphatic resolve to write a few more.

  After his coffee, he smoked awhile, and then set to work again, his desk
cluttered with notebooks, pens, and piles of paper. The skull, norma frontalis,
\footnote{ \textit{norma frontalis}: Latin, the outline of the skull viewed from
  the front.}
was even smiling its approval this morning, for before long Darconville was
fully re-engaged in that silent and solemn duel in which the mind sits
concentrated in the most fearless of disciplines, the tidiness of which, he
felt, life could never hope to emulate and the wonderful and deep delight of
which nothing whatsoever else could hope to match. It was worth the loneliness.
It was worth the time. And if the blazing rockets of his imagination came
whistling down mere sticks, as occasionally they did, it was worth it still, for
truth indeed was fabulous and man, he'd always thought, best knew himself by
fable.

  Was truth, however, discovered or constructed? Darconville, actually, was
never really sure, and, of so-called "experience," well, when he thought of it
he tended to believe that it had to be avoided in order to write---a matter, in
fact, that couldn't have been given clearer focus than during these last few
weeks in Quinsyburg when up in those rooms, conjugating the games of speech and
light, writing pages racily colloquial, classically satirical, stinging and
tender in robust and ardent sequence, be had been interrupted with a frequency
that almost bordered on devotion: students-in-crisis; evangelists---two ruby-nosed
Baptanodons, 
\footnote{ a genus of large Jurassic enaliosaurs from the Baptanodon beds of the
Rocky Mountain region. These animals attained a length of 8 or 9 feet, and had
toothless jaws, orbits of great size, and a conic sclerotic ring of 8 plates.
The genus is closely allied to Ichthyosaurus.
}
especially, whose particular theory of disputation was that one
should aim, not to convince, but rather to silence one's opponent; and even
several locals who knocked to inquire, alas in vain, about the possibility of
buying his car, the black 1948 Mark VI Bentley in front of the house whose tires
they invariably kicked, crying, "What a mochine!" So he had his telephone pulled
out. He kept his shades pulled. He locked his door. And the Lords of Pleroma
\mynote{pleroma}{(n.): totality of the divine power}
stood by.

  When not teaching, Darconville worked very hard on his manuscript and, in
doing so, was entirely free from any feeling of having committed sacrilege
\mynote{scrilege}{(n.): The crime or sin of stealing or misappropriating what is consecrated
  to God's service.}
against the vow he had made with himself before coming there. His was a kind of
asceticism. 
\mynote{asceticism}{(n.):  The principles or practice of the Ascetics; rigorous self-discipline,
  severe abstinence, austerity. }
The writer was not a person, Darconville felt, rather an amanuensis
of verity, 
\mynote{verity}{(n.): Truth, either in general or with reference to a particular fact;
  conformity to fact or reality.}
who would only corrupt what he wrote to the extent, that he yielded
to passion or shirked 
\mynote{shirk}{(v.) OED 4.a. To evade (one's duty, work, obligations, etc.).}
the discipline of objectivity.

  The noon bell sounded from the library clock, work continued, and before too
long it was midafternoon, reminding him quite poignantly 
\mynote{poignantly}{(adv.): In a poignant manner; piercingly, pungently, acutely, keenly. }
of how slow the
imaginative struggle was---for were not artists those few of us flung 
\mynote{fling}{(v.): To move with haste or violence from or
  towards an object; to go or run violently or hastily; to dash, rush. }
down from the heavens into mortal garments?---and how difficult it was to 
return. To do, however, was not necessarily to make, nor to shape, to shape 
correctly. A maniacal 
\mynote{maniacal}{(adj.):  Affected with mania.}
stylist, Darconville worked to shape what he wrote---contour of form with
respect to beauty, coherence of matter with respect to blend---and to dig in
matter the furrows 
\mynote{furrow}{(n.): A narrow trench made in the earth with a plough, esp. for the
  reception of seed.}
of the mind, for in all creation matter sought form, form
matter, and that was as profound an exhortation 
\mynote{exhortation}{(n.): The action or process of exhorting, of earnestly admonishing or
  urging to what is deemed laudable conduct; an instance of this. }
to the artist as any: form
matter! The Greeks, he reminded himself, designated the world by a word that
means ornament, κόσμος, and the Romans gave it the name of mundus for its
finish, its grace, its elegance. And caelum? The word itself meant a tool to
engrave!

  The horizontal sun, shooting its rays through great dark banks of western
clouds, sent a last coppery glow under the shade, the fiery 
\mynote{fiery}{(adj.):  Consisting of or containing fire; flaming with fire.}
reflection of what was left of a good day. Dareonville closed his eyes, 
strained from concentration, and leaned back. With a furtive 
\mynote{furtive}{(adj.): Done by stealth or with the hope of escaping observation;
  clandestine, surreptitious, secret, unperceived. }
movement of his shoulders, he
turned, feeling suddenly a girl's phantom presence in the room. But he ignored
it and continued working on into the night, his face a shadow above the
gooseneck lamp---the cat snoring---rewriting the pages he'd spent the day on. 
It was abundance, to be alone, in the solitude of night, watching what you 
fashioned and fashioning by the miracle of art what was nothing less than 
giving birth by parthenogenesis. 
\mynote{parthenogenesis}{(n.): Process in which an unfertilized egg develops into a new individual;
  common among insects and some other arthropods; virgin birth.}
At last, in the middle of the close and quiet night, he saw he
was done. He looked up at the old watch hanging on the nail: late, late---the
tortoise of the hour hand, the hare of the minute hand epiphanizing 
\mynote{epiphanize}{(v.): to represent in a literary \textit{epiphany} 
  (A manifestation or appearance of some divine or superhuman being. ) }
the ambivalence of time that both weighed on him and bore him up. But then 
there on the desk, completed, lay the finished pages, washed with silver, 
wiped with gold. And the phantom?

  The light was out, and he was fast asleep, happier than anyone deserved to be,
and the only phantoms he could see were the benevolent ones he found in the
fleeting fancies of his dreams. And that was fine with him. Accident he would
leave to life which specialized in it.
