\chapter{ The Garden of Earthly Delight}

\begin{quotation}
\begin{verse}
  \hspace{-1.5em}There foamed rebellious Logic, gagged 
\mynote{gag}{(v.) OED 2.c. transf. and fig., esp. to deprive of power or freedom
    of speech; to stop the mouth of.}
and bound, \\
  There, stripped, fair Rhetoric languished
\mynote{languish}{(v.): OED 1. intr. Of living beings (also of plants or
    vegetation): To grow weak, faint, or feeble: to lose health, have one's vitality
    impaired; to continue in a state of feebleness and suffering. †In early use
    often: To be sick (const. of). }
on the ground.
\end{verse}
\sourceatright{--- ALEXANDER POPE, The Dunciad}
\end{quotation} 
\vspace{0.2cm}

  FACULTY MEETINGS are held whenever the need to show off is combined with the
imperative of accomplishing nothing. It was the case at Quinsy College. There
were five every year---a rump 
\mynote{rump}{(n.): OED 3.a. fig. A small, unimportant, or contemptible remnant
    or remainder of a body of persons (esp. of a Parliament).}
parliament of "educators," all bunged up with
complaints and full of prefabricated particulars, who sat around on their buns
for several hours trying to correct various problems, whether of school business
or personal bagwash, 
\mynote{bagwash}{(n.): The rough unfinished washing of clothes; a laundry that
   undertakes such rough washing; the ‘bag’ of clothes prepared for, or given, this
   type of washing. Also attrib.} 
and to confront problematical variables, every one of which
was always greeted with a salvo 
\mynote{salvo}{(n.): an outburst resembling the discharge of firearms or the
    release of bombs.}
of idle questions, nice distinctions, sophisms,
\mynote{sophism}{(n.): A specious but fallacious argument, either used
    deliberately in order to deceive or mislead, or employed as a means of
    displaying ingenuity in reasoning.} 
obs and sols, 
\footnote{obs and sols: objections and solutions; scholastic or subtle
    disputation.}
and last-ditch 
\footnote{last-ditch: last line of defense.}
stratagems. The presiding genius? Brizo, goddess of
sleep. The style? Extemporaneous. 
\mynote{extemporaneous}{(adj.): Not premeditated or studied, off-hand,
    extempore; esp. of discourse, prayer, etc. Rarely of a person: Speaking
    extempore; also, inclined to promptness of action. }
The participants? Teachers, who, like
whetstones, would make others cut that could not cut themselves.

  There were some preliminary matters, first. President Greatracks, sitting up
front like Buddha in a bad mood with his hands clasped in front of his belly as
indifferently as tweezers, 
\mynote{tweezers}{(n.pl.): OED 2. Small pincers or nippers (orig. as included in
  the contents of an etui) used for plucking out hairs from the face or for
  grasping minute objects.}
was being presented a gift (an unidentifiable rhomboid 
\mynote{rhomboid}{(n.): OED B.1. A quadrilateral figure having only its opposite
    sides and angles equal.} 
of pewter)
\mynote{pewter}{(n.): OED 1.a. A grey alloy of tin and lead, usually containing
    one fifth of its weight of lead, for which other metals are sometimes
    substituted, partly or entirely, in the composition of different varieties.}
---for his "unswerving 
\mynote{unswerving}{(adj.): OED, Not turning aside; steady, constant.}
leadership"---in the name of the senior
class whose metonym, 
\mynote{metonym}{(n.): a word that denotes one thing but refers to a related
   thing.}
Miss Xystine Chappelle, student body president and
sweetheart-of-the-school, after tweaking the tips of an organdy 
\mynote{organdy}{(n.) A very fine and translucent kind of muslin.}
the color of an orange fruit-jumble, 
\mynote{jumble}{(n.): a small thin usually ring-shaped sugared cookie or cake.}
curtseying, and delivering her message with cute little
eye-rolls, breathlessly concluded:
 
  ". . . as a class we really couldn't think of a more deserving leader. The
real meaning of college, hopefully, will remain with us even when we have to
face the reality of the world after graduation, the realization of which, I
realize, is not very realistic now. Anyway, I'm really proud to bestow this gift
in behalf of all of us who can never thank you enough for helping us reach our
realizable goals." [Applause]
 
  The agenda was full: it was the first such meeting of the year, and there was
a great deal to do. Swinging his nose, President Greatracks motioned to her feet
his secretary, a blinking anablept 
\footnote{\textdbend four-eyed anableps: some kind of fish?}
with mis-mated eyes and slingback 
\footnote{a shoe that has a strap that wraps around the heel.}
shoes who read the minutes from the previous spring and then handed out the updated
"Faculty Profile" booklets, after which Greatracks, his cheeks puffed out like a
Switzer's breeches, 
\footnote{puffed breeches wore by Swiss mercenaries}
stepped forward on his huge drawbridge 
\mynote{drawbridge}{(n.): A bridge hinged at one end and free at the other,
  which may be drawn up and let down so as to prevent or permit passage over it,
  or allow passage through the channel which it crosses.}
feet and said he wasn't going to be wordy because where was the damfool 
percentage in it? No, his only grouse, 
\mynote{grouse}{(n. slang.): OED, A grumble or complaint: a reason for
    grumbling. }
he said, was about the budget. Money, he pointed out, didn't grow
on trees, was the root of all evil, lost one friends, and talked!---too damned
much, he said. This year, he continued---it was prolegomenon 
\mynote{prolegomenon}{(n.): OED a. A preliminary discourse prefixed to a
  literary work; esp. a learned preface or preamble; chiefly in pl. introductory
  or preliminary observations on the subject of a book. From Gr. Gr. προλεγόµενον,
  neut. of pres. pple. pass. of προλέγειν to say beforehand, f. πρό, +
  λέγειν to say.}
to every faculty meeting---money was going to keep its little mouth shut, 
and they could all paste that in their hats, OK? Mr. Schrecklichkeit, an 
assistant biology professor with a white squill 
\mynote{squill}{(n.): OED 1.a. A bulb or root of the sea-onion or other related 
  plant. Chiefly in pl.}
of a nose, leaned over to Darconville and snapped, 
\mynote{snap}{(v.): OED 2.a. To utter sharp, tart, or cutting words or remarks;
   to speak or reply irritably or abruptly. Usu. with at.}
"There goes my course in Vegetable Staticks, that son of a bitch!"

  The meeting, then, was called to order---with several raps 
\mynote{rap}{(n.): A sharp and pretty loud knock, such as is produced by
  striking on a wooden surface with something hard; esp. a knock at a door, or (in
  recent use) one supposed to be made by a spirit. }
of the gavel 
\mynote{gavel}{(n.): OED 1.b. A president's mallet or hammer.} 
by President Greatracks, appearing rather like the Turk of legend who, ready to
drink a bottle of wine, first made loud noises and screwed out filthy faces to
warn his soul of the foul anti-Koranic act he was about to commit. First, Dean
Barathrum, a born remittance man 
\mynote{remittance man}{OED 2. an emigrant who is supported or assisted by
  remittances from home.}
and author of several out-of-date arithmetics,
introduced the newest members of the faculty. They were received, Darconville
and several others, with eye-watering yawns. 
\mynote{yawn}{(n.): OED 2. The or an act of yawning: a. Gaping or opening wide.
    b. Involuntary opening of the mouth, as from dwowsiness.}
One man was continuously tracing the sweephand 
\footnote{a hand marking seconds on a timepiece mounted concentrically with the
  other hands and read from the same dial as the minute hand.}
of his watch with his fingers. Another, staring out a window, was
actually sucking his thumb.

  Darconville's attention, however, had been drawn to the shiny-paged faculty
booklet, alphabetically listing everyone with photo and credentials. It looked
like a medieval bestiary: 
\mynote{bestiary}{(n.): OED 2. A treatise on beasts: applied to the moralizing
  treatises written during the Middle Ages.} 
skipjacks, 
\footnote{any of various fishes (as a ladyfish or bluefish) that jump above or
  are active at the surface of the water; especially :  skipjack tuna}
groutnolls, 
\footnote{=grout-head, A blockhead, thickhead, dunce. }
hysterical-looking circumferentors, 
\footnote{An instrument consisting of a flat brass bar with sights at
  the ends and a circular brass box in the middle, containing a magnetic needle,
  which plays over a graduated circle; the whole being supported on a staff or
  tripod.}
frumps 
\mynote{frump}{(n.): OED 5.a. A cross, old-fashioned, dowdily-dressed woman.
    Also rarely, said of a man. }
and filiopietistic 
\mynote{filiopietistic}{(adj.): of or relating to an often excessive veneration
  of ancestors or tradition.}
longheads, micelings, 
\footnote{\textdbend related to mice?}
whipsnades, 
\footnote{\textdbend maybe whipsnake?}
and many another whose eyes showed a very short limit of accommodation. Several had
actually taken no college degrees. Others were part-time evangelists, ex-army
colonels, and car salesmen. And the various titles of their scholarly
publications---books, articles, monographs, etc.---were scarcely believable:
"English Nose Literature"; Stephen Duck: More Rhyme Than Reason; "The American
Disgrace: Overabuse of the Verb `To Get'"; "Fundavit Stones in Crozet, Va.";
Much Ado About Mothing; 
\footnote{hunt for moths}
"The Psychopathological Connection Between Liquid Natural Gas and Agraphia"; 
\mynote{agraphia}{(n.): OED, Inability to write (as a manifestation of
  brain-disease). }
The Story of Windmill Technology; "The Significance
of Head Motions in Peking Ducks"; "Infusions as Drinks"; "Abraham Lincoln,
Quadroon?" 
\mynote{quadroon}{(n.): OED 1.a. One who is the offspring of a white person and
    a mulatto; one who has a quarter of Negro blood.}
and several other inventions, thought Darconville, of which necessity
was hardly the mother.

  Finally, the Great Consult began. They discussed tenure procedure. They
revised policy on sabbaticals. They rehearsed, to palsied 
\mynote{palsied}{(ppl.adj.): OED, Affected with palsy, paralysed; fig. deprived
  of muscular energy or power of action; rendered impotent; tottering, trembling.}
lengths, curriculum changes, cross-registration, crises in enrollment. It 
suddenly became a great din 
\mynote{din}{(n.): A loud noise; particularly a continued confused or resonant
  sound, which stuns or distresses the ear.} 
of objections, fierce denials, and loud peevishness 
\mynote{peevishness}{(n.): OED 3. Disposition to be vexed at trifles;
  moroseness, querulousness; fretfulness, petty or childish ill-temper.} 
all expressed in noises like the farting of laurel in flames 
\footnote{Don D. Nibbelink's Fearsome Folklore of Farting, p.33, attributed to 
  Swift, "My Lord, on Fire amidst the Dames, / Farts like a Laurel in the
  Flames."}
with everybody going 
at it head to head as if they were all trying right then and there to solve the 
problem of circular shot, perpetual motion, and abiogenesis!
\mynote{abiogenesis}{(n.): OED, The (supposed) origination or evolution of
living organisms from lifeless matter without the action of living parents;
‘spontaneous generation.’ (Introduced by Prof. Huxley in addressing Brit. Assoc.
at Liverpool, Sept. 1870.)} 

  Staring in disbelief, Darconville looked on in a kind of autoscopic
\mynote{autoscopic}{(adj.): from autoscopy—the experience in which the
  individual while believing himself to be awake sees his or her body position
  outside of his body.}
hallucination 
\mynote{hallucination}{(n.): OED 1. The mental condition of being deceived or
  mistaken, or of entertaining unfounded notions; with a and pl., an idea or
  belief to which nothing real corresponds; an illusion. }
as each of the faculty members rose in turn to make a point that
never seemed to have an acute end. It was all queer, makeshift, 
\mynote{makeshift}{(adj.): With which one makes shift; serving as a temporary
  and inferior substitute. }
and unpindownable, for all the cube-duplicators, angle-trisectors, and
circle-squarers seemed to keep busy avoiding any question that hadn't sufficient
strength to throw doubt on whatever answer couldn't have been offered anyway lest 
\mynote{lest}{(conj.): for fear that}
an inefficacious solution only prove to muddle a problem that couldn't be
raised in the first place. The discussion, rarely deviating into sense, grew
round with resolutions and amendments as they sacrificed the necessary to
acquire the superfluous 
\mynote{superfluous}{(adj.): That exceeds what is sufficient; of which there is
    more than enough; excessively abundant or numerous.} 
and did everything twice by halves, for, like Noah, they
had two of everything---two, it might be said, they didn't need so much as one of:
two policies, two excuses, two faces and, always, forty-eleven reasons to prop
up both.

  There was, for instance, Miss Shepe the witty, Miss Ghote the wise
---educatresses both, departments sociology and art education respectively---who
fell swiftly to reviewing the college motto: should it be "We Preach to Teach"
or "We Teach to Preach"?---a rabid 
\mynote{rabid}{(adj.): Furious, raging; madly violent in nature or behaviour.
  Also transf. of things or parts of the body. }
grace/ free-will discussion growing out of
their sudden but sustained failure to settle on the primacy of one over the
other. They squared off, adjusting their plackets and glaring into each other's
pinched and penny-saving faces. "I'm for less grapes and more fox," exclaimed
Miss Shepe, confusing everybody. Furious, Miss Ghote---brekekekekek!--- snapped 
\mynote{snap}{(v.): OED 16.a. To break (something) suddenly and cleanly; to
    break in two; to cause (a rope, etc.) to part or give way.}
her pencil in two.

  "So much for your deduction," said Miss Ghote.

  "I deduce nothing," sniffed Miss Shepe. "You've simply induced that yourself."

  "Induced, yes, what you'd implied."

  "You dare," snapped 
\mynote{snap}{(v.): OED 2.a. To utter sharp, tart, or cutting words or remarks;
   to speak or reply irritably or abruptly. Usu. with at.}
Miss Shepe, twisting her cramp-ring, 
\footnote{A ring held to be efficacious against cramp, falling sickness, and the
  like; esp. one of those which in pre-reformation times the kings and queens of
  England used to hallow on Good Friday for this purpose.}
"you dare infer I've implied what you yourself have induced, Miss Ghote?"

  "Put it this way," replied Miss Ghote with an icy-sweet smile, "you've only
surmised 
\mynote{surmise}{(v.): To form a notion that the thing in question may be so, on
  slight grounds or without proof; to infer conjecturally. Const. obj. cl. or
  simple obj.} 
I infer what you've implied I induced---and I do believe your bra strap's showing."

  Miss Shepe banged down her heel. "Then you conclude wrongly, Miss
Nothing-in-the-World-Could-Make-Me-Care-Less, that I surmised you infer what you
think I've implied you've induced!"

  "But you only assume that I conclude wrongly that you've surmised what---"

  "Elephant balls!" howled President Greatracks, the fat in his eyeballs
quivering. A group of old fishfags 
\footnote{A Woman who sells fish, more commonly called a fishwife}
from the home economics department, dosed 
\mynote{dose}{(v.): OED 2.b. transf. To add or apply a dose of something to}
to sleep by their own heavy perfume, immediately woke to clap their mouths in
horror.

  "I'm not certain I heard what he said," whispered Miss Swint to someone behind
her. It was a faint voice, some staring ghost suddenly exclaiming upon
Rhadamanth. The world to Miss Swint, piano teacher, her face two subtle shades
of oatmeal, backlit both by a monocle, consisted merely of music, her collection
of wheat-sheaf pennies, and the responsibility of playing the organ every Sunday
at the Presbyterian church, the very place in which, years ago, she'd long since
become convinced that maidenhead and godhead were indivisible.

  There were soon other matters on the docket: 
\mynote{docket}{(n.): OED 5. Law. A list of causes for trial, or of names of
  persons having causes pending. Hence phr. \textit{on the docket.} (U.S.)}
dining-hall duty, election to committees, chaperon 
\mynote{chaperon}{(n.): OED 3. fig. A person, esp. a married or elderly woman,
  who, for the sake of propriety, accompanies a young unmarried lady in public, as
  guide and protector.}
assignments. And some few raised questions about general
reform, and yet while only a mere fraction of the lot were actually concerned
with change---it was a subject met by children, with reform as the wicked
uncle---they all jumped up like minorités, 
\footnote{\textdbend a friar minor of Franciscan?}
jurisprudentes, and tub-thumping Sorbonnists 
\footnote{=Sorbonist, OED A doctor or student at the Sorbonne. Orig., a
theological college at Paris founded by Robert de Sorbon early in the 13th
century; the faculty of theology in the old University of Paris, of great
importance down to the 17th century; now, a constituent part of the University
of Paris (Paris IV), also housing departments from other sections of the
University.}
to debate it, all reinforcing the "yo-he-ho" protoglottological
\footnote{\textdbend glottology is the science of tongues or languages, 
  \textit{proto} indicates the first, earliest or original.}
theory that words initially began as shouts. No aspect was overlooked, no fine
point ignored, no issue diminished. It was complete havoc once again as they
stood in coalition or squatted 
\mynote{squat}{(v.): OED 6, Also jocularly, to sit (down).}
in caucus, breaking down every proposition like reformational hairsplitters 
\mynote{hairsplitter}{(n.): OED, One who 'splits hairs'; one who makes minute or
  over-refined distinctions.}
into partitions, sections, members, subsections,
submembral sections, submembral subsections and denouncing each other with
mouthfuls of rhetoric warped by quiddling, 
\mynote{quiddling}{(vbl.) a trifling}
diddling, 
\mynote{diddle}{(v.) OED, to waste time in the merest trifling}
and undistributed middling.
One third believed what another third invented what the other third laughed at.
Quid the Cynic argued with Suction the Epicurean, Suction the Epicurean argued
with Sipsop the Pythagorean, Sipsop the Pythagorean argued with Quid the Cynic,
and the whole afternoon dwindled away with one saving at the spigot 
\mynote{spigot}{(n.) OED 1. A small wooden peg or pin used to stop the vent-hole of a barrel or cask; a vent-peg; a similar peg inserted into and controlling the opening or tube of a faucet and used to regulate the flow of liquor. }
and another letting out at the bunghole.
\footnote{A bunghole or bungehole is a hole bored in a liquid-tight barrel to remove contents. The hole is capped with a large cork-like object called a bung. }

  It occurred to Darconville that the expression, "ignorance is bliss," was a
perception curiously unavailable to the ignorant as he considered, with great
misgivings, the sad and inarticulate desperviews there whose identities had gone
soft on them and whose grasp of reality was so slight and so arbitrary and so
grotesque that each could have easily stepped from there into the hot
pornofornocacophagomaniacal 
\footnote{\textdbend A.T's neologism.}
set of Bosch's unmusical hell and fit, snugly.

  The meeting, finally, adjourned. 
\mynote{adjourn}{ }
A partition was rolled back, revealing at
another section of the long room several tables with bottles of soft drinks (no
liquor) and plates of cookies, candy, and cake. Papers rattled, 
\mynote{rattle}{ }
people coughed, and chairs shifted as everyone withdrew.

  "Y'all ever see so dang much shuck for so little nubbin?" complained President
Greatracks who---a fake giant among real pygmies ---stomped toward the food in the
company of several pipe-smoking lackeys, all dodging about in his circumference.
No, no, they agreed, no, they hadn't. His arms upon the boobies' shoulders, you
quickly saw the gudgeon bite. No, no, they repeated, not at all. "Sombitches,"
he said, picking up five thumbprinted marshmallows and stuffing them into his
mouth. "I can hear their brains rolling about like B-B's in a boxcar."

  The Quinsy faculty, during the refreshments, took the occasion to gossip, but
faction didn't really constitute disunity. It was as if, somehow, they had all
been destined by some temporal and spatial anti-miracle of history to come
together at the same place and time and to know each other on the instant by
some mystic subtlety for mates, perhaps in the paradox of peculiar faces which,
while each was different, all looked as though God had smeared them when still
wet. They didn't have to talk. They weren't speakers, really. But they were
experts at malversation.

  "I know about budgets," said Miss Throwswitch, the drama teacher, with a
scrawl of mockery under the rims of her eyes as she poured a soda. "This year,
you watch, we'll be doing Chekhov in clothesbags and clunkies."

  "They are saying"---Prof. Fewstone's standard opening line, never without its
veiled threat---"they are saying that over in Richmond the legislature is cutting
us off without a dime unless we put some Mau-Maus on the faculty." A history
professor, Fewstone always kept to that one note, like a finch; his sour
doctrines, masquerading as brands of economics and politics, fit his reputation
as a miser far better than the hackberry-colored jacket he never took off,
perhaps for that elegant pin (compass, square) awarded him by some Brotherhood
of Skinks or other who worshipped trowels and pyramids. "Now in my book, The
Rehoboths: Reform or Reglementation? I put forth the theory---"

  "Attitudinizing," murmured Miss Gibletts, the Latin teacher whose thin, dry,
hectic, unperspirable habit of body made her somewhat saturnine. She looked like
St. Colitis of the Sprung Chair. Turning away to find another conversation, she
carefully managed to avoid the dark stranger with the French name who'd just
been introduced and who was standing all alone.
  Miss Dessicquint, the assistant dean, looked like Nosferatu---a huge mustachioed
godforgone, inflexible as a Dutch shoe, who was given to lying about her age
before anybody asked her. She closed her eyes and begun to hum. "Look," she
sneered, nudging her secretary and sometime bowling partner, Miss Gupse, an
unhealthy-looking little poltfoot with one of the longest noses on earth.
"There, by the table, in harlequin shorts."
  "Oh no. Floyce."
  "The flower of fairybelle land."
  "He irks me, that one."
  "He?"
  Miss Gupse, smiling, caught the irony and licked her nose. "She."
  "It!" said Miss Dessicquint, wagging a tongue long as a biscuit-seller's
shovel. Her secretary had to turn around to stop the smile glimmering down the
flanks of her steep nose. "He looks like a Mexican banana-split."
  "I wonder," swallowed little Miss Gupse, "how those people"---she looked up over
a cookie---"well, do it."
  "Have you ever had the occasion," asked Miss Dessicquint, pausing for dramatic
effect, "to look at the east end of a westbound cow?"
  Floyce R. Fulwider, reputedly a ferocious alcibiadean, was a balding,
fastidious art teacher---he pronounced Titian, his favorite painter, for instance,
to rhyme with Keatsian---who could often be seen skipping across the Quinsy campus
holding by his fingertips the newly wet gouaches and undernourished
para-menstrual creations of his students or, as he called them, his "popsies."
The college teemed with stories about the felonious gender-switching parties he
threw, when he'd paint his windows black, put on his Mabel Mercer records, and
encourage everyone to don feather boas and run around naked. He couldn't have
liked it more. But Miss Dessicquint didn't like it and told Miss Gupse,
confidentially, that he was getting a terminal contract this year. "Rome," she
whispered into her companion's ear, "wasn't burnt in a day."
  Then, turning, they saw Darconville.
  "I hope he isn't a bumbie."
  "He looks normal to me," said Miss Gupse.
  "Well, you know what they say. And it's true," said Miss Dessicquint, with
eyes like frozen frass, "there's a thin line between madness and insanity."
  Miss Gupse, never strong, secretly thought herself the object of this cut and
that very evening would proceed to call her sister in Nashville to inquire
tearfully about a job there. True, that particular afternoon she had been in the
midst of her mois. But the memory of having a big nose from childhood is perhaps
at the bottom of more havoc than one can ever know.
  In one corner of the room, occupying much of it, stood Dr. Glibbery, a
walrus-arsed microbiology professor with a face, spanned by a dirty ramiform
mustache, that had the nasty whiteness of boiled veal. His body resembled the
shape in microcosm and odor in fact of one of those nineteenth-century lamps,
the kind with a reservoir that gurgled periodically, emitted a stench of oil,
and often exploded. A fanatical right-wing at-once-ist, he bullied his
students---calling them all "Dufus"---with incredible lies about the difficulty of
taking a doctorate and how, in pursuit of his own, he'd lived on ketchup
sandwiches, got hookworm from going barefoot, and had to study upsidedown to
keep awake. In point of fact, he had taken his degree from one of those
just-about-accredited polytechnical millhouses in Virginia, identified usually
by its initials, where the only requirements for graduation were to belong to
the DeMolay and have two thumbs and a reasonably erect posture.
  "None of these here dufuses work anymore. Well," pontificated Dr. Glibbery,
"I'm going to knock their tits right into their watch-pockets for them, OK? You
won't spy no grade of A in my roll-book," he said, "for love nor money."
  He mumped a fistful of horehound drops and winked.
  "Or money," corrected Prof. Wratschewe, doyen of the English department and
proud author of several monographs, still, alas, awaiting the recognition of an
indifferent world: "Bed-Wetting Imagery in Chatterton"; "Packed Earth as
Anti-Resurrection Symbol in Wordsworth"; "The Caesura: Rest Home of Rhetoric";
and one slim book, Menus from Homer (o.p.).
  But Dr. Glibbery only farted and grinned.
  Miss Pouce, standing there, almost died of embarrassment---and, turning quickly,
banged into a wall. Great snorts of laughter accompanied this, then grew---the
lackeys, going beetred, doubled over with hilarity---as Greatracks, mimicking her,
waddled yoketoed into the same wall, went crosseyed, and grabbed his nose. Qwert
Yui Op, a midget Tibetan-Chinese, and Miss Malducoit, of the dirigible hips,
both in math, applauded---and she spilled her drink. Darconville went to pick up
her cup. "Thank you very much," said Miss Malducoit coldly, stepping in front of
him, "I can do it myself."
  Darconville, unwittingly, backed into a group of old ladies. It was the
delegation, each old enough to carbon-date, from home economics. In the midst of
them, studiously arranged that way, loomed a huge unproliferative boffa with
tied oviducts and blimp-like breasts which sagged according to the law of S=½vt2
named Mrs. DeCrow, associate professor in American history and local U.D.C.
chairwoman, who was blowing a loud ranz des vaches across the room, subject: her
summer in Biloxi. She had a voice like a faulty drain. There was a cruel pin in
her hat, and she wore an eighteenth-century brooch with a painting of a litter
of pigs on it. Famously disagreeable, at least so Dr. Dodypol, a colleague, told
Darconville, she harbored a monomaniacal loathing---something not a few
Southerners shared---for Abraham Lincoln.
  "I know you but you don't know me," Mrs. DeCrow announced to Darconville,
hoping she was wrong.
  An arrangement, thought Darconville, that suits me fine.
  "Wait," she suddenly exclaimed, holding a finger to her nose and turning to
the other ladies, "but now who does this boy remind me of?"
  Everyone dutifully waited.
  "Not that awful Edwin Albert Poe who wrote horror poetry?" asked a sagcheeked
lady cretinously mis-shoving a wedge of cake into her face.
  "No," said bossy Mrs. DeCrow, cocking a quick snook at the lady. "I know," she
said, interrupting herself, "Sir Thomas More."
  "Saint Thomas More," corrected Darconville, for whom such distinctions meant
something. But he was smiling.
  Mrs. DeCrow, arching an eyebrow, didn't care. That remark, though she smiled a
hard smile, immediately cost him---and irrevocably---any invitation whatsoever to
one of her Fridays, even if he lived to be a hundred. She knew where to stick
the knife in.
  Darconville submitted to his uneasiness and in the contending noise, the
rambling incoherent flow of talk, found that the most delectable social delight
there seemed to be the quarrel that stopped just short of violence, discussions
ending in repeated barks and loud flat interrobangs. It was a tragedy of
language. There was no dialogue, only monologue, wherein each of the
thimbleriggers and pathologically self-defensive opsimaths gathered there, most
of them educated on the fly at one of those narrow and sectarian
chop-and-chance-it swot factories in that region which was named after some
square-capped fifteenth-century Protestant joykiller, only kept to a mad and
self-referent fixity of subject and blathered on to no consequence.
  Suddenly, there was a buzz of intrigue by the doorway: someone reporting a car
accident. Xystine Chappelle, calling from the Quinsyburg Hospital, had reported
that, while driving them home, Miss Shepe and Miss Ghote, sitting in the back
seat, had willfully resumed their "I-think-that-you-think-that-I-think"
argument, whereupon Miss Ghote leaped in frustration from the moving car and
broke her femur.
  "Them ol' beangooses," muttered Dr. Glibbery, fingering through the candies,
"I do believe they gone get married one day."
  "Are you?" Darconville, surprised, looked behind him. "Are you married?" It
was the German instructor, Miss Tavistock (a.k.a. "The Clawhammer"), a
flat-chested girl with a cataleptic smile who had the curious habit of
suggesting to perfect strangers that, frankly, she thought they were in love
with her.
  "No, I'm not," answered Darconville who, suddenly noticing Mrs. DeCrow
empty-handed, offered her some cookies, but she would have none of it and walked
past him in a snit---the snit that would last a lifetime.
  "The writer," sighed Miss Sally Bull Sweetshrub, the tournure of her phrases
as precise as cut glass, "has no time whatsoever for such things as marriage.
No, I'm afraid she has not. The devotion which asks her to feel the deliberation
of art asks also that she choose the single life." A senior member of the
creative writing department, Miss Sweetshrub carefully lifted off her
half-glasses and, patronizingly turning her head from person to person, lectured
the few people nearby on the difficulty in question. She was---how to put it?---an
"unclaimed blessing." "I am a novelist, you see," she said, "which is spelled
d-e-v-o-t-i-o-n."
  "I see," said Mrs. McAwaddle, the registrar, who passing by with a tidy of
licorice-whips on a napkin and a paper cup of purple soda unintentionally got
hooped into the conversation.
  "But do you? Do you really?" asked Miss Sweetshrub.
  Mrs. McAwaddle, to be kind, gave the mandatory sign of ignorance by pretending
to reconsider and said, "Mmmmmmm." Long having been a fretful auditor to Miss
Sweetshrub's explicit polysemantic chats, she knew the hopelessness of trying to
extricate herself. "Well," she reflected, charitably, "I have no doubt but that
I might not."
  Prof. Wratschewe, ducking his head in, interjected. "Excuse me, Mrs.
McAwaddle? The word `but,' I believe, is used unnecessarily after the verbs `to
doubt' and `to help.' Strunk \& White, Book Four, I think you'll find."
  "I wouldn't have made that mistake," riposted Miss Sally Bull Sweetshrub,
bowing her head and fluttering her eyelids with the advantage Prof. Wratschewe's
old, but as yet inconsequential, crush on her allowed.
  Miss Sally Bull Sweetshrub, aet. 50 or so, was classically affected and wore
out-of-date dresses that looked like Morris wallpaper. She had a perched
Jacobean nose, like a dogvane, a waltzing oversized bottom, and wore her
pissburnt-colored hair in an outlandish bun at the back anchored by several
severe combs that matched in horn her conversational brooches. A bottomless fund
of clarifications and cruelly sharpened pencils, she was known for giving
strawberry teas and answering all her mail on Featherweight Antique Wove, Double
Royal 90 when she couldn't attend a function, which, of course, she always
could---and, when so doing, always did in the company of her pet beagle, "Howlet,"
and a monstrous alligator friend who some years ago left its native land in the
form of a handbag. She was the pillar of Quinsy College, a woman both
impregnable and unbearable (what she might have borne were she pregnable would,
indeed, be unimaginable!)---one of those starboard-leaning High Anglicans, famous
and innumerable in the Old Dominion, whose factitious sense of vanity in a
private conversation makes it frankly public: the raised and punishingly
assertive voice of the insane queen which includes, willingly or not, everybody
within earshot.
  "You perhaps know my work?"
  She was looking at Darconville with her eyes shut, her lavender gloves
buttoned shut and crossed immaculately in her lap.
  "Possibly. Could you name---?"
  "My novels? Let me see. My first was Answer Came There None, then,
respectively, I, a Stranger; Also but Not Yet the Wombat Cries; The Big Regret;
The Interrupted Woman; The Same, Only Different; and The Black Duchess." It was
the genre of course of Hoodoo, Hackwork, and Hyperesthesia, the popular
dustjacket for which always showed a crumbling old mansion-by-moonlight and a
frightened beauty in gossamer standing before it, tresses down, never knowing
which way to turn.
  Darconville, listening in spite of himself, was trying to repress an image of
the prattling angel, Glossopetra, who reputedly had fallen from heaven to spend
his eternal punishment wagging like a human tongue.
  "There was a volume of my early verse, Naps Upon Parnassus. I followed that,"
continued Miss Sweetshrub, "with a critical work on Robert Browning called The
Snail on the Thorn." Miss Throwswitch, standing behind her, periscoped two
little fingers over her head and wiggled them, while Mr. Schrecklichkeit and
Qwert Yui Op smiled into their sleeves.
  "Mmmmm," said Mrs. McAwaddle, breathing her hum back in, a sound more of
fatigue than recognition.
  "Thankfully nothing on Abraham Lincoln!" piped in Mrs. DeCrow, making a loud
transition to the powder room.
  Miss Sweetshrub chastely crossed her legs. "My most widely reviewed book?
Difficult to say. I guess I'd have to say it was my Tlot! Tlot! The Biography of
Alfred Noyes."
  "My," said Mrs. McAwaddle, hooing low but wisely, like Minerva's owl, "you've
sure written a lot."
  "Permit me? A lot is a parcel of land. I have written"---she drew her little
finger fastidiously across her eyebrow---"much." Then she smiled around sweetly at
everyone, especially Prof. Wratschewe whose approval was never in doubt if the
handful of tiny white valentine candies (inscribed "Skidoo," "Be Mine," "Kiss
Me," etc.) he solicitously extended to her meant anything. For eleven years
there had run the possibly not unfounded rumor they'd be married any weekend
now---who knew, perhaps the five-hundred-and-seventy-third would be it?
  "Good grief," cried Prof. Wratschewe over his pocket-watch. "It's late---and
I've a lecture to prepare this weekend."
  "Weekends few," sighed Miss Sally Bull Sweetshrub, recrossing her hands and
anticipating, perhaps, only the title of another novel she'd one day write, one
she saw somewhat wistfully that there'd be plenty of time for as she sat there
silently, as if straining to hear the celesta of wedding bells she felt could
now peal only across the landscape of her fiction.
  And so it went.
  The last rays of the afternoon sun brought a feeble light through the
meeting-room, and Darconville, heading toward the doorway, found a last group of
teachers to meet. He was introduced to Miss Ballhatchet, a muscular valkyrie
from the physical ed department; Dr. Excipuliform, a history professor with
afflicted eyes and a Transylvanian hop in his speech who despised the left-wing
American press ("Intellectschultz! Velfare! Gummonism! Rewolt! Neekroes!");
cheery Miss Skait, from physics, who, immune to paradox, bouncingly assured him
that Quinsy College was actually one big family; the Weerds---Aldo and Dodo---a
young Pekinese-faced couple in the English department who spoke only to each
other, wore rucksacks, and had occasional bouts of purpura haemorrhagica which
they insisted on calling poetry; a six-foot maliarda named Miss Porchmouth,
chemistry, who had a handshake that felt like pebbledash-siding; a certain Mr.
Bischthumb, head of anthropology, who, tucking into an enormous piece of cake,
bit off half the doily which got stuck in his weasand---and, wheezing tearfully,
had to be raised by four men and lifted out of the room kicking skyward
cataleptically in one of the most indecent postures Darconville had ever seen.

  There were others. Darconville met Mr. Thimm, government department, who told
him, several times, about his wife's egg-in-the-armpit deformity; Drs.
Knipperdoling and Pindle, two sententious ballachers from economic geography,
who in a little vaudeville skit of theirs constantly pitched back and forth
various adages and truisms about life, invariably beginning, "Well, I've learned
one thing from the rough-and-tumble . . ." or "Now, I've been shinnying up this
old stick for forty some odd years, and . . ."; and, finally, Dr. Speetles from
the general education department, an anti-intellectual gepid who claimed the
study of Skakespeare to be completely worthless, preferring instead the applied
sciences and confessing wittily, "I teach all my classes, see, using only a toy
fire-engine. It's a multi-concept factor, you know? With this as a point of
departure," he said, sucking a candy-spangle from his thumb, "I can teach
sociology, government, civics, visual and environmental studies---hell, you name
it." He tapped the side of his nose and nudged Darconville, who at that point
was seriously on the brink of calling upon the angel, Hodniel, who supposedly
had the power of curing stupidity in men. Instead, he put on his coat.

  Walking out, one of the unaligneds of the faculty who may have stood, perhaps,
for the several Darconville hadn't yet managed to meet, confided to him, "You
know, when ol' Greatracks bangs down the gavel in these meetings and says,
`Begin!' well, I always think, shoot, that's about the longest word in the damn
books---and ain't that the truth."

  The thought hadn't been uttered a second when Prof. Wratschewe, grammarian,
suddenly jumping-jacked out of nowhere and said, "Actually, I think you'll find
that the longest word found in literature, Aristophanic in origin, is:
%\textit{Lopadotemachoselacogaleokranioleipsanodrimhypotrimmatosilphioparaomelitokatakechymenokichlepikossyphophattoperisteralektryonoptekephalliokigklopeleiolagoisiraiobaphetragonopterygn}."
\textit{
Lopadotemachoselacogaleokranioleipsanodrimhypotrimmatosilphioparaomeli-\\
tokatakechymenokichlepikossyphophattoperisteralektryonoptekephalliokigklo-\\
peleiolagoisiraiobaphetragonopterygn}."

  Somehow, it summed up the day.
 
