\chapter{ Quinsy College}

\begin{quotation}
  A hen is only an egg's way of making another egg.
\sourceatright{---SAMUEL BUTLER}
\end{quotation}
 
  QUINSY COLLEGE, est. 1839, was a quaint 
\mynote{quaint}{(adj.):  unusual or uncommon in character or appearance, but at the same time
having some attractive or agreeable feature, esp., having an old-fashioned
prettiness or daintiness.
}
old respectable school for girls. It
stood in the seminary tradition of the female academy: a chaste 
\mynote{chaste}{(adj.): chastened, modest, restrained from all excess. }
academic retreat, moral as peppermint, 
\mynote{peppermint}{(n.): A species or subspecies of mint (Mentha piperita), 
  cultivated for its essential oil. }
built in semi-colonial red brick and set back in a deep green delling 
\footnote{ \textdbend Probably related to \textit{dell}, A deep natural hollow
  or vale of no great extent, the sides usually clothed with trees or foliage.}
where, alone---at least so felt the Board of Visitors (ten FFVs
with swimming eyes, three names, and hands with liverspots
\mynote{liverspot}{(n.): age spot}
)---one's daughter could
be lessoned in character and virtue without the indecent distractions that
elsewhere, everywhere else, wherever led to vicious intemperance, 
\mynote{intemperance}{(n.):  Lack of moderation or restraint; excess in any kind of action;
  immoderation; spec. excessive indulgence of any passion or appetite. }
Bolshevism, and free thought. There were other girls' schools in the area 
---Falcon Hall, Longwood College, St. Bunn's---but none was quite so singular 
as Quinsy.

  It had been strictly private years ago, one of those dame 
\mynote{dame}{(n.):  informal terms for a (young) woman}
schools in the
South, usually called something like Montfaucon or Thirlwood or Miss Tidy's
Establishment for Young Ladies and run by a woman with a name like Miss
Monflathers, a bun-haired duchess of malfeasance 
\mynote{malfeasance}{(n.):  wrong-doing}
from the English-Speaking Union who was given to wearing sensflectum 
\footnote{ \textdbend Theroux's coinage.}
crinolines 
\mynote{crinoline}{(n.): a stiffened or structured petticoat
designed to hold out a woman's skirt, popular at various times since the
mid-19th century. Originally, crinoline described a stiff fabric made of
horsehair ("crin") and cotton or linen which was used to make underskirts and as
a dress lining.
}
and horsehair jupon 
\mynote{jupon}{(n.): a short kirtle worn by women. }
and whipping her girls at night. At the turn of the century, however, Quinsy 
came under state receivership and, although suffering the shocks of democracy, 
remained yet blind
to change. It was an institution, still, whose expressed intention was to
diminish in distance and time the dangers of creeping modernity and with prudes
\mynote{prude}{(n.):  A woman who maintains or affects excessive modesty or propriety in
conduct or speech; one who is of extreme propriety: usually applied adversely
with implication of affectation. 
}
for proctors 
\mynote{proctor}{(n.): In modern use, as at Oxford and Cambridge, each of two officers
appointed annually to discharge various functions in connexion with the meetings
of the University and its various Boards, the examinations and conferment of
degrees, and the like; they are also charged with the discipline of all persons
in statu pupillari, and the summary punishment of minor offences.
}
and dowagers for deans to produce girls tutored in matters not only
academic but on subjects touching on the skillet, 
\mynote{skillet}{(n.): a frying pan, a (heavy) cooking pan.}
the needle, and, though
strangers yet to pain, even the nursery, a matter, it was confidentially given
out, not unrelated to that regrettable but thankfully fleeting moment during
which they would simply have to bite on a bullet and endure.

  A legendary respect for the Southern lady---doubt it who dares!--- was all 
through the histories. The War Between the States proved it. Robbed they might 
have been, subjected to privation, 
\mynote{privation}{(n.): the action of depriving or taking away; the fact or condition of
  being deprived of or cut off from something; deprivation.}
yes, and burned out of hearth 
\mynote{hearth}{(n.): home symbolized as a part of the fireplace as in \textit{hearth and
  home}.}
and home, but NEVER
once had they been set upon by masked outlaws, howling and rapacious 
\mynote{rapacious}{(adj.): giving to grasping or taking for oneself; inordinately greedy. }
Negroes, or
drunken Yankee soldiers who couldn't see straight anyway. And would you
perchance like to know why? Their manners protected them. And those same
standards of conduct would always prevail in the South.

  The Quinsy handbook---a little bluebird-colored affair which bore on its cover
the sphragistic 
\mynote{sphragistic}{(adj.):  Of or pertaining to, relating to or dealing with, seals or signet
  rings. From F. sphragistique or Gr. σϕρᾱγιστικ-ός, f. σϕρᾱγίς seal.}
of a dove rising through hymeneal 
\mynote{hymeneal}{(adj.):  pertaining to marriage. From L. hymenæ-us (also hymenēius), a. Gr.
  ὑµέναιος belonging to wedlock, also as n. a wedding, wedding-song. }
clouds and carrying a banner
with the college motto, "We Preach, You Teach"---codified behavior for the girls.
They were not merely to have a type-and-file appearance. They were asked to wear
white gloves pouring tea, to perfume the wrists, and to maintain custody of the
eyes. They were advised to tithe, 
\mynote{tithe}{(v.): to grant or pay one tenth of (one's goods, earnings, etc.), esp. to
  the support of the church; to pay tithes on (one's goods, lands, etc.).}
to avoid boisterous 
\mynote{boisterous}{(adj.): roughly massive, bulky, big and cumbrous. }
hats, and to use the neglected herb, cerfeuil. 
\mynote{cerfeuil}{(n.):  chervil, a leafy herb, Anthriscus cerefolium, resembling parsley.}
They were asked neither to lisp, 
\mynote{lisp}{(v.): to speak with that defect of utterance which consists in substituting
for (s) and (z) sounds approaching (θ) and (ð); either by reason of a defect in
the organs of speech or as an affectation. Also, loosely, to speak with
child-like utterance, falteringly or imperfectly. 
}
squint, wink, talk loud, look fierce or foolish nor bite the lips, grind the 
teeth, speak through the nose, nor guffle 
\footnote{ \textdbend Probably means to blow on, from \textit{guff}, a puff, 
wiff. The only other place with this usage is found in \textit{The Letters 
of Mercurius}, p.103.
}
their soup. They were encouraged, on the other hand, to sew turkeywork, 
\mynote{turkeywork}{(n.): Turkish tapestry work, or an imitation of this.}
to refer to their young men as "gentlemen callers," and to
accumulate, with a view to future use, egg-frames, salvers, 
\mynote{salver}{(v.): a tray, used for handing refreshments or for
  presenting letters, visiting-cards, etc.}
muffineers, and knife-rests. Above all, they were to familiarize themselves with 
the history of the school.

  The Virginian's was a record of which to be proud. Tradition! Custom!
History!---a meal of fresh heritables 
\mynote{heritable}{(n.): naturally transmissible or transmitted from parent to offspring;
  hereditary. }
all to be washed down with flagons 
\mynote{flagon}{(n.): a large bottle for holding wine or other liquors.}
of the fermented wine of the past. It was a living heritage. There were no 
limits, furthermore, to the historicogeographical importance of Quinsyburg 
itself, and it had long been a matter of great pride to the townsfolk there to 
reflect on the fact that Mr. Jefferson, once stopping by overnight, had found 
the old Timberlake Hotel "clubbable" and that, in 1865, General Robert E. Lee 
with his brave soldiers, on the march from Saylor's Creek to fateful Appomattox,
straggled 
\mynote{straggle}{(v.): to wander or stray from the proper road, one's
companions, etc.; to rove without fixed direction; to go up and down 
dispersedly. 
}
through this very town, at which time the little sisters of mercy in
the college dorms flew with unspeakable horror to the sides of the wounded and
selflessly gave of themselves, cradling 
\mynote{cradle}{(v.): to receive or hold as a cradle.}
their hurt heads, applying cupping glasses, plasters, and bandages, and humming 
strains of "The Bonnie Blue Flag." But today?

  Today was another story. Few even bothered to put flowers on the Confederate
monument anymore. The historical society, its funds dwindling, had been removed
to a room over the theatre. And who ever took time anymore to visit the
rare-book room in Smethwick Library where Miss Pouce, not without effort, had
carefully gathered in a row of glass cases all that Quinsy memorabilia? It was
primarily a collection of old photographs, gum bichromate 
\mynote{bichromate}{(chem.): a compound containing the divalent negative ion $Cr_2 O_7$, usually 
  having a characteristic orange-red colour.}
prints, and bent platinotypes 
\mynote{platinotype}{(n.): platinum prints, also called platinotypes, are photographic prints
made by a monochrome printing process that provides the greatest tonal
range of any printing method using chemical development.
}
preserving the memory of so many dear girls, a thousand blushing apparitions, 
\mynote{apparition}{(n.): the action of appearing or becoming visible.}
who would later go on to make their mark in the world, whether in
the cause of society, Stopesism, 
\footnote{ Marie Charlotte Carmichael Stopes (15 October 1880 – 2 October
1958) was a British author, palaeobotanist and campaigner for eugenics and
women's rights. She made significant contributions to plant palaeontology and
coal classification, and was the first female academic on the faculty of the
University of Manchester.
With her second husband, Humphrey Verdon Roe, Stopes founded the first birth
control clinic in Britain. Stopes edited the newsletter Birth Control News,
which gave explicit practical advice. Her sex manual \textit{Married Love} 
(1918) was controversial and influential, and brought the subject of birth 
control into wide public discourse. Stopes opposed abortion, arguing that the 
prevention of conception was all that was needed.
}
or the suffragettes: 
\mynote{suffragette}{(n.): a female supporter of the cause of women's political enfranchisement,
  esp. one of a violent or `militant' type.}
a group of languorous
\mynote{languorous}{(adj.): full of, characterized by, or suggestive of, languor. Lacking spirit
  or liveliness.}
girls, sitting cross-legged with hockey sticks, staring into the middle distance
with eyes pale as air and jelly-soft cheeks; one dear thing, oversized, rolling
a hoop somewhere; two husky 
\mynote{husky}{(adj.): tough and strong (like a corn-husk); big, strong, and vigorous.}
tsarinas 
\mynote{tsarina}{(n.): the wife of a tsar, a Russian empress. }
posed humorlessly on the old athletic field pointing in mid-turn to a third 
with a faint mustache and a bewildered expression mis-gripping a croquet 
\mynote{croquet}{(n.): a game played upon a lawn, in which wooden balls are driven by means
of wooden mallets through iron arches or `hoops' fixed in the ground in a
particular order.
}
mallet, one high-buttoned shoe poised on a small striped ball; a marvelous 
wide-angle shot, none the worse for time, of forty or so students in bombazine
\mynote{bombazine}{(n.): a twilled or corded dress-material, composed of silk and worsted;
sometimes also of cotton and worsted, or of worsted alone. In black the material
is much used in mourning. 
}
---Quinsy girls all!---trooping like mallards 
\mynote{mallard}{(n.): a wild drake or duck.}
in pious, if pointless, gyrovagation 
\footnote{ \textdbend probably from \textit{gyrovague}, One of those monks who
were in the habit of wandering from monastery to monastery; or from
\textit{vagation}, the action of wandering, straying, or departing from the
proper or regular course; an instance or occasion of this; a wandering,
rambling, roaming; an aberration.
}
along the path of a field called now, as then,
"The Reproaches"; and many many others. ( Miss Pouce had secretly boxed three of
the lot, offensive ones which she kept down with the discards in the basement:
one, a girl in a droopy 
\mynote{droopy}{(adj.): dejected, sad, gloomy, drooping. }
bag swimsuit à la Gertrude Ederle 
\footnote{Gertrude Caroline Ederle (October 23, 1905 – November 30, 2003) was an
American competition swimmer, Olympic champion, and former world record-holder
in five events. On 6 August 1926, she became the first woman to swim across the
English Channel.
}
pitching off a diving board, certain of her parts having been circled in 
neurotoxified purple ink by some poor twisted Gomorrhite 
\mynote{Gomorrhite}{(n.):  An inhabitant of Gomorrah.}
years ago; and two others, shamelessly thumbed,
showing ( 1 ) three girls in chemistry lab smirking 
\mynote{smirk}{(v.): to smile; in later use, to smile in an affected,
  self-satisfied, or silly manner; to simper. }
into the camera while they held up a guttapercha 
\mynote{guttapercha}{(n.): the inspissated juice of various trees found chiefly in the Malayan
  archipelago, now extensively used in the arts.}
object of unambiguous size and shape and (2) the same
girls but one---and she, in the distance, screaming with laughter, and the object
gone.) Smethwick was open until 10 P.M. It would be 9 P.M. on Saturdays, the
rare-book room, of course, by appointment only. Miss Pouce would be ever so
pleased if you came by. Had no one such time for things anymore?

  That was a fair question, for if Quinsyburg had a wealth of anything it was
certainly time, and, beyond that---as the handbook so sagely put it---didn't 
sloth,
\mynote{sloth}{(v.): to waste, pass away (time) in idleness. }
like rust, consume faster than labor wears, with the key that's often used
remaining always bright?

  The girls in that little gradus, 
\mynote{gradus}{(n.):  A handbook used as an aid in a difficult art or practice,
specifically, a dictionary of Greek or Latin prosody used as a guide in writing
of poetry in Greek or Latin.
}
if they paid attention, had their rules for
life. The most trifling actions, they were reminded, if good, increased their
credit but if bad became a matter, when done, no apology could rectify. 
\mynote{rectify}{(v.): to put or set right, to remedy (a bad or faulty condition or state of
  things). }
They were not to fork for bread, dry their underpinnings by the fire, leave
lip-prints on drinking glasses, touch the teeth with the tines 
\mynote{tine}{(n.): each of a series of projecting sharp points on
some weapon or implement, as a harrow, fork, eel-spear, etc.; a prong, spike,
tooth. 
}
of a fork, use rampant 
\mynote{rampant}{(adj.):  of things: Unchecked, unrestrained, aggressive, etc.; esp. of a
quality, belief, state of things, etc.: Having full sway or unchecked course in
the individual or (more commonly) in general society. 
}
witticism, stipple 
\mynote{stipple}{(v.): to paint, engrave, or otherwise design in dots; to produce gradations
  of shade or colour in a design by means of dots or small spots.}
the shower stalls, nor effect shadowgraphs with the
fingers at the Saturday movie, neither were they ever to thrust out the tongue,
sigh aloud, gape, swap underwear, eat fish with the knife, use French words as
that is apt 
\mynote{apt}{(adj.):  likely }
to grow fatiguing, nor cultivate mimicry which was the favorite
amusement of little minds. They must never strike out wantonly 
\mynote{wantonly}{(adv.): in a wanton manner; Recklessly, unadvisedly; without proper excuse or
motive; without regard for right or consequences; in a way that betrays delight
in wrong-doing or mischief, wilfully. 
}
nor snip 
\mynote{snip}{(v.): to cut, to cut up or off, by or as by scissors or some similar
  cutting instrument. }
their nails in public, neither hawk, 
\mynote{hawk}{(v.): to make an effort to clear the throat of phlegm; to clear the throat
  noisily.}
spit, sniff, crack fleas nor drum their fingers.
They were asked not to jerk their hair out of their eyes nor sip audibly nor
effect the branch of a tree for walking purposes nor indicate assent or dissent
by motions of the head, as was the wont of Northern girls. Neither must they
spit on their irons, crunch 
\mynote{crunch}{(v.): To crush with the teeth (a thing somewhat firm and brittle); to chew
  or bite with a crushing noise. }
on cracknels, 
\mynote{cracknel}{(n.):  a light, crisp kind of biscuit, of a curved or hollowed shape. }
use primroses 
\mynote{primrose}{(n.): a well-known plant (Primula veris var. acaulis Linn., P. vulgaris
Huds., P. acaulis Jacq., P. grandiflora Lam.), bearing pale yellowish flowers in
early spring, growing wild in woods and hedges and on banks, esp. on clayey
soil, and cultivated in many varieties as a garden plant. Also, the flower of
this plant. Sometimes extended to include other species of the genus primula. 
}
for floral decoration, say rude things like "Stir your stumps!" 
\mynote{stump}{(n.):  jocularly used for: A leg. Chiefly in 
\textit{to stir one's stumps}, to walk or dance briskly, to do one's duty 
zealously. 
}
or "Tarnation!" or mutter anything
whatsoever disrespectful about the universe. In sharply turning a corner, coming
suddenly in contact with another, they had already abused a right. They were not
to whistle, toast cheese in their rooms, make memorandum knots in their
handkerchiefs, wangle 
\mynote{wangle}{(v.): to move loosely or shakily on its base or in its place of attachment. }
their fingers during conversation, or indulge in parades
of learning, for they would surely live to see verified that a woman who is
negligent at twenty will be a sloven 
\mynote{sloven}{(n.): one who is careless or negligent in respect of dress, personal
  appearance, or cleanliness; an untidy or dirty person. }
at forty, intolerable at fifty, and at
sixty a hopeless mental case. To flee affectation and to be circumspect 
that she offend no gentleman caller in her jesting 
\mynote{jest}{(v.): to utter gibes or taunts; to give utterance to ridicule; to scoff,
  jeer, mock. }
and taunting, 
\mynote{taunt}{(v.): To reproach (a person) with something in a
  sarcastic, scornful, or insulting way. }
to appear thereby of a ready wit, was, above all, paramount.
\mynote{paramount}{(adj.): superior to all others in influence, power, position, or importance;
  pre-eminent. }

  There was, of course, this business of young men, a matter nearest their
hearts because most agreeable to their ambitions, for a Southern girl without a
man was like a pushwainling 
\mynote{pushwainling}{(n.):  A perambulator (A machine for measuring distances, consisting of a
large wheel trundled by a handle along the ground, with attached clockwork and
dial for recording the revolutions; a hodometer. )
}
without wheels. Southern boys, superintended parent
and patron alike, must be polished at all points and trained correctly lest 
\mynote{lest}{(conj.):  Used as a negative particle of intention or purpose, introducing a
clause expressive of something to be prevented or guarded against; = L.
\textit{ne}, Eng. \textit{that $\dots$ not}, for fear that.
}
they flag in those battles for which they were now in the very preparation, 
whether against the enemy agents of this country or in behalf of Southern 
ideals, of which the girls at Quinsy College, and like schools, were the most 
charming synecdoche, 
\mynote{synecdoche}{(n.): a figure by which a more comprehensive term is used for a less
comprehensive or vice versa; as whole for part or part for whole, genus for
species or species for genus, etc.
}
and while there was a generous plenty---an epidemic, in fact---of prep
schools and Presbyterian colleges all over Virginia, anti-intellectual rest
homes which taught overadvantaged quidnuncs 
\mynote{quidnuncs}{(n.): one who is constantly asking: `What
  now' `What's the news'; hence, an inquisitive person; a gossip; a newsmonger.}
how to wear rep ties and smile, the
closest parallel to the female academy, juxtaposing 
\mynote{juxtapose}{(v.): to place (two or more things) side by side,
  or close to one another, or (one thing) by the side of another. }
chivalry with charm, was the military academy.

  Virginia is famous for its many military academies. All, in point of fact, are
one. It is at best a technique mine of Prussian fanaticism, an encampment of
stibnite-colored 
\mynote{stibnite}{(n.): native trisulphide of antimony, `gray antimony', 
  the most common ore of the metal. }
barracks and halls sticking up dolefully 
\mynote{dolefully}{(adv.): in a doleful manner; sorrowfully, mournfully, sadly; drearily,
  dismally. }
in the middle of acres of castrate 
\mynote{castrate}{(adj.): castrated}
lawns, and, as advertised, usually in the rearward pages of national
magazines---always showing either a little chevalier midway over a baffle or some
lost, disappointed boy, too old for his age, staring out in parade dress---it is
invariably named something like Stirrup-and-Halter Hall or St. Bugle's Academy
or Furlongville and run by a man called Colonel Forksplit, a vole-eyed martinet
\mynote{martinet}{(n.): a military or naval officer who is a stickler for strictness of
discipline; hence in wider sense, a rigid disciplinarian. From the name of
General Martinet, a French drill-master of the reign of Louis XIV.
}
with a back straight as a Hepplewhite 
\footnote{ The name of George Hepplewhite (died 1786), who was succeeded by A.
Hepplewhite and Co., used attrib. to designate an English style of furniture of
the latter part of the eighteenth century, characterized by lightness, delicacy,
and graceful curves, being an adaptation of current French styles. 
}
chair and a mouthful of sententious
\mynote{sententious}{(adj.): of discourse, style, etc.: Abounding in pointed maxims, aphoristic.
  In recent use sometimes in bad sense, affectedly or pompously formal. }
stories, all lies, about Stonewall Jackson's boyhood. His charges---you can 
always see them standing alone, glum, 
\mynote{glum}{(adj.):  of persons: Sullen, frowning; having an air of dejection or
  displeasure.}
in the Washington, D.C. bus terminals after holidays---must adhere to the 
regimental uniform which is nothing more than a bit of jerry jingle 
\mynote{jingle}{(n.): a noise such as is made by small bells, a chain of loose links, or
loose pieces of metal when struck; a sound intermediate between clinking and
ringing. 
}
stolen from the Yeomen 
\mynote{yeoman}{(n.): an attendant or assistant to an official, etc. }
of the Guard. But uniforms have plenty of
buttons, and Southern girls, whose adoration for uniforms must be listed, after
amour-propre, 
\mynote{amour-propre}{(n.): self-love which is ready with its claims, and sensitive to causes of
  offence; good opinion of oneself, self-esteem.}
as the most pronounced regional hobble 
\mynote{hobble}{(n.): fig. An awkward or perplexing situation from which extrication is
  difficult.}
against the First Commandment, 
\footnote{The first of the Ten Commandments, which is "Thou shalt have no other
  gods before me".}
always hold you by the button when they speak to you.

  Were Quinsy girls, then, familiarized with all the social graces, schooled to
realize that by a failure of either fashion or forthputfulness 
\mynote{forthputfulness}{(n.):  From \textit{forthput}, put forth.}
it might very well cost them an M.R.S. 
\footnote{ A MRS Degree is a term used to describe when a young woman attends
college or university with the intention of finding a potential spouse.}
degree?

  The handbook, encouraging power without aggression, covered all contingencies.
The caveats were long and letter-perfect. They were never to dip sippets, 
\mynote{sippet}{(n.): a small piece of toasted or fried bread, usually served in soup or
broth, or with meat, or used for dipping into gravy, etc.; a small sop.}
lap
\mynote{lap}{(v.): to take up liquid with the tongue.}
stamps, or chew gum, and upon the occasion of being invited to dine out to wear
dotted Swiss, eat little, and remember that one variety of meat and one kind of
vegetable was the maximum. At least one half of the fare, a sop 
\mynote{sop}{(n.): something given to appease or pacify the recipient; a bribe.}
of grace to gluttony, must be left in the plate. They were told that game 
bones must never be lifted to the mouth nor strenuously 
\mynote{strenuously}{(adv.): in a strenuous manner. \textit{strenuous}(adj.): 
Of action or effort: Vigorous, energetic; now with stronger sense, persistently and ardently
laborious. Of conditions, periods, etc.: Characterized by strenuous exertion.
}
attacked, scraped, 
\mynote{scrape}{(v.): to remove (an outer layer or something
excrescent or adhering) by drawing across the surface the edge of some
instrument held nearly perpendicularly.
}
or twirled. 
\mynote{twirl}{(v.): to rotate rapidly, to spin; to be whirled round
or about; also to turn round quickly so as to face or point the other way; also
fig. of the mind or head: to be in a whirl, be confused or giddy. 
}
They were always to use palliatives 
\mynote{palliative}{(n.): that which gives superficial or temporary
relief; that which serves to alleviate or abate the violence of pain, disease,
or other evil. 
}
when giving opinions of consequence and yet, at the
same time, encouraged to shape the gentlemen callers who were over-saucy 
\mynote{saucy}{(adj.): impertinent or disrespectful, often in a way
  that is regarded as entertaining or amusing; smart.}
with them, or who had small respect in their talk, such an answer that they 
may well understand they were offended with them, not so sharply, however, 
that the escort be irrevocably turned away. They were asked not to crake 
\mynote{crake}{(v.): to utter a harsh grating cry: said of the crow, quail, 
  corn-crake, etc.}
or boast, 
\mynote{boast}{(v.): to extol; to speak of with pride or ostentation; to brag of, vaunt.}
not to use any fond sauciness 
\mynote{sauciness}{(n.): the quality of being saucy, in various senses of the adj. In early
use as a term of serious reprobation: Insolent presumption, haughtiness,
arrogance. Now with milder sense: Impertinence, rudeness to superiors; often
used playfully in mock dispraise 
}
or presumption 
\mynote{presumption}{(n.): the taking upon onself of more than is warranted by one's position,
right, or (formerly) ability; forward or over-confident opinion or conduct;
arrogance, pride, effrontery, assurance. 
}
but, if dancing, say, to dance well without over-nimble 
\mynote{nimble}{(adj.): Quick and light in movement or action; agile,
  active, swift, rapid.}
footings or too busy tricks---and it was advised they neither
reverse in waltzing nor dip.

  Curtseying was encouraged, as long as every girl remembered that there have
been many women who have owed their ruin to an awkward attempt at such. She was
counseled against shingling 
\mynote{shingle}{(v.): to cut (hair), properly so as to give the
effect of overlapping shingles, by exposing the ends of hair all over the head;
also absol. U.S.; to cut (women's hair) so that it tapers from the back of the
head to the nape of the neck; also absol., to have the hair so cut. 
}
the hair, sipping audibly, effecting a need in public for toilet facilities, 
and always, in instance without number, to take heed 
\mynote{heed}{(n.): careful attention, care, observation, regard. (Now chiefly literary.)}
that she give none the occasion to make ill report of her, whereupon, for
example, if she went riding, she must never ride astride, 
\mynote{astride}{(adv.): in a striding position; with the legs stretched wide apart, or so
that one leg is on each side of some object between, as when a person is on
horseback.}
whoop 
\mynote{whoop}{(v.): to utter a cry of `whoop!' or a loud vocal sound resembling this; to
shout, hollo (as in incitement, summons, exultation, defiance, intimidation, or
mere excitement). 
}
untowardly, 
\mynote{untowardly}{(adv.): unbecoming, improper}
or button the third button of the hacking jacket. The carriage of a young lady
must, at all times, be respectful without meanness, easy without familiarity,
genteel without affectation, and insinuating 
\mynote{insinuating}{(adj.): that artfully works his way into company, position, favour, etc.;
  wily, wheedling, ingratiating. Of persons, their manner, etc. }
without design. Finally, they were
asked not to speak of themselves, for nothing could ever be said to varnish
\mynote{varnish}{(v.): to cover or overlay with a specious or deceptive appearance; to gloss
  over, disguise. }
one's defects nor add luster 
\mynote{luster}{(n.):  =lustre, fig. in various applications, esp. Brilliance or 
splendour of renown; glory. Often in phrases, to add lustre to, to shed or throw 
lustre on, etc. Also, splendid beauty (of language, sentiments, etc.). (OED 4.)
}
to one's virtues, whereas, on the contrary, it
would only make the former more visible and the latter more obscure. She who
lived upon talk would die fasting. Good manners will minister to the shop, and
the shop will minister to thee. "Industry," as Stonewall Jackson once
precociously 
\mynote{precociously}{(adv.): In a precocious manner; with premature development. }
lisped to his mother from his bassinet, 
\mynote{bassinet}{(n.): an oblong wickerwork basket, with a hood over one end, used as a
  cradle for babies. Also, a form of child's perambulator of the same shape. }
"needs not wish."

  Quinsy College would not only endure, but prevail. It had before in times of
trial and would again. Its policies, fashioned out of an impatience with this
new age of permissiveness, said it all, with this hope expressed, this
continuity dearly wished, that in such schools---with the laws of both discipline
and decorum 
\mynote{decorum}{(n.): an act or requirement of polite behaviour; a decorous observance.}
meted 
\mynote{mete}{(v.): to apportion by measure; to assign in portions; to
  portion or deal out; esp. to allot (punishment, praise, reward, etc.).}
out by the best teachers of bienséance
\mynote{bienseance}{(n.): French for Decorum; propriety. Fr., f. \textit{bienséant}, f. 
\textit{bien} well + \textit{séant}, pr. pple. of \textit{seoir} to befit. 
Rather common in Eng. use about end of 18th and beginning of 19th century.
}
---there would surely be
an eventual return to the good old American Way. Had times changed? They had,
yes, they had indeed, but if President Greatracks himself, as he so often said,
had to patrol the campus by night in specially made sneakers snooping 
\mynote{snoop}{(v.): to go around in a sly or prying manner. }
for socialists, so be it! Had customs changed? They had, yes, but if one could no
longer catch a glimpse, as in days of yore, 
\mynote{yore}{(n.):  \textit{of yore}: in adj. use: Belonging to time long past, ancient,
  former.}
of Southern belles holding parasols,
\mynote{parasol}{(n.): a light portable screen or canopy carried as a defence from the sun;
a sunshade: used by persons of high rank in the East, and hence, by women in
Europe, etc., in the form of a small light umbrella, often ornamental or gaily
coloured. 
}
wearing frilled 
\mynote{frilled}{(adj.): having, wearing, or adorned with a frill, or something like a frill.
In sewing and dressmaking, a ruffle, frill, or furbelow is a strip of fabric,
lace or ribbon tightly gathered or pleated on one edge and applied to a garment,
bedding, or other textile as a form of trimming.
}
bonnets, 
\mynote{bonnet}{(n.): An article of apparel for the head; `a covering for the head, a hat, a
  cap' (J.)}
and tripping across lawns in fragile blue-and-white
prints with handkerchief-pointed tiers of feminine chiffon 
\mynote{chiffon}{(n.): ornamental adjuncts of a lady's dress, `fal-lals'; colloquially
extended to matters of feminine dress considered as an object of ornament or
display. 
}
and cascades of quivering ruffles, 
\mynote{ruffle}{(n.): a strip of lace or other fine material, gathered
on one edge and used as an ornamental frill on a garment, esp. at the wrist,
breast, or neck.
}
it didn't matter---they were still worshipfully kept alive in
the rotogravure 
\mynote{rotogravure}{(n.): a method of printing by means of a rotary press with intaglio
  cylinders, usu. used at high speed for long print runs. }
section of every true Virginian's heart. Had laxity 
\mynote{laxity}{(n.): Looseness or slackness in the moral and intellectual spheres; want of
  firmness, strictness, or precision. }
set in? Yes, yes, yes, laxity, slackness, 
\mynote{slackness}{(n.): Laxity; want of strictness. }
looseness! It was the age. But if ideals were
honored more in the breach than in the observance nowadays, it took nothing more
than a quick look up at the college rotunda to have one's faith restored, for
there the flag still flew. The American flag, sir!

  The flag, indeed, still flew. And yet a paradox presented itself, for the
democracy which that flag represented was somehow the same democracy that the
admission board, the administration, and the alumni were under jurisdiction,
sometimes, in fact, by federal writ, 
\mynote{writ}{(n.): Law. A written command, precept, or formal order issued by a court in
the name of the sovereign, state, or other competent legal authority, directing
or enjoining the person or persons to whom it is addressed to do or refrain from
doing some act specified therein. 
}
to oblige---resulting in a perceptible latitudinarianism 
\mynote{latitudinarianism}{(n.): Latitudinarian doctrine, opinions, principles, or practice; the
professions or practice of a latitudinarian or the latitudinarians.
\textit{latitudinarian}(adj.): Allowing, favouring, or characterized by latitude
in opinion or action, esp. in matters of religion; not insisting on strict
adherence to or conformity with an established code, standard, formula, etc.;
tolerating free thought or laxity of belief on religious questions;
characteristic of the latitudinarians.
}
which included, among other things, extended curfews, 
\mynote{curfew}{(n.): any regulation requiring people to be off the streets and in their
  homes by a certain time. Also the time when such restriction begins.}
widening of student privileges, faculty raises, the lowering of entrance 
requirements (two new highrise dorms had to be filled) and now a 
consideration for the actual admission of black students!

  The South was trying to rise again. But where was the yeast? 
"Freedom," as President Greatracks had said on many occasions,"is all very 
well and good, but---"
