\chapter{He Enters the House of Rimmon}

\begin{quotation}
My child, how didst thou come beneath the murky darkness, being still alive?\\
\sourceatright{--- \textit{Odyssey}, XI, 155-156}
\end{quotation}
\vspace{0.2cm}

MISS THELMA TRAPPE, spinster, 
\mynote{spinster}{(n.): old maid}
had a pitted 
\mynote{pitted}{(adj.): Having pits or small depressions on the surface; marked 
or spotted with pits; dimpled;}
nose. An ex-schoolmistress from
Quinsy College, having been forced because of age to retire, she could often be
seen walking eccentrically up and down the streets of the neighborhood in her
wide straw hats, peering over her spectacles and repeating like a mantra 
\mynote{mantra}{(n.): a sacred text or passage.}
to the sun, "Let me suffer, just keep shining. Let me suffer, just keep shining."

  One Saturday morning, she simply walked over to Darconville, who was sitting
on a porchswing, and with her little pyewacket 
\footnote{Pyewacket was one of the familiar spirits of a witch detected by the
"witchfinder general" Matthew Hopkins in March 1644 in the town of Manningtree,
Essex, England.}
of a head turned sweetly to the
side asked him, as he was new there, would he care to see the town? A little
walking tour, perhaps?

  It was the first friend Darconville had made in Quinsyburg. An exile herself,
she had come down from New Hampshire many years ago, stayed on to teach, and now
lived by herself in rented rooms at the top of the hill where the loneliness,
she said, always seemed worse. She wore a dress like a teepee, 
\footnote{A tipi(also tepee or teepee) is a cone-shaped tent,
traditionally made of animal skins upon wooden poles.
}
loved frequently
to quote from her favorite literary piece, "Mrs. Battle's Opinions on Whist,"
\footnote{An essay by Charles Lamb(1775-1834). \textit{whist} means a whistling
sound, a whistle.}
and although she once had red hair and fair skin, with the passage of time and
more than periodic though secret infusions of parsnip 
\mynote{parsnip}{(n.): The parsnip (\textit{Pastinaca sativa}) is a root vegetable closely 
related to the carrot and parsley.}
wine they had reversed, creating a face rather like a crabapple. 
\mynote{crabapple}{(n.): wild apple}
It was a glorious day, and so they went off together on the jaunt, 
\mynote{jaunt}{(n.):  an excursion, a trip, a journey, especially one taken for pleasure.}
exchanging confidences freely from the very start.

  Quinsyburg was the county seat. The old courthouse stood behind a short lawn
in the square. The place hadn't changed much since the long-gone days of the
Civil War, and its townsfolk ---ardent lifelong drys --- lived out their small 
agonies or quietly went to the dogs in the proper behind-the-curtains manner 
of shabby genteel 
\footnote{\textit{shabby gentil} means attempting to look genteel and keep 
          up appearances in spite of shabbiness.}
respectability. There was an odor of decay there, of custom, of brittle
endurance, a sort of banality, 
\mynote{banality}{(n.): anything trite or trivial; a commonplace.}
with yet something sinister, waiting below the bleak checkerwork 
\mynote{checkerwork}{(.): sequence of changing fortune, vicissitude.}
of vacant yards, used-car lots, gas stations, and the panmoronium 
\footnote{Theroux's coinage, i.e. an all-encompassing assembly of morons.}
of faded motels (the "Bide-A-Wee," the "Sleepy Hollow," etc.), to
the rooms of which, studiously 
\mynote{studiously}{(adv.):  With studied attention or care; carefully, solicitously, diligently.}
obliterative 
\mynote{obliterative}{(adj.): Having the quality of obliterating; tending to obliterate.
          \textit{obliterate} is a verb meaning to cause to disappear, to 
          efface (anything visible or perceived by the senses).}
of every trace of pretense, came the intrepid 
\mynote{intrepid}{(adj.): Of persons and personal qualities: Fearless; undaunted; 
daring; brave. }
Polos and pan-animated wanderers of America who, in point of fact, had
usually taken the wrong road, missed the right bus, blinked the incidentals of a
highway sign, or somehow got delayed or waylaid 
\mynote{waylay}{(v.): to wait for (a person) in the way and accost;
to stop (a person) in order to converse with him. 
}
at the eleventh hour. It was a
little world unpardonably misled by fundamentalist drivel, 
\mynote{drivel}{(n.): idiotic utterance; silly nonsense; twaddle.}
a stronghold of biblicism, and one drowned in the swamp of its execrable 
\mynote{execrable}{(adj.): of persons and things: Deserving to be execrated or cursed;
abominable, detestable.}
simplicities. Nowhere could be found anything in the way of adornment. It was a 
place that liked its coffee black, its flapjacks 
\mynote{flapjack}{(n.): a flat cake, a pan-cake.}
dry, its adjectives few, its cheeses hard, its
visits short, its melodies whistleable, and its dreams in black and
white --- preferably the latter.

  Darconville was amused to find Miss Trappe setting a good pace, her hat so
large, wagging, 
\mynote{wag}{(v.): to oscillate, shake, or sway alternately in opposite directions, as
something working on a pivot, fitting loosely in a socket, or the like.
}
that she looked like a tip under a plate. "'I like a
thorough-paced partner,'" quoted Miss Trappe, "as Mrs. Battle would say." They
crossed through the nicer part of town, an area of well-treed properties and
rows of colonnaded 
\mynote{colonnaded}{(adj.): having a series of columns arranged at regular intervals.}
houses, handsomely appointed in old brick that served also the formal front walks, 
chimneys, and no-longer-used slave quarters out back. The patrician 
\mynote{patrician}{(adj.): upper class, aristocratic}
section of Quinsyburg was small.

  This was not the doo-dah South of the Camptown Races, 
\footnote{ "Gwine to Run All Night, or De Camptown Races" (popularly known as
"Camptown Races") is a minstrel song by Stephen Foster (1826-1864). 
It was published by F. D. Benteen of Baltimore,
Maryland, in February 1850. Benteen published another edition in 1852 with
guitar accompaniment under the title, "The Celebrated Ethiopian Song/Camptown
Races".}
good bourbon, 
\mynote{bourbon}{(n.):  Whisky of a kind originally made in Bourbon County, 
Kentucky; a glass of this whisky.
}
and the smell of honeysuckle 
\mynote{honeysuckle}{(n.):  any of the many species of arching shrubs and climbing vines of the
genus \textit{Lonicera} in the \textit{Caprifoliaceae} family, many with sweet 
smelling, bell shaped flowers.}
in old shambling 
\mynote{shambling}{(ppl.): that shambles or is characterized by an awkward, irregular gait or
motion. \textit{Shamble} as a verb means to go with an awkward ungainly gait, to
walk awkwardly or unsteadily.}
yards where at dusk one heard the sound of risible 
\mynote{risible}{(adj.): capable of exciting laughter; laughable, ludicrous, comical. From late
          Latin rīsibilis, f. rīs-, ppl. stem of rīdēre to laugh. So F. 
          risible (14th c.), Sp. risible, It. risibile, Pg. risivel.}
Negroes pocking 
\footnote{\textdbend pock=poke ?}
out "Dixie" 
\footnote{ "Dixie," also known as "Dixie's Land", "I Wish I Was in Dixie" and
other titles, is a popular American song. It is one of the most distinctively
American musical products of the 19th century, and probably the best-known
song to have come out of blackface minstrelsy. Although not a folk song at
its creation, "Dixie" has since entered the American folk vernacular. The song
likely cemented the word "Dixie" in the American vocabulary as a toponym for the
Southern United States.}
on hand-hewn 
\mynote{hewn}{p.p. of \textit{hew}, fashioned by hewing (i.e. striking) with a chisel, 
axe, or other tools.} 
banjos. 
\mynote{banjo}{(n.):  a stringed musical instrument, played with the fingers, having a head
and neck like a guitar, and a body like a tambourine; a modification of the
bandore. }
It was far more dreadful and far less eloquent: a kind of cimmeria, 
\footnote{ \textdbend a fictional land of barbarians in antediluvian earth. }
a serviceable huggermugger
\mynote{huggermugger}{(n.): disorder, confusion; a medley, muddle. }
of old wooden tobacco sheds; 
\mynote{shed}{(n.):  a slight structure built for shelter or storage, or for use as a
workshop, either attached as a lean-to to a permanent building or separate;
often with open front or sides. }
auction barns; too many hardware stores; a dismal shoe factory; and a run-down 
dairy bar into whose neon "foot-long hotdog" sign, at night, sizzled 
\mynote{sizzle}{(v.): To make a kind of hissing sound, esp. in the process of frying,
roasting, or burning.
}
bugs blown in by the stale 
\mynote{stale}{(adj.): That has lost its freshness; altered by keeping. }
breezes of the dung-drab 
\footnote{dung(n.) means manure, drab(adj.) means dull. }
Appomattqx River which sludged 
\mynote{sludge}{(v.): to move slowly by solifluction. }
along its fosses 
\mynote{fosse}{(n.):  an excavation narrow in proportion to its length; a canal, 
ditch, or trench;
}
of spatterdock 
\footnote{ Nuphar advena (spatterdock or cow lily or yellow pond-lily) is a
species of Nuphar native throughout the eastern United States and at least some
parts of Canada, such as Nova Scotia.}
and alligator weed 
\footnote{\textit{alligator weed}: Prolific South American aquatic weed having grasslike
leaves and short spikes of white flowers; clogs waterways with dense floating
masses emergent broad leaves and globe-shaped yellow flowers; in sluggish fresh
or slightly brackish water.}
and milfoil. 
\mynote{milfoil}{(n.):  The common yarrow, \textit{Achillea Millefolium.} }
The freight train Darconville had heard but hadn't yet seen chugged
\mynote{chug}{(v.):  To make an intermittent explosive sound as of the escape of exhaust
       gases from an engine cylinder; to move with a sound characteristic of a
       steam-engine or electric motor at work.}
through town once in a while on its way to Cincinnati, but as it had long ago
stopped taking on passengers, the station had fallen into disrepair. The town
came to an abrupt halt at both ends, a foolish watertank marking the limits on
one side and the other giving way to a region of fat-farms and open fields
which, several times a year, suddenly sprouted 
\mynote{sprout}{(v.):  To grow, issue, or proceed as a sprout or sprouts; to shoot forth or
spring up by natural growth.
}
up tents soon to be all faffed 
\mynote{faff}{(v.):  to fuss, to dither; to waste time on an unproductive activity. }
up with the trivialities of the camp-meeting and the chatauqua 
\footnote{\textdbend possible typo of chautauqua --- any of various 
travelling shows and local assemblies that flourished in the United 
States in the late 19th and early 20th centuries. }
harangue, 
\mynote{harangue}{(n.):  a speech addressed to an assembly; a loud or vehement address, a
tirade; formerly, sometimes, a formal or pompous speech.}
the county fair and the vote-rousing picnic. But these were \textit{special}
events.

  During most of the year, the brass-jewelry tastes of its citizens --- you knew
them by string ties, brutal haircuts, and snap-brim hats, 
\footnote{a hat whose brim can be turned up or down}
with fishhooks and lures, 
\mynote{lure}{(n.): a means of alluring animals to be captured; in \textit{Angling} 
    (the action or art of fishing with a rod) a more general term than 
    \textit{bait}, which strictly denotes only something that fishes can eat.}
advertising things like "Funk's Hybrid" or "Wirthmore Feeds" --- ran to
little more than a general enthusiasm for church bake-offs, barbershop gossip,
and all that hand-me-down bumpkinry 
\footnote{\textdbend clownlike act? \textit{bumpkin} means an awkward country
    fellow, a clown.}
touching on Bryanism, 
\footnote{\textdbend William Jennings Bryan (March 19, 1860 - July
26, 1925) was an American orator and politician from Nebraska, and a dominant
force in the populist wing of the Democratic Party, standing three times as the
Party's nominee for President of the United States (1896, 1900, and 1908). He
served two terms as a member of the United States House of Representatives from
Nebraska and was United States Secretary of State under President Woodrow Wilson
(1913-1915). He resigned because of his pacifist position on World War I. Bryan
was a devout Presbyterian, a strong advocate of popular democracy, and an enemy
of the banks and the gold standard. He demanded "Free Silver" because he
believed it undermined the evil "Money Power" and put more cash in the hands of
the common people. He was a peace advocate, a supporter of Prohibition, and an
opponent of Darwinism on religious and humanitarian grounds. With his deep,
commanding voice and wide travels, he was perhaps the best-known orator and
lecturer of the era. Because of his faith in the wisdom of the common people, he
was called "The Great Commoner."
}
vice-crusading, and prohibition. It was a town nonascriptive, 
\mynote{nonascriptive}{(adj.): not attributable, not ascribable.}
nonchalant, 
\mynote{nonchalant}{(adj.): Wanting in warmth of feeling; lacking in enthusiasm 
or interest; indifferent. }
and nonentitative, 
\mynote{nonentitative}{(adj.): not entitative; \textit{entitative} means of the 
nature of an entity; having real existence.
}
one of those places that lent itself to uneasy jokes or gave rise to 
dismissive quips,
\mynote{quip}{(n.): a sharp or sarcastic remark towards a person}
like "I spent a whole week there one Sunday" or "It'd be a great place to live
if you were dead" or "I visited there once, but it was closed."

  "Look," pointed Miss Trappe, coming to a halt. "I never see him without
thinking, for some reason, of my father." The statue of a sinople-green
\footnote{ \textit{sinope} means the green color}
Confederate soldier, so common in Southern communities, stood above them on a
granite pediment, 
\mynote{pediment}{(n.):  referred to L. pēs, pedem ‘foot’, and used for: 
A base, foundation; a pavement.
}
surrounded by cannonballs, with a dapper 
\mynote{dapper}{(adj.):  neat, trim, smart, spruce in dress or appearance.}
Van Dyke beard, a bandolier, 
\mynote{bandoleer,-ier}{(n.): a belt that is worn over the shoulder and across
the body and that holds bullets.}
a rifle-at-the-ready, and a chivalric 
\mynote{chivalric}{(adj.):  of or pertaining to chivalry; chivalrous. }
squint 
\mynote{squint}{(n.):  of looks: Characterized by oblique vision.}
into the heart of the
legitimacy of states' rights, honoring those who died --- so read the
inscription --- "in a just and holy cause." He was the Defender of State
Sovereignty. He stood there in all weather, unphased 
\footnote{ =\textit{unfazed}, not frightened or hesitant; undaunted; not put off}
by birdlime 
\mynote{birdlime}{(n.):  A glutinous substance spread upon twigs, by which birds may be caught
and held fast. 
}
or pigeons. He never flinched. 
\mynote{flinch}{(v.):  To give way, draw back, yield ground in a combat; to draw back or
  turn aside from a course of action, a duty or enterprise. 
}
"My father left us, you know. I was only a child, but almost
died of shame. Oh yes, but that was long ago, and, besides," she sighed, "that,
as they say, was in another country."

  The main buildings of Quinsy College could now be seen across the street, a
cloister 
\mynote{cloister}{(n.): an enclosed place or space, enclosure; close, compass.}
of white columns running along by way of a portico. Disquisiting,
\mynote{disquisite}{(v.):  to make a disquisition. }
somewhat abstractly, on the college's history, Miss Trappe stepped off the
sidewalk. Suddenly leaping back to the curb --- peevishly 
\mynote{peevishly}{(adv.):  in a peevish manner; foolishly, spitefully, perversely, 
skittishly (obs.); with petty vexation or discontent; morosely, querulously, 
petulantly. }
screeching, 
\mynote{screech}{(v.):  to utter a sharp, piercing cry, as of pain or alarm; to scream or call
out with a shrill voice;}
"You!" --- she saved her toes, just, as a green pickup truck with an armament 
\mynote{armament}{(n.):  Military equipments, munitions of war.}
rack at the back window whipped out onto High St. and raced toward Main, pedal 
to the metal. The driver, an underscullion 
\mynote{underscullion}{(n.):  a kitchen servant employed to do menial tasks 
(especially washing).  More commonly referred to as simply a \textit{scullion}. 
}
with a face like a knife, called out something vile.
\mynote{vile}{(adj.):  of actions, conduct, character, etc.: Despicable on moral grounds;
deserving to be regarded with abhorrence or disgust; characterized by baseness
or depravity.}

  Miss Trappe and Darconville continued walking, a strange little mock-up --- a
skeptical Dante, a wizened 
\mynote{wizened}{(adj.):  of persons or animals, their features, etc: Shrunken and dried up,
thin and shrivelled. 
}
Beatrice --- in a most un-paradisaical world: a matchbox-sized theatre, an 
ice-cream shop, and the old Timberlake Hotel, with its chintz 
\mynote{chintz}{(n.):  a name for cotton cloths fast-printed with designs of flowers, etc.,
in a number of colours, generally not less than five, and usually glazed. 
}
curtains, upon whose shaded veranda 
\mynote{veranda}{(n.):  An open portico or light roofed gallery extending along the front
(and occas. other sides) of a dwelling or other building, freq. having a front
of lattice-work, and erected chiefly as a protection or shelter from the sun or
rain. 
}
sat several cut-to-the-pattern townies slumped 
\mynote{slump}{(v.): to lump; to put, place, regard, deal with, etc., as one quantity,
mass, or group. cf. slump in baseball language.} 
in black-lacquered wicker 
\mynote{wicker}{(n.):  a pliant twig or small rod, usually of willow, esp. as used for
  making baskets and various other objects; an osier; a withe. 
}
chairs and several careworn 
\mynote{careworn}{(adj.): showing the wearing effects of overwork or care or suffering.}
arteriopaths, 
\footnote{ \textdbend arteriopathy---any disease of an artery.}
hunched up like angry hawks, fussily 
\mynote{fussily}{(adv.):  in a fussy manner. \textit{fussy} means exacting 
especially about details. 
}
presiding over a game of dominoes. "Percy," came a squawk, 
\mynote{squawk}{(n.):  A loud grating call or cry; a hoarse squall. }
"you ain't got enough strength to pull a greasy string out of a goose's ass." A slam 
\mynote{slam}{(n.):  An insult or ‘put-down’. U.S. slang. }
followed. "Move!" Miss Trappe shook her head. "I had a brother," she said, 
out of the blue, "who always played chess
with me. We wouldn't consider dominoes." She paused. "He married, lost his wife
to another, took to drink ---" A distinct sorrow came into her eyes.

  "And is he ---"

  Miss Trappe made a cataphatic 
\mynote{cataphatic}{(adj.):  defining God positively or by positive statements, opp.
  \textit{apophatic}. From Gr. καταϕατικ-ός affirmative (καταϕῆναι to assent).}
nod. "By his own hand. He was twenty." She
sighed and stumped along. " `The rigor of the game,' " she said, "as Mrs. Battle
would say."

  They came to Quinsyburg's main street. It was a contingent, 
\mynote{contingent}{(adj.):  a gathering of persons representative of some larger group.}
down both sides, of shoulder-to-shoulder shops, a frontage 
\mynote{frontage}{(n.):  Land which abuts on a river or piece of water, or on a road. Also,
  the land between the front of a building and the road, etc. }
dull and repetitious but saved from the blight 
\mynote{blight}{(n.):  any malignant influence of obscure or mysterious origin; anything
  which withers hopes or prospects, or checks prosperity. }
of uniformity by cute mercantile jingles 
\mynote{jingle}{(n.):  a short verse or song in a radio or television commercial or in
  general advertising.}
painted on each window --- the poetic effusions 
\mynote{effusion}{(n.):  a pouring out, a spilling (of liquid); shedding (of tears). }
of various local struldbrugs 
\mynote{struldbrug}{(n.):  Originating from Gulliver's Travels, this name is given to people
  born in Luggnagg who are immortal but continue to age as if they were not.}
and place-proud retailers --- which in small towns, for some peculiar reason, 
become such a rich source of humor:
United Dixiebelle Cup Co. ("Even Our Name Begins with You"); Quinsyburg Bedding
Co. ("We Give You a Lot of Bunk"); 
\mynote{bunk}{(n.):  A box or recess in a ship's cabin, railway-carriage, lodging-house,
  etc., serving for a bed; a sleeping-berth. Freq. one of two or more beds
  arranged in a tier.}
The Old Dominion Outlet ("If Your Clothes Aren't Becoming to You, You Should 
Be Coming to Us"); Stars `N' Bars Exterminating 
\mynote{exterminate}{(v.):  to destroy utterly, put an end to.}
Co. ("All Our Patients Die"); Piedmont Travel Service ( "Please Go Away" ); 
Southside Rug and Linoleum 
\mynote{linoleum}{(.):  A kind of floor-cloth made by coating canvas with a preparation of
  oxidized linseed-oil.}
( "The Best Floor Show in Town" ); The Virginia Shook 
%\footnote{A set of pieces for making a cask or box, usually wood.}
\footnote{‘A set of staves and headings sufficient for one hogshead, barrel, or
the like, prepared for use and bound up in a compact form for convenience of
transport. Boards for boxes prepared or fitted for use and packed in the same
way bear the same name’ (\textit{Cent. Dict.} 1891).
}
Co. ("We'll Stave 
\mynote{stave}{(v.):  To break up (a cask) into staves; to break into and let out the
  contents. }
You In"); The Quinsyburg Gun Shop ("The First to Last"); and The Prince 
Edward Lumber Co. ("May We Strike a Cord for You?")

  The Southern town, a parody of itself, is the prototype from which every other
one is copied. Where is the one, for instance, that doesn't have a radio
announcer named Don Dale; a private white academy; Muddy Creek; a chili-dog
emporium; 
\mynote{emporium}{(n.):  a place in which merchandise is collected or traded in. Often as
  applied to towns or countries: A principal centre of commerce, ‘a mart’. }
the State theatre; Jaycees 
\footnote{ The United States Junior Chamber (JCs or more commonly Jaycees) is a
leadership training and civic organization for people between the ages of 18 and
40.}
with berry-knotted ties; a sheriff called "Goober"; something like the ol' 
Shuckcorn Place (it's always supposed to be haunted by J. E. B. Stuart); a 
popular delivery boy-cum-halfwit named Willis Foster; and a local NRA enclave 
\mynote{enclave}{(n.): a portion of territory entirely surrounded by foreign dominions.}
that meets upstairs in the gunshop every Friday
night to tell lies and make up stories about niggers, nymphomania, 
\mynote{nymphomania}{(n.):  a feminine disease characterized by morbid and 
  uncontrollable sexual desire. } 
and New York City?

  Darconville and Miss Trappe took time for tea at the Seldom Inn ("A Place to
Remember for Cares to Forget") --- a popular meeting place downtown for the
professional tie-and-jacket faction (booths) who rotated matchbooks and told
loud interminable tales and various peckerwoods 
\footnote{ Peckerwood (or simply wood) is a Southern US slang term for
  woodpecker. It is used to describe poor rural whites similar to white trash.}
(stools) who gripcruppered 
\footnote{ \textdbend grip-crupper? \textit{crupper}(n.) originally means a
  strap, looped under a horse's tail, used to stop a saddle from slipping.}
their coffee cups from the non-handle side and stared into a stippled 
\mynote{stippled}{(adj.):  painted, engraved, etc. with dots.}
wall-mirror at their chinless faces and pointed ears. The jukebox 
was blaring 
\mynote{blare}{(v.): of a gramophone, loudspeaker, etc.: to sound loudly
  and stridently. }
country music --- Kitty Wells, "Honky Tonk 
\footnote{ \textit{honky-tonk}: a tawdry drinking-saloon, dance-hall, or 
  gambling-house; a cheap night-club.}
Angels" --- making it impossible to talk, so
Darconville and Miss Trappe together watched through the window as the
Quinsyburg townsfolk passed by, peculiar people on the hop, 
\mynote{hop}{(n.):  a short jump}
remarkably alike all, with faces like the trolls on German beer mugs, 
the curious result, perhaps, of poultry-like 
\mynote{poultry}{(n.): domestic fowl (e.g. chickens, ducks, turkeys and
  geese) raised for food (either meat or eggs).}
inbreeding (farmers, farmers' daughters, farmers' daughters' farmers) 
that had transmogrified 
\mynote{transmogrify}{(v.): to alter or change in form or appearance; 
  to transform, metamorphose (utterly, grotesquely, or strangely).}
a once vital eighteenth-century Protestant Celtic stock into a hedgecreeping 
\mynote{hedgecreeping}{(adj.):  That creeps or sneaks by hedges; clandestine, base; }
lower-class breed of joltheads 
\mynote{jolthead}{(n.):  a heavy-headed or thick-headed person; a blockhead.}
and jusqu' aubouts
\footnote{ \textdbend \textit{jusqu'au bout}, Fr., to the end }
and then metastasized 
\mynote{metastasize}{(v.):  Of a disease, esp. a tumour: to pass from one part or organ to
  another; to undergo metastasis}
into one huge gene pool which seemed to reach from the bulletheaded truckers 
of Mississippi to the triple-named senators of Virginia, slackjawed 
\mynote{slackjawed}{(adj.):  Of a person, overly talkative, indiscreet. }
and malplasmic 
\footnote{\textdbend related to \textit{metaplasmic} (word changing)?}
to a one. It seemed an orgy of kin, with everybody anybody's cousin.

  It was a burlesque 
\mynote{burlesque}{(adj.): of the nature of derisive imitation; ironically bombastic,
mock-heroic or mock-pathetic; now chiefly said of literary or oratorical
compositions and dramatic representations; formerly (quot. 1712) also of
pictorial caricatures. 
}
subordinating individuality to a constant reference of type. 
\textit{Quaeritis habitantes}?
\footnote{ (Latin) “Are you looking for the residents?”}
Rotarians; 
\mynote{Rotarian}{(n.):  a member of a Rotary Club.}
wood-hewing gibeonites; 
\mynote{gibeonite}{(n.): one of the inhabitants of Gibeon who were condemned by Joshua to be
'hewers of wood and drawers of water' for the Israelites (Josh. ix. 27). Hence,
a menial, a drudge. 
}
32° Masons and their ball-jars; 
\footnote{ A Mason jar, named after John Landis Mason who first invented and
patented it in 1858, is a molded glass jar used in home canning to preserve
food. Mason jars are also called Ball jars, in reference to the Ball
Corporation, an early and prolific manufacturer of glass canning jars.
}
pushing tradesmen; zelators and zélatrices; 
\mynote{zelator}{(n.):  zealous defender or supporter; one who zealously 
furthers the cause of. \textit{zélatrice}: a female zelator.}
Odd Fellows of indecipherable worth; Hemerobaptists; 
\footnote{ A Jewish sect which practised daily baptism as a spiritual means of
cleansing from sin; also an early obscure Christian sect who followed the same
practice. 
}
racist Elks (B!P!O.E.) 
\footnote{The Benevolent and Protective Order of Elks (BPOE; also often known as
  the Elks Lodge or simply The Elks) is an American fraternal order founded by
  Joseph M. Norcross in 1868 originally as a social club in New York City, 
  United States.}
and their shovelmouthed wives, usually named Lorinda or Moxone; psalm-snufflers; 
\mynote{snuffler}{(n.): one who snuffles or speaks through the nose; one who speaks 
  cantingly.} 
longnosed umbrella-carrying joykillers; widows with applepandowdy 
\mynote{pandowdy}{(n.):  a kind of apple pudding, variously seasoned, but
  usually with molasses, and baked in a deep dish with or without a crust. }
faces; Volsteaders;
\footnote{ \textit{Volstead}: The name of Andrew J. Volstead (1860–1947),
American legislator, originator of the legislation to enforce prohibition (sense
4) which was passed in 1919 by the U.S. Congress, used attrib. to designate this
legislation or the period during which prohibition was in force. 
}
rattle-toothed almsters; 
\footnote{ \textdbend Probably related to alms?}
gout-footed Shriners; 
\footnote{ A member of the Order of Nobles of the Mystic Shrine, established in
  the U.S. in 1872. Also Shrinite. }
tiny birdheaded clerks in red suspenders; supposititious 
\mynote{supposititious}{(adj.): put by artifice in the place of another; fraudulently substituted for
the genuine thing or person; hence, pretended (to be what it is not), not
genuine, spurious, counterfeit, false. }
chamberers of commerce; pullulating 
\mynote{pullulate}{(v.):  to be developed or produced as offspring, to 
  spring up abundantly.}
boosters; and cretinous, 
\mynote{cretinous}{(adj.):  of or pertaining to a cretin; of the nature of cretinism.
\textit{cretin}: One of a class of dwarfed and specially deformed idiots found
in certain valleys of the Alps and elsewhere. Also in weakened sense (esp. in
form crétin): a fool, one who behaves stupidly. Also attrib. and transf. 
}
peasant-like Colin Clouts 
\footnote{\textit{Colin Clouts Come Home Againe} (also known as 
  \textit{Colin Clouts Come Home Again}) is a pastoral poem by the English poet 
  Edmund Spenser and published in 1596.}
on every street corner who slunched 
\footnote{ A combination of slouching and hunching usually brought on as a
  result of excessive consumption of alcohol or narcotics.}
against poles squinting and chewing down toothpicks in a slow watchful rhythm.

  Growing depressed, Miss Trappe suggested she resume showing Darconville the
town she simultaneously warned him against, arguing, convincingly, that a writer
in staying too long would go mad there. The suicide rate in Quinsyburg, she
said, was --- she stopped and, in the reflection of a window, retied under her 
chin the wide straw hat.

  "High?"

  The crabapple wrinkled. "Astronomical."

  I will stay here for only a year, thought Darconville, and try to do my work.
He told Miss Trappe he'd take the chance, but she told him that Mrs. Battle said
chance is nothing. And yet, he reasoned, wasn't the price for privacy anonymity?
\textit{Un altro, un altro, gran'Dio, ma più forte}.

  They now stood in front of the Wyanoid Baptist Church, a plain white affair
with the usual homiletic 
\mynote{homiletic}{(n.):  The art of preaching; sacred rhetoric.}
menu out front and at the peak of its steeple, spiritual guerdon 
\mynote{guerdon}{(n.):  a reward, requital, or recompense. }
to a whole community, a weathervane 
\mynote{weathervane}{(n.):  mechanical device attached to an elevated structure; 
  rotates freely to show the direction of the wind.}
in the shape of a metal cricket (has anyone ever figured that one out?).

  Quinsyburg, Va. was one of those places where pulpit 
\mynote{pulpit}{(n.): a raised structure consisting of an enclosed platform, usually
supplied with a desk, seat, and other accessories, from which the preacher in a
church or chapel delivers the sermon, and in which in some denominations the
officiating minister conducts the service. }
and drum ecclesiastick were beat with a fist instead of a stick, and whatever 
the persuasion --- whether Wycliffites, Old Order Dunkers, 
\footnote{The Old German Baptist Brethren (Old Order Dunkers) is a conservative Plain church
that emerged from a division among the German Baptist Brethren in 1881 being
part of the Old Order Movement.}
Stundo-Baptists, 
\footnote{an adherent of a Russian religious movement originating in the 19th
  century with the union of various Stundists and Russian Baptists.}
or the International Church of the Foursquare Gospel 
\footnote{The International Church of the Foursquare Gospel (ICFG), commonly
  referred to as the Foursquare Church, is a evangelical Pentecostal Christian
  denomination founded in 1923 by preacher Aimee Semple McPherson.}
--- religion was religion as long as it had been scoured 
\mynote{scour}{(v.):  to remove, get rid of}
of any whim 
\mynote{whim}{(n.): an odd or fanciful or capricious idea}
or wishet 
\footnote{ \textdbend wish?}
that flirted with Rome or ritual or racial equality. It was,
in fact, a reactionary little town filled with stiff-nosed Galatians,
\footnote{ A native or inhabitant of Galatia, an ancient country of central Asia
  Minor. Refered in St. Paul's \textit{Epistle to the Galatians}. }
circuit-riders, and reformers with upsidedown bibles, all looking up hill and
down dale 
\mynote{dale}{(n.):  a valley}
for a chance to save someone's soul. The place teemed 
\mynote{teem}{(v.): to be full, as if ready to give birth; to be prolific or fertile; to
  abound, swarm. Usually cont. \textit{with}.}
with Presbyters 
\footnote{An elder in the Christian church.}
Writ Large, 
\footnote{\textit{writ large}: on a large scale}
and on every Sunday this very church, become a hotbed of tracasseries 
\mynote{tracasserie}{(n.): a state of disturbance or annoyance; a turmoil, 
  bother, fuss; an embroilment, petty quarrel. (Chiefly in pl.) }
and dissent, swelled to overflowing with singing, ringing wonglers,
\footnote{\textdbend Theroux's coinage. }
diehards from the U.D.C. 
\footnote{ \textdbend}
looking for fellowship, and hundreds of bag-in-hand geriatrics 
\mynote{geriatrics}{(n.): an old person receiving geriatric care; a patient in a geriatric ward
  or institution. }
with voices like hoopoes 
\mynote{hoopoe}{(n.): any of several crested Old World birds with a slender 
  downward-curved bill.}
who preferred their theology muscular, their ministers mousy, 
\mynote{mousy}{(adj.):  as quiet as a mouse.}
and their church quite definitely in the majority.

  Darconville bent forward to read the little marquee 
\mynote{marquee}{(n.): a canopy projecting over the main entrance to a building, spec. one at
a theatre, cinema, etc., on which details of the entertainment or performers are
featured. (Orig. in circus use).
}
--- and cocked 
\mynote{cock}{(v.): to raise, to turn up}
an eyebrow. It read:
 
%\begin{verse}
\begin{center}
            Sermon \\
    "Did God Wink?" (Acts 17:30) \\
  W. C. Cloogy, Pastor and Evangelist \\
      Wyanoid Baptist Church \\
        Bethel of Blessings
\mynote{bethel}{(n.):  a hallowed spot; a place where God is worshipped; the 
  pillar that marks such a place. }
%\end{verse}
\end{center}
 
  God help us, thought Darconville, who quite frankly, if somewhat surprisingly,
had yet to be convinced that the Edict of Nantes 
\mynote{edict}{(n.): that which is proclaimed by authority as a rule of 
action; an order issued by a sovereign to his subjects; an ordinance or 
proclamation having the force of law; esp. the edicts of the Roman prætors, and 
subsequently of the emperors, and of the French monarchs. \textit{Edict of
Nantes}, an edict issued by Henry IV of France on 30 April 1598, granting 
toleration to the Protestants; revoked by Louis XIV.}
hadn't perhaps caused more trouble on earth than original sin.

  Darconville and Miss Trappe hadn't gone two feet when a woman came suddenly
shooting out of the side-door of the church. She looked like the wife of a
manciple, 
\mynote{manciple}{(n.):  an officer or servant who purchases provisions for a college, an inn
  of court, a monastery, etc. }
frazzled, 
\mynote{frazzle}{(v.):  to fray, wear out, tear to rags or ribbons.}
with shocked eyes. Clutching 
\mynote{clutch}{(v.): to hold tightly in the bent or closed hand; to hold with a tight or
  determined grasp.}
a fistful of pamphlets, she
identified herself as an evangelist's helper and quickly began batsqueaking
\footnote{ bat+squeak? \textit{squeak}: to emit a high pitch sound.}
about God's love, in support of which topic she swiftly presented to each one of
her little tracts: "Crumbs 
\mynote{crumb}{(n.): a small particle of bread (or other friable food), such as breaks or
  falls off by rubbing, etc.} 
from My Table" by W. C. Cloogy, Evangelist. As all
three stood there, two bewildered, the third --- intense with eyeshine --- 
spontaneously improvised a wee 
\mynote{wee}{(adj.): extremely small, tiny.} 
sermonette 
\mynote{sermonette}{(n.):  a short sermon }
on The Deluge, 
\mynote{deluge}{(n.): the great Flood in the time of Noah.}
she playing Noah, her voice the animals it knew, and the air was soon filled with 
a most ingenious array of barks, oinks, 
\mynote{oink}{(n.):  grunt of pig}
croaks, 
\mynote{croak}{(n): the deep hoarse sound made by a frog or raven.}
snarls, 
\mynote{snarl}{(n.): growl, as of an angry or surly dog, or similar; grumbling sounds}
cheeps, 
\mynote{cheep}{(n.): a faint shrill sound, such as the voice of a young bird or a mouse.} 
and moos, 
\mynote{moo}{(n.): the sound made by a cow or bull}
all articulating 
\mynote{articulate}{(v.): to pronounce distinctly; to utter, give
  utterance to; to express in words. }
the same curious complaint, that this world was too corrupt and wretched to live 
in, the unavoidable implication of which seemed to be that the lesser creatures 
of this earth shared, if not the same size or shape, then at least the same 
agony and accent. And when, she asked, would they make their assent to faith? 
Did they know Jesus for their personal savior? Were they willing to be born 
again?

  And, pray, were they in need of revival?

  Revival? The word sprouted a capital letter. It was bad enough, thought
Darconville, to suggest anything to perfect strangers, but to dare to suggest
one of those punk 
\mynote{punk}{(adj.): devoid of worth or sense; poor in quality; disappointing; nonsensical;
  ‘rotten’. colloq.} 
kick-ups and premillennial 
\mynote{premillennial}{(adj.):  Occurring before the millennium; particularly 
said of the Second Advent of Christ; also, pertaining to the world as it now 
is before the millennium. }
antihomologoumena? 
\footnote{ \textdbend from homologoumena --- books of the New Testament which 
  were acknowledged as canonical by the early church.}
He had a sudden vision of all those bible-thumping
\mynote{thump}{(v.):  to strike or beat heavily, as with the fist, a 
club, or any blunt instrument, producing a dead, dull, somewhat hard sound; 
also, without reference to the sound produced, to hammer, pound, knock 
forcibly. }
wompsters, 
\footnote{ \textdbend Theroux's insults.}
unscrupulous 
\mynote{unscrupulous}{(adj.):  without scruples or principles}
sharpers, 
\mynote{sharper}{(n.): A cheat, swindler, rogue; one who lives by his wits and by taking
  advantage of the simplicity of others; esp. a fraudulent gamester. }
and pigeon-faced decretalicides 
\footnote{ from \textit{Gargantua \& Pantagruel}, Book IV. "So burn 'em, claw 
'em, lop 'em, drown 'em, hang 'em, skewer 'em, bash 'em, rip 'em apart, gut 'em, carve
'em, fry 'em, roast 'em, chop 'em, crucify 'em, boil 'em, broil 'em, quarter 'em,
squash 'em, tear 'em limb from limb and grill 'em: wicked heretics,
decretalifugitives, decretalicides -- worse than homicides, worse than parricides
--- the devil's own decretali-slaughterers." -- translation by Screech.}
who, having weaseled 
\mynote{weasel}{(v.):  to render (a word, phrase, etc.) ambiguous or equivocal; 
  to remove or detract from (its meaning) intentionally. 
  from weasel (n., 鼬); so used because the weasel sucks out the contents of eggs, 
  leaving the shell intact.
}
into the narrow existentialate
\footnote{\textdbend existence?}
of the American South, had for so long impunitively 
\mynote{impunitively}{(adv.): adopting an attitude of resignation towards frustration;
characterized by blaming neither oneself nor others unreasonably. Contrasted
with \textit{intropunitive} and \textit{extrapunitive}. 
}
burked 
\mynote{burke}{(v.):  (British slang) to smother; to conceal, hush up,
  suppress. From William Burke, the name of a notorious criminal executed at 
  Edinburgh in 1829, for smothering many persons in order to sell their bodies 
  for dissection. 
}
reason, honesty, and truth and set up false gods to whom, like rats toward platters 
\mynote{platter}{(n.): a flat dish or plate for food; in later usage often a wooden plate.}
of meazled 
\mynote{meazled}{(adj.): =measled, infected with measles, spotted}
pork, the illiterate \textit{faex populi}
\footnote{ Latin for dregs (i.e. lowest part) of people. \textit{faex} means
  grounds, sediment, lees, dregs of liquids.}
had swarmed only to be bilked, 
\mynote{bilk}{(v.):  to balk (hope, expectation); to cheat, deceive, betray. }
beggared, and buccaneered 
\footnote{ \textdbend \textit{buccaneer}(v.): to engage in piracy against 
  any but one's own nation's ships.}
right on the spot. Was that religion? Miss Trappe, agreeing, said
she would rather take her own life --- at least that way, she added, she would 
not need to be scared anymore about what would happen if she didn't. They walked
away in silence.

  Then Miss Trappe adjusted her spectacles, waited until her optic axes grew
coincident, and took one last painful look to the far end of Main St. She shook
her head.

  "You know," she said, "a thought just crossed my feeble 
\mynote{feeble}{(adj.): of persons or animals, their limbs or organs: Lacking strength, weak,
infirm. Now implying an extreme degree of weakness, and suggesting either pity
or contempt.}
old mind, dear."

  "Yes?"

  "Well," offered Miss Trappe, "the act of committing suicide may be very easy."

  Darconville gently took her arm.

  "When you do it," she said, shrugging and looking up at him with eyes pale as
air, "just simply pretend it isn't you!"

  There was nothing Darconville could find to say, search his heart though he
might. They slipped behind the courthouse and walked through an alley past the
Quinsyburg jail where, high above their heads, they both noticed a series of
black fists gripping the bars of the grills. A lonesome song drifted from one of
the cells across the afternoon.
 
\begin{verse}
      "What a beautiful mornin' that will be,  \\
      \hspace{2em}    Let my people go;  \\
      When time breaks up in eternity,  \\
      \hspace{2em}    Let my people go."
\end{verse}
 
  Miss Trappe halted again on the edge of a thought, the shadow of the building
darkening her face. She looked up at him like Van Eyck's pinchfaced Amolfini 
\footnote{ The Arnolfini Portrait (or The Arnolfini Wedding, The Arnolfini
Marriage, the Portrait of Giovanni Arnolfini and his Wife, or other titles) is a
1434 oil painting on oak panel by the Early Netherlandish painter Jan van Eyck.
It forms a full-length double portrait, believed to depict the Italian merchant
Giovanni di Nicolao Arnolfini and his wife, presumably in their home in the
Flemish city of Bruges.
}
in horizontal hat, and as Darconville again took her arm --- the skin seemed to 
crumble between his fingers, like burial earth --- she stared past him and 
placed a finger on her chin.

  "And you know, the strangest thing of all is," said she, "you may not even
have to pretend."

  It was clear, Darconville now saw, that from her lugubrious 
\mynote{lugubious}{(adj.): characterized by, expressing or causing mourning; 
  doleful, mournful, sorrowful.}
pronouncements Miss Trappe had seen to more terrible depths than the town at first glance
afforded, and, asking her various questions, he began to learn more of what
oppressed her. Quinsyburg was a closed account. It was a place, apparently,
reduced to total irrelevance, to terms that, while having fallen to the category
of the tedious and the negligible, were yet maintained in the hollowness of
their churchianity, the religion, deportmentalized 
\footnote{ \textdbend From \textit{deportment}(n.): manner of conducting 
  oneself}
and moralistic, of the Prodigal Son's brother, and reinforced in a terrible 
irony of reciprocity by the cynical and nonconformist whingeings 
\mynote{whinge}{(v.):  to whine; esp. to complain peevishly.}
of Luther, Melanchthon, 
\footnote{
Philipp Melanchthon (Gr. transl. of Schwarzerd ‘black earth’), a German reformer (1497–1560).
}
Bucer, 
\footnote{
Martin Bucer (early German: Martin Butzer; 1491-1551) was a German Protestant reformer in the Reformed tradition based in Strasbourg who influenced Lutheran, Calvinist, and Anglican doctrines and practices. 
}
Pomeranus, 
\footnote{
Johannes Bugenhagen (1485-1558), also called Doctor Pomeranus by Martin Luther, introduced the Protestant Reformation in the Duchy of Pomerania and Denmark in the 16th century. 
}
Knox, 
\footnote{
John Knox (c.1513-1572) was a Scottish minister, theologian, and writer who was a leader of the country's Reformation. He is the founder of the Presbyterian Church of Scotland.
}
Flacius Illyricus, 
\footnote{
Matthias Flacius Illyricus (Latin; Croatian: Matija Vlačić Ilirik) (1520-1575) 
was a Lutheran reformer from Istria, present day Croatia. He was notable as a 
theologian, sometimes dissenting strongly with his fellow Lutherans, and as a 
scholar for his editorial work on the \emph{Magdeburg Centuries}.
}
and various other sons of revolt. If it seemed, at
first, the kind of town that customarily salted its appeal with a sort of
grassroots neighborliness, the notion didn't last. Insiders were in, outsiders
out. When and if friendship were shown, she said, it was the type of
loving-kindness that uncomfortably verged 
\mynote{verge}{(v.):  to be contiguous or adjacent to; to lie on the
  verge of. Const. on or upon, along.}
on tyranny. They wanted not so much to
convince you of their opinions as to deprive you of your own, and typical of so
many other one-crop regions that had taprooted 
\mynote{taproot}{(v.):  of a plant, to send down a tap-root}
out of the normal world --- pockets of homogeneity bypassed by culture and 
change --- most of its townsfolk were in the grip of an acute xenophobia 
\mynote{xenophobia}{(n.):  a deep antipathy to foreigners.}
that filled the vocabularies of their villageois
\footnote{ village-like}
language with a repetition of paranoid they/them pronouns and convinced them
that beyond the particular borders of their town spread the etceterated 
\footnote{ from etc.}
sloughs
\mynote{slough}{(n.):  A state or condition (esp. of moral degradation) in which a person,
  etc., sinks or has sunk. }
of godlessness where people drank swipes, 
\mynote{swipe}{(n.): poor, weak beer; small beer.}
backslid, 
\mynote{backslide}{(v.):  to slide back, in a figurative sense; to fall
away from attained excellence, esp. of religious faith and practice; to relapse. 
}
and gambled away their
wives. They stuck together in the same way that piranhas, 
\mynote{piranha}{(n.):  a carnivorous freshwater fish of the genus 
  Serrasalmus, belonging to the family Characidae and native to South America;}
seeing certain maculate 
\mynote{maculate}{(adj.):  spotted or blotched}
spots of identification, would not attack each other. Politically, the
community was so far to the Right it jimmied 
\mynote{jimmy}{(v.): to move or force, especially in an effort to get
  something open.}
the Left. In terms of actual religious belief, Quinsyburg's, oxymoronically, 
\mynote{oxymoronically}{(adv.): suggestive of oxymoron; incongruous, self-contradictory.}
was in fact a civic faith, for spiritual obligation had devolved 
\mynote{devolve}{(v.): to cause to pass down by the revolution of time (into some state or
condition). }
to the concept of good citizenry, a quo ad sacra 
\footnote{ \textit{quo ad sacra}: Latin for concerning sacred matters. A quoad
sacra parish is a parish of the Church of Scotland which is not a civil parish.
That is, it had ecclesiastical functions but no local government functions.
}
invariably performed as an endorsement for, and in the name of, the
American Way of Life. Knowledge to them was the parent of malice. 
\mynote{malice}{(n.): the desire to injure another person; active ill-will or hatred. In
mod. use sometimes in weaker sense: Sportively 'mischievous' intent, desire to
tease.}
Ideas they met with derision, 
\mynote{derision}{(n.): the action of deriding or laughing to scorn; ridicule, mockery. }
truth with suspicion, and differences with fear. For any other
way of life one couldn't raise their temperatures a therm. 
\mynote{therm}{(n.): a proposed unit of heat: the quantity of heat required to raise the
temperature of one gramme of water at its maximum density one degree centigrade;
}
They lived, they
knew, the way it was done, and the devil, the great disturber of our faith in
this world, couldn't raise in the townsfolk there one scruple whatsoever leading
to alternatives. They will be there yesterday. They were there tomorrow. They
are.

  Miss Trappe dodged 
\mynote{dodge}{(v.): to move to and fro, or backwards and forwards; to keep changing one's
position or shifting one's ground; to shuffle.}
into a dark overheated little store, bought her paper, and
they continued on past the Quinsyburg post-office --- the walls within painted 
over in flat WPA murals: frigid 
\mynote{frigid}{(adj.): destitute of ardour or warmth of feeling, lacking enthusiasm or zeal;
  cold, indifferent, apathetic; formal, stiff.}
square-jawed men in overalls, holding trowels 
\mynote{trowel}{(n.): a tool consisting of a flat (or, less commonly, rounded) plate of
metal or wood, of various shapes, attached to a short handle; used by masons,
bricklayers, plasterers, and others for spreading, moulding, or smoothing
mortar, cement, and the like.
}
and staring off at horizons --- and cut over by a semi-residential area toward 
the back of the college. The houses, large and desolate, 
\mynote{desolate}{(adj.):  Having the characteristics of a place deserted or uninhabited.}
were all clapboarded 
\mynote{clapboard}{(v.):  to cover or line with clapboards.}
egg-brown affairs with sunken porches overcome with wisteria, 
\mynote{wisteria}{(n.): any of several woody climbing vines, of the genus Wisteria, native to
the East Asian countries of China, Korea, and Japan and the eastern United States.
}
an eruption 
\mynote{eruption}{(n.): the bursting forth (of water, fire, air, etc.) from natural or
artificial limits. }
of domiciliary
\mynote{domiciliary}{(n.):  Pertaining to, relating to, or connected with a domicile or
  residence.}
pasteboards rising up in shingled 
\mynote{shingled}{(adj.):  Arranged tile-wise, imbricated. }
capuches, 
\mynote{capuche}{(n.): the hood of a cloak; spec. that of the Capuchin monks. }
far too close together, and although
the curtains were always pulled one could almost look through them by way of
imagination to see hooded furniture; engravings of stags 
\mynote{stag}{(n.): the male of a deer, esp. of the red deer; spec. a hart or male deer of
the fifth year. }
in the hallways;
perhaps an obsolete oil stove; a clawfoot bathtub upstairs (with elongated
\mynote{elongated}{(adj.): That is excessively long in proportion to its breadth, as if drawn out
or extended. }
orange stains under the faucets); a single bookshelf, with copies of Law's
\textit{Serious Call}, Doddridge's \textit{Rise and Progress}, Orton's
\textit{Discourse on the Aged},
etc., and some poor someone sitting in a calico 
\mynote{calico}{(n.):  a kind of rough cloth, often printed with a bright pattern.}
apron, shelling peas, or all
alone in the darkness of a backroom, desperately praying for forgiveness.

  The streets suddenly were then no longer paved. There were, in fact, no
streets at all, only stamped-out paths of red Virginia soil winding through low
scrubby 
\mynote{scrubby}{(adj.): covered with scrub or brushwood. Also, consisting of or in the form
  of scrub. }
bushland into an outdistrict, poor, rundown, aimless, that dropped away
to an alfalfa 
\footnote{ \textit{Alfalfa}, Medicago sativa, also called lucerne, is a
perennial flowering plant in the pea family Fabaceae cultivated as an important
forage crop in many countries around the world.
}
field ballpark, where the trees were the color of dirty money and
the dust sifted 
\mynote{sift}{(v.): to pass or fall as through a sieve. }
into your shoes like talc. 
\mynote{talc}{(n.):  a fine grained mineral having a soft soapy feel and consisting of
hydrated magnesium silicate; used in a variety of products including talcum
powder.}
Rusted old cars were humped on blocks in narrow driveways. Tarpaper 
\footnote{Tar paper is a heavy-duty paper used in construction.}
hatcheries 
\mynote{hatchery}{(n.): a hatching establishment; spec. one for hatching 
  the ova of fish by artificial means. }
were wired smack 
\mynote{smack}{(n.):  sounding blow delivered with the flat of the hand or something
  having a flat surface; a slap.}
up to the tilted wooden shacks that had either dirt or puncheon 
\mynote{puncheon}{(n.):  A piece of timber with one face roughly dressed, or a split trunk,
  used for flooring and rough building.}
floors, no front doors, and
through the dim breezeways drifted the odors of frying bread, simmering
collards, 
\mynote{collard}{(n.):  a variety of cabbage which does not heart; }
and sweet potato pone. 
\mynote{pone}{(n.):  The bread of the N. Amer. Indians, made of maize flour in thin cakes,
and cooked in hot ashes; now, in southern U.S., any bread made of maize, esp.
that of a coarse or poor kind; also, very fine light bread, enriched with milk,
eggs, and the like, and made in flat cakes.
}
It was the black ghetto.
\mynote{ghetto}{(n.): a quarter in a city, esp. a thickly populated slum area, inhabited by
a minority group or groups, usu. as a result of economic or social pressures; an
area, etc., occupied by an isolated group; an isolated or segregated group,
community, or area.}

  Most of the shacks had no indoor plumbing, and the listing outhouses in the
back of each yard had simply been clapped together with dull, misshapen planks.
\mynote{plank}{(n.): a long flat piece of smoothed timber, thicker than a board; spec. a
length of timber sawn for building or other purposes to a thickness of from two
to six inches, a width of nine inches or more, and eight feet or upwards in
length.}
There were pipe chimneys, makeshift windows, covered in plasti-sheet, and broken
stairs, a pauperization --- the direct result of racism in Quinsyburg --- 
that kept the blacks, because poor, servile. 
\mynote{servile}{(adj.):  of, belonging to or proper to a slave or slaves.}
It was a little world of fatigue, inanition, 
\mynote{inanition}{(n.): the action or process of emptying; the condition of being empty;
spec. the exhausted condition resulting from want or insufficiency of nourishment.
}
and wasted minds. Quinsyburg, only a few years previous, had closed its schools 
for half a decade rather than integrate, simultaneously building a private white
academy, notwithstanding federal pressure, to maintain racial purity. The blacks
were forced into separate schools, separate churches, and even a separate
cemetery. The rents were adjusted: if a black family aspired to fix up its
house, that meant it had money; if it had money, that meant not only a decrease
in servility --- "uppityness" --- but also that it could pay more rent. An 
adjusted rent cured that. It was a "ceiling" theory, for people 
who had no ceilings. Indeed, they hadn't much of anything. 
They lived out their lives as they had for
centuries, cutting up logs, washing toilets, scrubbing doorways, and quietly
knocking on white folks' backdoors with a nickel to ask if the noise of their
old gear-hobbing 
\mynote{hob}{(v.): to create (a gear) by cutting with a hob. }
lawn-mowers would disturb the peace of the nobles inside.

  The afternoon sun began to turn coppery as Darconville and Miss Trappe crossed
downhill toward the ring-road. Old sambos, 
\footnote{ A nickname for a Negro. Now used only as a term of abuse. Also
attrib., esp. with reference to the appearance or subservient mentality held to
be typical of the black American slave.
}
with napkins on their heads, sat on their warped 
\mynote{warp}{(v.): to bend, curve, or twist (an object) out of shape; spec. to curve
(timber) by the application of steam; also, to distort, contort (the body or a
limb, the features). 
}
cane-chairs and waved, while out front little black girls --- their
hair braided in corn-row tight plaits, their legs ashen 
\mynote{ashen}{(adj.): Ash-coloured, whitish-grey, deadly pale. }
--- either played with their pedaps 
\footnote{ \textdbend related to footwear?}
or skipped barefoot, hand in hand, to the Piggly Wiggly 
\footnote{Piggly Wiggly is an American supermarket chain operating in the
  Midwestern and Southern regions of the United States. }
for gumballs. 
\mynote{gumball}{(n.): a ball of chewing gum with a coating of colored sugar.}
A few young men sloped back from the A.B.C. liquor store with bottles of fruit
wine in crinkled 
\mynote{crinkle}{(v.):  to form numerous short twists or turns; to wind
or twist in its course; to contract surface wrinkles or ripples; to wrinkle or
shrink up. 
}
bags and joked as they passed doorways where buxom 
\mynote{buxom}{(adj.): full of health, vigour, and good temper; well-favoured, plump and
comely, 'jolly', comfortable-looking (in person). Chiefly on women.}
young mothers in bandanas, 
\mynote{bandana}{(n.):  kerchief}
looking away as they smiled, rocked their carriages with
one foot and gave pieces of fatback 
\mynote{fatback}{(n.):  salt pork from the back of a hog carcass.}
to their children for pacifiers. There was a
life here that would forever go on unchanged, immutable to pain, to policy, to
the passing of pleasure, and there was perhaps in the constancy of it all, if
finally in nothing else, at least something on which they could depend.

  A loud chorus of "Lift Every Voice and Sing" 
\footnote{
"Lift Ev'ry Voice and Sing" -- often referred to as the "Black National Anthem" 
-- is a song written as a poem by James Weldon Johnson (1871–1938) in 1900 and 
set to music by his brother John Rosamond Johnson (1873–1954) in 1905.
}
could be heard from the small
mal-shingled 
\mynote{shingle}{(v.): to cover, roof (a house, etc.) with shingles (a 
thin piece of wood having parallel sides and one end thicker than the other,
used as a house-tile.)  
}
Negro church that suddenly came into view between several live-oaks
and beyond which, as Darconville and Miss Trappe turned, the two large
dormitories of Quinsy College rose across a half mile or so of woodland. The
bell in the church's squat 
\mynote{squat}{(adj.):  Short and thick; disproportionately broad or wide; podgy; thick-set.}
steeple was bonging 
\mynote{bong}{(v.): ring loudly and deeply}
slowly, a mournful, solitary peal
\mynote{peal}{(n.): the loud ringing of a bell, or of a set of bells; spec. a series of
  changes rung on a set of bells.}
that seemed, echo upon echo, to get lost in the sunshine and become, ironically,
the more forsaken. A group of phantom ladies in weepers, 
\mynote{weeper}{(n.): the long black crape veil of a widow. }
supporting each other,
began to leave the church. They were crying, trying to stifle sobs with their
handkerchiefs. What could have happened?

  Detective, Miss Thelma Trappe stood stock-still in the middle of a hunch. She
went suddenly skewbacked, unflapping 
\footnote{ \textdbend unfasten?}
her copy of the \textit{Quinsyburg Herald}, and ran
her eyes over the front page. She turned to page two. She turned to page three.
She turned to page four, bent down, newt-eyed, 
\footnote{ \textit{newt}: small usually bright-colored semiaquatic salamanders
  of North America and Europe and northern Asia}
and sighed. Looking over her
shoulder, Darconville followed her finger --- for she never said a word ---
to a simple photograph. It was the quizzical 
\mynote{quizzical}{(adj.): of the nature of a quiz or oddity; causing amusement; comical. }
face of a spoon-headed black boy, about
thirteen years old, and underneath were given brief reportorial 
\mynote{reportorial}{(adj.): consisting of, pertaining to, or characteristic of, reporters. }
facts: name, age, address, place of burial. 
He had been struck and instantly killed the day
previous by a hit-and-run pickup truck at the corner of Main and High streets.
The several witnesses from the Timberlake, dominoists by avocation, could give
no description of the driver, so the sheriff closed the investigation. It was
only God's way, he no doubt felt, of turning a leaf in the Book of Eternal
Decrees.
\mynote{decree}{(n.): an ordinance or edict set forth by the civil or other authority; an
  authoritative decision having the force of law. }

  "Precisely," said Miss Trappe.

  Darconville heard the irony but kept counsel, feeling only the vibrations of
her increased step in awkward silence as she stumped along homeward, pausing
only to eye the sky at intervals from under her wide hat as if searching the
heavens for any indication of justice or balance, a clue, of any degree, to the
moderating power of the universe, some kind of proof of the stamp celestial.
They were soon not far from the street where they'd first begun their walk when
Darconville's companion decided to speak,

  "I lost my stepmother in the same way," said Miss Trappe, her little mouth
trembling. She tried, unsuccessfully, to shape her hands --- carelessly hopping
up --- to an attitude of resignation. "She was coming home at dusk with several
just-packed jars of autumn honey and was run down by a passing motorist, who
simply drove on." Her eyes were filling up, her hat quivering."She was not
killed right away --- the report from the hospital was that she kept feverishly
repeating, `my honey. My honey.' Just that. `My honey.' I mustn't fail to tell
you, Darconville," sobbed Miss Trappe, "I was neither the pretty nor the favored
child, but how I ran --- " She gasped to breathe out pain. "--- ran to her."
She smiled up through a runnel 
\mynote{runnel}{(n.): a small stream of water; a brooklet, rivulet, rill, or trickle. }
of tears. "Hearts," she said, "was Mrs. Battle's
favorite suit." And wiping a tear from her pitted nose, she repeated, "`My 
honey.' "She swallowed. "I wanted to believe, you know, that she was asking for 
me."

  "She \textit{was} calling you."

  Miss Trappe lifted up two little gerbil 
\footnote{A gerbil is a small mammal of the order Rodentia. Once known simply
as "desert rats", the gerbil subfamily includes about 110 species of African,
Indian, and Asian rodents, including sand rats and jirds, all of which are
adapted to arid habitats. 
}
eyes and simply shook her head. Her nose was dripping. 
\mynote{drip}{(v.): to let (a liquid) fall in drops; to let fall
  (drops; rarely other objects). }
And as Darconville took her hand, soft as bird's-eye, she heaved 
\mynote{heave}{(v.): to lift with exertion (something heavy); to 
  raise with effort or force; to hoist.}
up the sorrow weltering 
\mynote{welter}{(v.): to be sunk or deeply involved in. (OED 2.c.)}
in her heart. "No," she wept, "she said she never
loved me, then \textit{screamed} it, Darconville, laying there like dead metal, 
she sere 
\mynote{sere}{(adj.):  dry, withered}
---" Darconville's heart almost misgave. 
\mynote{misgive}{(v.): of one's 'heart', mind, etc.: To suggest (to
one) doubt or apprehension; to cause to be apprehensive (that); to incline to
suspicion or foreboding.
}
"I was s-so ashamed," she whispered into his chest with a tiny humiliated 
voice, shattering with sobs, "\textit{I completely w-wet myself, all over.}"

  Weeping to her feet, she suddenly turned her face up defiantly 
\mynote{defiantly}{(adv.): In a defiant manner; with defiance; daringly. 
  \textit{defiance}: the action of defying.}
toward the sky and looked with a hideous grimace 
\mynote{grimace}{(n.): a distortion of the countenance whether spontaneous or involuntary,
expressive of some feeling (esp. annoyance, embarrassment, ill-humour or pain)
or tending to excite laughter; a wry face.  
}
into infinity, as if to say not balanced yet, O crafty universe, not balanced 
yet.

  It became impossible to think. Darconville could say nothing: so overcome with
pity, he could find no words adequate to consolation and knew beyond reason that
any poor stuttering attempt on his own part must fall terribly short, for
touching things sometimes can only be felt, and yet before he could make any
show of what he felt, she kissed him on the neck, turned her little crabapple of
a head toward the westering 
\mynote{wester}{(v.): of the sun, moon, or a star: To travel westward
  in its course; to draw near the west.}
sun, and then disappeared over the hill like a dot.
 
