\chapter{President Greatracks Delivers}

%\vspace{4cm}

\begin{quotation}
"He hath builded towers of superarrogation 
\mynote{arrogation}{(n.): the action of claiming and assuming without just
  reason; unwarrantable assumption.}
in his owne head."  \\
\sourceatright{--- Gabriel Harvey, Pierce's Supererogation
\mynote{supererogation}{(n.): transf. and gen. Performance of more than duty or circumstances
  require; doing more than is needed. (OED 1.b.) From super+erogation (pay out).}
}
\end{quotation}
\vspace{0.2cm}

  "College! There's a little word for you, dear girls, college: neatly
pronounced, pronounced as spelt, and correctly spelt s-a-c-r-i-f-i-c-e---the
battle cry of the United States of God Bless America! And who's telling y'all
this, some up-spoutin' 
\mynote{spouting}{(n.):  declamation or recitation; speech-making, speechifying. }
no account dingdong 
\mynote{dingdong}{(n.):  a tumultuous party or gathering. colloq.}
in a string tie pointing his head out
in the direction of his face and looking for to impress you? Exactly wrong! It's
a man who knows what sacrifice means, and costs! Now, button back your ears,
little cousins, for I only aim to say it once---I know all the tricks, every last
one of them, and unless someone out there can tell me how a brown cow can eat
green grass, give white milk, and make yellow butter, she best just sit down
tight and listen, OK? I been evywhere but the moon and seen evything but the
wind. I been to the edge of the world and looked over. I mean, I can fight,
shout, win, lose, draw, turn on a dime, and meet you coming back for change, you
hear? Good, now you mark time on that and we gone get somewhere."
 
  IT WAS PRESIDENT GREATRACKS, the college headmaster, a man fat as a Fugger: 
\footnote{ \textdbend The Fugger family is a
German family that was a historically prominent group of European bankers,
members of the fifteenth- and sixteenth-century mercantile patriciate of
Augsburg, international mercantile bankers, and venture capitalists. 
}
a bun, a ham, 
\mynote{ham}{(n.):  An inexpert performer; (also ham actor, ham actress) an ineffective
  or over-emphatic actor, one who rants or overacts. slang (orig. U.S.). }
a burgher. 
\mynote{burgher}{(n.):  a middle class citizen}
He was a charming and resourceful academic illiterate,
politically appointed, his brain a \textit{pot-au-feu}
\footnote{ a French beef stew}
of boomism, 
\footnote{ \textdbend related to boom?}
bad grammar, and prejudice, his face very like that of the legendary Leucrota 
\footnote{ The crocotta (or corocotta, crocuta, or leucrocotta), is a mythical
dog-wolf of India or Ethiopia, linked to the hyena and said to be a deadly enemy
of men and dogs.
}
whose mouth opened as far as its ears. The school auditorium was decked 
\mynote{deck}{(v.): to clothe in rich or ornamental garments; to cover
  with what beautifies; to array, attire, adorn.}
out for the opening
assembly, with all eyes fixed on the keynote speaker. He showed the conviction
of a roundhead, 
\footnote{ A member or adherent of the Parliamentary party in the Civil War of
  the 17th century, so called from their custom of wearing the hair close cut.}
and, having traversed the dais 
\mynote{dais}{(n.): a platform raised above the surrounding level to give prominence to
  the person on it.}
with an oafish 
\mynote{oafish}{(adj.):  of the nature of an oaf, dull-witted, stupid, loutish. }
and peasant-like lumber 
\mynote{lumber}{(n.):  useless or cumbrous material.}
that betrayed his grim-the-collier 
\footnote{ A species of hawk-weed (Hieracium aurantiacum). Also The name of a
character in an Elizabethan play (modernized as 'Grim the Collier of 
Croydon' (1662). 
}
background, he bulked now over the lectern around which had been slung 
\mynote{sling}{(v.):  to place in, or secure with, a sling or slings in
order to admit of or facilitate hoisting or lowering; to raise up or let down by
means of a sling or slings. 
}
a banner lettered in blue and white
 
\begin{center}
            QUINSY   \\
           WELCOMES  \\
             YOU
\end{center}
 
\noindent
and then hamfistedly 
\mynote{hamfistedly}{(adv.): lacking physical movement skills, especially with the hands.}
fussbudgeted 
\mynote{fussbudgeted}{(adj.): complained or fussed a great deal, especially about unimportant
  matters;}
back and forth in a suit the color of sea-fowl guano, 
\mynote{guano}{(n.):  A natural manure found in great abundance on some sea-coasts, esp. on
the Chincha and other islands about Peru, consisting of the excrement of
sea-fowl. 
}
wagging a finger like Elijah the Tishbite and trooping out his dockets, 
\mynote{docket}{(n.): the calendar of a court; the list of cases to be tried or a summary
  of the court's activities.}
posits, 
\mynote{posit}{(n.): a statement which is made on the assumption that it will prove
  valid.}
and quiddits 
\mynote{quiddit}{(n.): the real nature or essence of a thing; that which
  makes a thing what it is. }
like a costermonger 
\mynote{costermonger}{(n.): an apple-seller, a fruiterer; esp. one that sold his fruit in the
open street. Hence, b.b Now, in London, a man who sells fruit, vegetables, fish,
etc. in the street from a barrow.
}
his pippins.
\mynote{pippin}{(n.):  The name of numerous varieties of apple, raised from seed.}
 
  "When I was a little wagpasty 
\footnote{ \textdbend =wagpastie (a rogue)? }
of a lad back in Free Union, Va. in them days of
the Depression, which you wouldn't know about, my tiny ol' mammy wore galoshes,
\mynote{galosh}{(n.): a waterproof overshoe that protects shoes from water
  or snow}
used thorns for fishhooks, and buck-washed me in a hopper. 
\mynote{hopper}{(n.):  a funnel-shaped or hopper-like hollow. }
We were so poor we couldn't even pay attention, and it was a dang 
\footnote{ =damn}
holiday, nothin' else, just to go and play stoopball 
\mynote{stoopball}{(.):  A game played by bouncing a ball off a step or stoop.}
with the swivel-eyed 
\mynote{swivel-eyed}{(adj.):  squint-eyed, squinting;}
halfwit next door or hop along down to the grocery to splurge 
\mynote{splurge}{(v.): to spend money extravagantly. Freq. const. on.}
on a box of penny chicle."
\mynote{chicle}{(n.):  the milky latex of the sapodilla tree, used to make chewing gum.}

  "I worked the nubs 
\mynote{nub}{(n.):  a small lump or protuberance}
off my little fingers for wages the coloreds laughed at, trundled 
\mynote{trundle}{(v.): to move along on a surface by revolving; to roll. }
out of bed at dawn like a filthy sweep, blinking and looking for the
life of me just as white and hairless as a egg, a little eyesore in my mussed
\mynote{muss}{(v.): to make untidy; to crumple, to ruffle; to smear,
  mess; to entangle, confuse. Also with up. }
overhalls and my sweater hind-side-to---so low I had to reach up to touch
bottom---and not a soul, but for mammy, who had a good word to say to me. But you
are what you are and you ain't what you ain't, right? We thanked the good Lord,
and thanked Him and how! Schoolrooms? Who said anything about schoolrooms?
Shoot, no one thought of them, sir! Only thing I knew was my 15¢ Dicky Deadlight
picture book---that is, until mammy took me in tow to learn me my devotions and
reading. And if I didn't oblige the dear thing? Why, she caned me scarlet in the
attic and set me to kneel on peppercorns for punishment, and if you think that
tickles! 
\mynote{tickle}{(n.): an act of tickling, in various senses of the vb.; a touch that
  tickles; a tickling sensation; a tickled or pleasantly excited feeling. }
I want to tell you, many's the night---O, it seems like yesterday!---I sat
up into the wee hours in my ripped jim-jams, 
\mynote{jim-jam}{(n.):  pl. Fantastic ways, peculiarities. colloq. (OED 2.) }
my eyes pinched tight from
candlesmoke, going over and over again the sentences in my mustard-colored copy
of Edward Clodd's \textit{Tom, Tit, Tot}. But, dagnabbit, 
\mynote{dagnabbit}{colloq. Expressing annoyance or amazement: `Darn it!'}
I got my sums, didn't I? I got my Bible, didn't I?"
 
  Greatracks rose up like a huge fat glyptodont, 
\mynote{glyptodont}{(n.): an extinct South American quaternary mammal, allied to the
  armadillos, with tessellated scales and fluted teeth.}
capitalizing every word with
his voice. He chop-gestured. He beckoned 
\mynote{beckon}{(v.) OED. 1.  intr. To make a mute signal or significant gesture 
with the head, hand, finger, etc.; now esp. in order to bid a person approach.
}
to the ceiling. He took oaths and blew
air and circled his arms, all with a jumped-up and inquisitorious duncery 
\mynote{duncery}{(n.): the state, character, or practice of a dunce or dullard; intellectual
  dullness, stupidity. }
that thumbfumbled 
\footnote{ =thumb+fumble? }
truth and opened up a museum of bygone pictorial mediocrities which
magnified puddles 
\mynote{puddle}{(n.): a small body of standing water, foul with mud, etc. or with a muddy
bottom, now always shallow, as those left in depressions of the ground in a road
or footpath after rain; a small dirty pool.
}
by rhetoric into blue fairy lakes and fobbed 
\mynote{fob}{(v.):  trans. To cheat, deceive, delude, trick, impose upon, ‘take 
in’; also with up. Also, to fob (a person) of or out of (something). colloq.
}
off hawks for handsaws.
\mynote{handsaw}{(n.): a corruption of heronshaw or hernsew, dial. harnsa, heron. Only found
in Shakespeare, \textit{Hamlet}, ii. ii. 367, I am but mad North, North-West: 
when the Winde is Southerly, I know a Hawke from a Handsaw.}
 
  "Now, I'm telling you all this for a mighty good reason, girls---and paying it
no mind, you'll find me as unsociable a creature as ever chawed gum, I
guarandamntee you! At Quinsy College here, it's either fish or cut bait. Ain't
enough insurance in this here world better than perspiration, 
\mynote{perspiration}{(n.): the excretion of moisture through the pores of the skin (originally
  applied to the insensible excretion, now also to the sensible); sweating. }
which means nothin' more than good ol' sweat to right-thinkin' folks, which is 
just what you better be aimin' to be, hear? Yes? Then stand on it! I mean, I 
want everybody---evahbody!---on automatic here! You mope. 
\mynote{mope}{(v.): (archaic) to act in a dazed or stupid manner}
You get bounced. You get up to
tomfoolery. You gone tom-foolerize yourself right out into the street, see?"

  "Why, only last year, now to mention it, one of those self-important little
undergraduate cinderbritches with a cheek rubied like a dang poppy pads into my
office fixin' to have at me, see, stands there wheel-high to a rubbish truck,
and then comes out with, `I'm fed up with college!' I remember standing there
amazed, thinking what the sam hill? What the tarnation 
\mynote{tarnation}{as adj. Damned, damnable, execrable. }
hell? O no. Said she was
hangin' it up. Said, shoot, she couldn't care a straw for book-learnin', not me,
no, I'm a big shot---failin' to realize, course, feelin' sorry for herself and
walkin' around on her lip, that what she really wanted was a 7 X 9 patch over
her mouth! The poor little trapes, 
\mynote{trape}{(n.):  a messy or untidy woman.}
obviously humble of brain, mistook college
for the handball court or your Five County Fair. I about died---dahd! Well, 
I read her the dang riot act for starters and then helped her into tomorrow 
with a kick strong enough to knock a billygoat off a gut-wagon! 
She spelled college f-u-n,
see, which is no kind of spelling at all. It's mis-spelling is what it is.
Mis-dang-spelling! And if you yourself ain't wrapped too tight, I best remind
you to prepare for the same kind of treatment, girls, or there ain't a hog 
\mynote{hog}{(n.): a swine, pig}
in Georgia, `cause it's only fifteen inches between a slap on the back and 
a swift kick on the place you wear no hat!"
 
  Perspiring, 
\mynote{perspire}{(v.): of a person (or the animal body): To give out
watery fluid through the pores of the skin. Originally of insensible
perspiration; later including sensible perspiration or sweating. (Now the
ordinary sense.) 
}
President Greatracks---wailing 
\mynote{wail}{(v.):  to utter persistent and bitter lamentations or
  complaints.}
like a preaching jebusite 
\mynote{jebusite}{(n.):  Name of a tribe of Canaanites, dispossessed of Jerusalem by David. In
  17th c., a nickname for Roman Catholics, esp. Jesuits. }
and orthonational
\footnote{ \textdbend Theroux's coinage.}
---drove in with his industrie-protestant creedal 
\mynote{creedal}{(n.):  =credal, Pertaining to or characterized by a creed, or formula of
  religious belief. }
formulations: a clattering 
\mynote{clatter}{(v.):  to make or emit a rapid succession of short
sharp noises in striking a hard and dry body; to rattle. Said either of the
material instruments or the agent. }
windmill of cheerings up, pep 
\mynote{pep}{(n.colloq.): vigor, energy, spirit, forcefulness, c.f. pep talk.}
tonics, 
\mynote{tonic}{(n.): fig. an invigorating or bracing influence. (OED 1.b.)}
%Pertaining or relating to tone or accent in speech;
%indicating the tone or accent of spoken words or syllables; characterized by
%distinctions of tone or accent.}
across-the-fence chat, and general protrepticos, 
\footnote{ \textdbend possibly related to \textit{protreptic}(n.), a book,
writing, or speech intended to exhort or instruct; an exhortation, instruction.
May also refered to Protrepticus (Greek: Προτρεπτικός), a philosophical work
by Aristotle that encouraged the young to study philosophy.
}
fanning, in crisis, all those who, in asking, seeking, or
knocking would receive, find, or have it opened unto them. The volume of gas
increased, according to physical principle, as his temperature did the same.
 
  "At Quinsy, we have no truck whatsoever with those so-called Northern, well,
views and, that being the case, aren't so all-fired anxious for revolution,
evolution, devolution or any other damn lution y'all want to come up with! I've
seen the lot of them, sneakin' around the campus at night with beards like
Stalin had and sandwich-boards broadcastin' `Freedom!' or `Down with God and His
Saints!'---all of them socialists, won't-work liberals, and bleedin' heart
sombitches---high-steppin' like coons 
\mynote{coon}{(n.):  i.e. the racoon (Procyon lotor), a carnivorous animal of North 
  America. }
and tryin' to turn this place redder than a hawberry 
\mynote{hawberry}{(n.):  山楂. Crataegus (from the Greek κράτος strength 
and άκης sharp, referring to the thorns of some species) commonly called 
hawthorn, thornapple, May-tree,whitethorn,or hawberry, is a large genus of 
shrubs and trees in the family Rosaceae, native to temperate regions of the 
Northern Hemisphere in Europe, Asia and North America.
}
with their radical Commie bushwa! 
\mynote{bushwa}{(n.):  (informal) rubbish; nonsense. May also mean bourgeois.}
Sowin' discord! Havin' into the coop!
Huh? Prit-near right, ain't I? You better know it. Shoot, I could quote you
chapter and verse!"
 
  His fat body shook like a balatron, 
\mynote{balatron}{(n.):  A buffoon, a contemptible fellow. }
as if his soul, biting for anger at a mouth inadequately circumferential, 
\mynote{circumferential}{(adj.): of, pertaining to, or of the nature of the circumference. }
desired in vain to fret 
\mynote{fret}{(v.): to eat} 
a passage through it. He blated. 
\mynote{blat}{(v.):  to bleat, or make similar sounds. Also fig., to talk 
  noisily or impulsively. }
He blaterated. 
\mynote{blaterate}{(v.): to babble or talk vainly.}
He blaterationed. 
\footnote{ \textdbend From \textit{blateration} (babbling chatter)? }
Out blasted a flash of oratorical n-rays and impatient oons 
\mynote{oons}{(n.):  A petty oath.}
while the echoes of his voice, pitched high, strident, 
\mynote{strident}{(adj.):  Making a harsh, grating or creaking noise; loud and harsh, shrill. }
like the hellish sounds of Vergil's Alecto, 
\footnote{ Alecto (Ἀληκτώ, English translation: "the implacable or
unceasing anger") is one of the Erinyes, or Furies, in Greek mythology. In the
Aeneid (Book 7), Juno commanded Alecto to prevent the Trojans from having their
way with King Latinus by marriage or besiege Italian borders. Alecto's mission
is to wreak havoc on the Trojans and cause their downfall through war. To do
this, Alecto takes over the body of Queen Amata, who clamors for all of the
Latin mothers to riot against the Trojans. She disguises herself as Juno's
priestess Calybe and appears to Turnus in a dream persuading him to begin the
war against the Trojans. Met with a mocking response from Turnus, Alecto
abandons persuasion and attacks Turnus with a torch, causing his blood to "boil
with the passion for war". Unsatisfied with her work in igniting the war, Alecto
asks Juno if she can provoke more strife by drawing in bordering towns. Juno
replies that she will manage the rest of the war herself: You're roving far too
freely, high on the heavens' winds, and the Father, king of steep Olympus, won't
allow it. You must give way. Whatever struggle is still to come, I'll manage it
myself. (Virgil, Aeneid, trans. Robert Fagles, Book 7, ll.646-649).
}
drumbeat through the auditorium and went right to the pit of the stomach.
 
  "You see that majestic piece of dry goods with the stars-and-stripes hanging
yonder? 
\mynote{yonder}{(adv.): at or in that place; there; usually implying that the object
  spoken of is at some distance but within sight: Over there, away there. }
That speak to you of revolution? The deuce, I say! And it aggravokes 
\footnote{ \textit{aggravoke}=aggravate + provoke. South Dial.}
me like you wouldn't believe to see these pseudo-intellectual puddingheads
---every one of them dumb as a felt boot---buddyin' up to Moscow! Well, 
put you in mind, we don't hold with this down South here. Eskimos eat the 
refuse out of their pipestems! 
\mynote{pipestem}{(n.):  the stem of a tobacco-pipe.}
Japs fry ice-cream! Them little puck-faced Zuni Indians from Mexico
drink their urine! Polish dogs bark like this: `Peef! Peef!' And instead of
sayin' hello in Tibet I'm told the poor jinglebrains just stick out their
tongues and hold up their thumbs! That mean we do it? Huh? Think!"

  "The Southran way, cousins, is the way we aim to follow. Item: we study here.
Item: we won't walk around here lookin' like boiled owls. Item: we'll be
sticking it through until we ain't got enough strength to blow the fuzz off a
peanut, and then we'll work some more! Thread and thrumme! 
\footnote{ \textit{thrumme}: =thrum? a short piece of waste thread or yarn 
  (including the unwoven ends of the warp). Thus \textit{thread and thrum}: good
  and bad together. }
Don't study and your chances of stayin' here are between slim and none
--- and slim is on a plane-ride to
Tahiti, you got it? I see any of them irritating thimbleheads, 
\mynote{thimble}{(n.): a small hard pitted cup worn for
  protection on the finger that pushes the needle in sewing.}
house-proud pippins, 
\mynote{pippin}{(n.): the name of numerous varieties of apple, raised from seed. }
and intellectual willopus-wallopuses 
\footnote{ Originating in the lumberjack communities of the developing United
States of America, the \textit{willopus-wallopus} is one of the fearsome
critters (i.e. creatures).
}
around here with signboards and
complaints, and it's goodbye Quinsy, hello world, and that's a promise, sisters,
that is a promise! You have to get up early, remember, to get out of bed. Now, I
always close with a quote from my favorite author of books, one Arthur de
Gobineau, 
\footnote{ Joseph Arthur, Comte de Gobineau (14 July 1816 - 13 October 1882) was
a French aristocrat who was best known by his contemporaries as a novelist,
diplomat and travel writer but is today most remembered for developing the
theory of the Aryan master race and helping to legitimise racism by scientific
racist theory and racial demography. 
}
a European person who once said, `Attaquez! Attaquez! Attaquez!' which
means attack, of course---in French. Gaze boldly into the past and put the future
behind you. Don't let your brains go to your heads. You'll thank yourself
someday---don't mind thanking me, I don't count. Now welcome to Quinsy College,
hear?"
 
  It was a rhetoric that would have taxed 
\mynote{tax}{(v.): To burden; to make serious demands upon; to put a
  strain on. }
Quintilian 
\footnote{ Marcus Fabius Quintilianus (c. 35 - c. 100 CE) was a Roman
rhetorician from Hispania, widely referred to in medieval schools of rhetoric
and in Renaissance writing.
}
himself: a few final admonitions, 
\mynote{admonition}{(n.): an act of admonishing; a warning, reproof; an utterance or statement
  of grave counsel or censure, esp. of ecclesiastical censure. }
accompanied by several rumplestiltskinian 
\footnote{=rumpelstiltskinian. \textit{Rumpelstiltskin} is the name of a 
  vindictive dwarf in German folk-tale, used allusively. }
stamps of anger---for the
particular hardcore few who, he thought, could not understand an order
unillumined by force---emphasized the need at the school of what his very manner
contravened, 
\mynote{contravene}{(v.): of things, actions, etc.: To run counter 
  to, be contrary to, come in conflict with.}
but this was by the by, for he had clearly argued himself into a
state of such broad magisterial cheek that he was virtually beyond not only the
accusation of such vulgarity but also beyond its being adduced, 
\mynote{adduce}{(v.): to bring forward (verbally) for consideration, to
  cite, to allege.} 
in the same way that, philosophically, at the exact moment of offense defense is 
clearly immoment. 
\mynote{immoment}{(adj.):  Of no moment; trifling. }
Not Berosus 
\footnote{ \textit{Berosus} was a famous Chaldean astrologer who flourished
during and after the lifetime of Alexander the Great. The Greeks honored Berosus
by erecting a statue of him with a tongue sculpeted from gold, signifying the
truth of his forecasts. Berosus was also a historian and wrote a Greek-language
work in three books on the history and culture of Babylonia. Pickover, 
\textit{Dreaming the Future: The Fantastic Story of Prediction}, p.251.
}
with tongue of gold was he, neither silver-throated Solon,
\footnote{Solon (Greek: Σόλων; c.638--c.558 BC) was an Athenian statesman,
lawmaker, and poet. He is remembered particularly for his efforts to legislate 
against political, economic, and moral decline in archaic Athens.}
rather a moody-sankeyan 
\footnote{
Dwight Lyman Moody (February 5, 1837 - December 22, 1899), also known as D. L.
Moody, was an American evangelist and publisher connected with the Holiness
Movement, who founded the Moody Church, Northfield School and Mount Hermon
School in Massachusetts (now Northfield Mount Hermon School), the Moody Bible
Institute, and Moody Publishers.
Ira David Sankey (August 28, 1840 - August 13, 1908), known as The Sweet Singer
of Methodism, was an American gospel singer and composer.
}
yammerer 
\mynote{yammer}{(v.): to make a loud unpleasant noise or outcry; to
  howl, yell; to roar, shout.  }
from the old school who, finishing now, wound
down to the conclusion that made up in volume what it lacked in finesse. He
jerked his head forward with one last glare, beady 
\mynote{beady}{(adj.): beadlike; (of eyes) small, round and glittering.} 
as a vole's, 
\mynote{vole}{(n.):  Any of various small mouselike rodents of the family Cricetidae
(especially of genus Microtus) having a stout short-tailed body and
inconspicuous ears and inhabiting fields or meadows.
}
then picked up his clatter 
\mynote{clatter}{(n.): Noisy talk; confused din of voices, gabble. Often applied
  contemptuously to what is treated as mere empty gabble.}
of clenches, 
\mynote{clench}{(n.):  In various senses of the vb.; e.g. a grasp, grip, clutch; tight
closure of the teeth, fist, etc.; secure fastening of a nail, etc.; fig.
conclusive confirmation of an argument, etc. (OED II.3.)
}
abstersives, 
\mynote{abstersive}{(n.): Something cleansing; detergent; abstergent.}
and céphalalgies 
\footnote{ those affected with a headache.}
and thumped out into the wings on his monstrous feet.

  One daring little beast in the back row frowned, 
\mynote{frown}{(v.): to express disapprobation or unfriendliness by a
  stern look. }
held her nose, and said, "Puke."
 
 
