\chapter{Bright Star}
 
\begin{quotation}
  The unthrift 
\mynote{unthrift}{(adj.): Prodigal, spendthrift.} 
sun shot vital gold, a thousand pieces.
\sourceatright{--- HENRY VAUGHAN, Silex Scintillans}
\end{quotation}
\vspace{0.2cm}
 
  THE CLASSROOM was old. It seemed in dark and incongruous 
\mynote{incongruous}{(adj.):  not congruous, disagreeing in charactor or qualities.}
contrast to the
delicate femininity it both isolated and yet protected with a fastness like that
of some battlemented watchtower. On the wall hung a portrait of the Droeshout
\footnote{ The Droeshout portrait or Droeshout engraving is a portrait of
William Shakespeare engraved by Martin Droeshout as the frontispiece for the
title page of the First Folio collection of Shakespeare's plays, published in
1623.
}
Shakespeare and a canvas map of Britain, pocked 
\mynote{pock}{(v.): To mark with pocks, or (fig.) with disfiguring spots.}
with red pins. This was English
100, a freshman section of girls who were almost all dressed à la négligence in
the present-day fribble-frabble 
\footnote{ \textit{fribble}: A trifling, frivolous person, one not occupied in
serious employment, a trifler. \textit{frabble}(n.): confused wrangling.
}
of fashion, mostly jeans and wee 
\mynote{wee}{(adj.): a little}
pannikins.
\footnote{ \textdbend a small metal drinking vessel.}

  Darconville strode 
\mynote{stride}{(v.): to walk with long or exteded steps.}
in and sat down. He placed his books on the desk, and, as
he smiled, the girls straightened around to squeaks, the click of shoes, the
scent of earth-flowers. The moment was immediately memorable, for instantly
aware at the corner of his eye of a sparkle, the fluorescence, of a jewel, he
looked up with sudden confusion, as if bewildered to discover art in nature's
province. It was she: a faery's child, the nameless lady of the meads, 
\footnote{  =meadow}
full beautiful, sitting in the front-row seat at the far right with her eyes 
lowered to the desk in a kind of fragrant prayer, her chin resting gently on 
the snowy jabot 
\mynote{jabot}{(n.):  An ornamental frill on a woman's bodice. }
of her blouse and her hair, tenting her face, golden as the Laconian's.
Prepared for her, he saw he really wasn't. The heart in painful riot omitted
roll-call.
\mynote{roll-call}{(n.): OED 1.a. The act of calling over a list of the names of
persons forming a military or other body, in order to ascertain who are present;
the marking of such a list at a particular time. Also (U.S.), a calling over of
a list of members of a legislative or similar body in order to ascertain how
each wishes to vote on a particular measure.} 

  "Shall we look at the Keats?" asked Darconville, quietly. It had been their
first assignment: to analyze one poem. There was a marked self-consciousness in
the straightening of shoulders, in the coughs, as the students settled down
resolutely to consider the poem.

  "As one must pronounce a Chinese ideograph in order to understand it," said
Darconville, "so also must a poem be read aloud. Would anybody care to do so?"
He waited.

  Silence.

  "Anyone?"

  The girls remained earnestly hunched over their books, submissive to the idea
that obscurity can be found in the solemnity of well-aimed concentration.

  "Anyone at all?"

  The linoleum snapped.
\mynote{snap}{(v.):  to fracture or break apart suddenly.}

  Darconville was amused. It seemed like vesper 
\mynote{vesper}{(n.): of or belonging to, characteristic of, occurring in, the evening;
  vespertinal}
hour at the Shaker Rest Home for Invalid Ladies. The discomfort was palpable, 
\mynote{palpable}{(adj.):  That can be touched, felt, or handled; apprehensible by the sense of
touch; tangible, sensible. From \textit{palp}(v.), to touch, feel; to handle
gently.
}
with not an eye on him. Then a hand shot up.

  "Miss---" Darconville looked at his roll-book.

  "Windt," the student provided.

  It was a girl in the third row who resembled Copernicus, the shape of her
pageboy, 
\mynote{pageboy}{(n.): A shoulder-length hairstyle with the ends of the hair curled under.}
its two guiches 
\footnote{ A spiral-shaped curl of hair upon the forehead.}
coming forward like tongs 
\mynote{tong}{(n.): An implement consisting of two limbs or 'legs' connected by a hinge,
pivot, or spring, by means of which their lower ends are brought together so as
to grasp and take up objects which it is impossible or inconvenient to lift with
the hand.
}
and swinging at the jawline, making her look small as a creepystool. 
\footnote{=\textit{creepie}, Sc. and dial., a low stool. }
She turned to one of her girlfriends for confidence, then stood up, and began.
 
\begin{verse}
        "Bright star! would I were steadfast as thou art---  \\
        Not in lone splendor hung aloft the night,  \\
        And watching, with eternal lids apart, \\
        Like Nature's patient, sleepless, um---"
\end{verse}
 
  "Eremite," said Darconville. "Hermit."

  Rachel Windt, wrinkling her nose, squinted 
\mynote{squint}{(v.): OED 2.a. To look with the eyes differently directed; to
  glance obliquely or in other than the direct line of vision; also, to glance
  hastily or casually, to peep. }
at the word. Darconville, repeating the word, prodded 
\mynote{prod}{(v.): fig. To goad mentally; to stir up, instigate,
incite; to irritate. 
}
her.

  "E-eree-ereem---"

  "Just pronounce the consonants," Darconville said, laughing, trying to relax
her, "and the vowels will fall into place." Two girls in the back row exchanged
cold glances; they didn't find that particularly funny. Rachel Windt
bewilderedly twisted up a noil 
\mynote{noil}{(n.): The short pieces and knots of wool combed out of the long staple. }
of her hair, shrugged, and continued.
 
\begin{verse}
        "The moving waters at their priestlike task  \\
        Of pure, um, ablution 
        \mynote{ablution}{(n.): the act or process of washing clean, originally as a
                   religious rite.}
          round earth's human shores,  \\
        Or gazing on the new soft fallen mask  \\
        Of snow upon the mountains and the moors
        \mynote{moor}{(n.): A tract of unenclosed waste ground.}
        ---"
\end{verse}
 
  "Perfect. Let's stop there," said Darconville, "where the octave 
\mynote{octave}{(n.):  the first eight lines of a poem}
and sestet
\mynote{sestet}{(n.):  the last six lines of a poem}
hinge. 
\mynote{hinge}{(v.): to attach or hang with or as with a hinge.}
This is a sonnet---composed, if you'll notice, in one sentence---in which the
poet expresses a wish. And what, let's ask ourselves, is that wish?" He paused.
"Anybody?"

  "He's wishin' he was a star?"

  Wishin'. Your thlipsis 
\mynote{thlipsis}{(n.): compression, and especially constriction of vessels by an external
cause. Oppression. From Gr. θλῖψις pressure, compression.}
is showing, thought Darconville, looking in the direction of the voice.

  It was a long-nosed piece of presumption 
\mynote{presumption}{(n.): The taking upon onself of more than is warranted by
  one's position, right, or (formerly) ability; forward or over-confident opinion
  or conduct; arrogance, pride, effrontery, assurance. }
in the last row, wearing an armory 
\mynote{armory}{(n.): a collection of weapons and material}
of Scandinavian nail-jewelry, who was less concerned with a Romantic poet's 
tragic wishes than after-shampoo flyaways if the comb she simultaneously 
shuttled through her hair to the repetitious snaps 
\mynote{snap}{(n.): A quick or sudden closing of the jaws or teeth in biting, or of
  scissors in cutting; a bite or cut made in this way.}
of chewing gum meant anything.
Darconville smiled and kindly suggested that she was, even if a trifle so,
somewhat wide of the mark. Snap. Another hand? Anyone?

  "Miss?"

  This voice came from the direction of---but it was not hers. Darconville, with
no small difficulty, fixedly tried to keep his eyes focused on this student
alone, a moon radiant only for the proximity of that adjacent sun whose beauty,
even in a condition of reflected light, seemed startling enough to destroy his
sight on the instant.

  "Trinley Moss," answered the girl, standing up, her six bracelets jangling, 
\mynote{jangle}{(v.): To make a discordant or unmusical noise; to
  sound or 'jingle' harshly or discordantly. }
in an opera blue turtleneck sweater and plaid wrap-around, safety-pinned at the
thigh. "Well, I believe he's right unhappy," she said, "and that this star he
mentions, symbolizin' brightness, is drivin' this poor child here, I don't know,
to wishin' he could either touch it or, I don't know, just plain ol' flop down
and become a"---she shrugged---"a eremite?"

  Darconville began to feel like St. Paul, watching errors creep in at Colossae.
It wasn't all that bad, hardly a matter for despair. He took teaching to be a
mission, not a trade, and sought to avoid the by-the-numbers Drillschule
\footnote{ \textdbend German for school?}
technique that, for one reason or another, had apparently long been held in
vogue there by various pompous 
\mynote{pompous}{(adj.): Characterized by an exaggerated display of self-importance or dignity;
boastful, vain-glorious, arrogant; consequential, pretentious, ceremonious; of
language: inflated, turgid.
}
doodles 
\mynote{doodle}{(n.): colloq. a silly or foolish fellow; a noodle.}
and burgraves-of-fine-print 
\mynote{burgrave}{(n.): The governor of a town or castle; later, a
noble ruling by hereditary right a town or castle, with the adjacent domain. 
}
on the Quinsy faculty, teachers, for whom students were the little limbs of 
Satan, who bored the girls to distraction repeating opinions seamed by 
cankerworm 
\mynote{cankerworm}{(n.):  A caterpillar that destroys buds and leaves. }
and reading from lecture notes long since parched 
\mynote{parch}{(v.): To dry to extremity, to make hot and dry; to
scorch, said esp. of the action of the sun's heat, or of fever or thirst.}
by the Dog Star. 
\footnote{The Dog Star, Sirius, in the constellation \textit{Canis Major}
   (Greater Dog)}
He would be patient. At Quinsy it was an essential requirement, for although 
the state constitution laid claim to its being a school of higher learning, 
he from the very first day of arrival wondered just what level they were 
measuring from, and if they took salt tablets when doing so.

  Darconville ranged the room with his eyes for a better answer--- they couldn't
as yet settle on la femme d'intérieur---and noticed a girl, her pencil poised
above her notebook, a response in her eyes, an answer on her lips. He nodded to
her.

  "I am Shelby Uprightly," she said, neatly pressing her glasses to her
zygomatic 
\mynote{zygomatic}{(adj.): pertaining to zygoma (the cheekbone) }
arch with one finger. "John Keats in his sonnet, `Bright Star,'
wishing to be steadfast, as he states, but not alone, is quite painfully
expressing, as he often did, the particular tension of feeling he suffered when
torn between a desire for the Ideal, the lovely star, in this case, and the
Real, the girl mentioned in the last lines, someone he must have---"

  "Excellent. You outdo yourself," exclaimed Darconville. "And would you read
those lines?"

  "I'm sorry, sir. I'm afraid I won't," she replied, confidentially lowering her
voice. "I do not read well enough for declamation." She tilted her head. "Is
that acceptable?"

  Darconville was nodding but continued to listen intently 
\mynote{intently}{(adv.): In an intent manner; with strained attention or close application;
  earnestly, eagerly. }
long after the girl
had finished speaking, as if struggling with an idea encountered, not in the
classroom, but at the innermost of his being, one he was unable to
comprehend---then, suddenly embarrassed, he looked up from the fanciful call of
that mysterious girl in the right front row whose beautiful hair, like that of
the Graces, enshrined the face he couldn't see. Who then, asked Darconville
quickly, would like to read the lines? He found a girl sitting directly in front
of him, bent down, and smiled into her eyes.

  "You are---?"

  She was speechless.

  It was Millette Snipes, a girl whose comic little face, perfectly round,
looked like a midway balloon. She blushed, tapping one of those
pencils-with-funheads, and with a farm-bright smile looked up with huge oval
eyes that told instantly just how deeply she'd been smitten 
\mynote{smitten}{(ppl.adj.): That has been smit; beaten, struck. }
with love for her teacher.

  "Shall I thay the whole poem, thir?" she asked, brushing aside a cinnamon 
\mynote{cinnamon}{(n.): A yellowish-brown colour, the color of cinnamon.}
curl that had stuck to her nose.

  A wave of giggles broke over the class, until Darconville, soberly, tapped the
desk once with his pointer like Orbilius 
\footnote{Lucius Orbilius Pupillus (114 BC - c. 14 BC) was a Latin grammarian of
the 1st century BC, who taught a school, first at Benevento and then at Rome,
where the poet Horace was one of his pupils. 
}
plagosus. 
\mynote{plagosus}{(adj.): Latin for fond of beating or striking.}
Then Scarlet Foxwell, her eye
inset with something of a humorful character, leaned over to show her where to
begin, but not before glancing back at DeDonda Umpton and Elsie Magoun who were
both pinching their noses to keep from exploding with laughter.

  Then Millette Snipes, her elbows akimbo, 
\mynote{akimbo}{(adj.): Of the arms: In a position in which the hands rest on the hips and
  the elbows are turned outwards.}
her book high, spoke in a voice from the family Cricetidae.
\footnote{ A taxonomic family within the order Rodentia - the hamsters, voles,
  lemmings and some mice and rats. }
 
\begin{verse}
        "No---yet thtill thteadfatht, thtill unchangeable, \\
        Pillowed upon my fair love'th ripening breatht,  \\
        To feel forever itth thoft fall and thwell,  \\
        Awake forever in a thweet unretht---"
\end{verse}
 
  "Crisper, child," said Darconville.

  Inevitably, she spoke louder.
 
\begin{verse}
        "Thtill, thtill to hear her tender-taken breath,  \\
        And tho live ever---or elthe thwoon to death."
\end{verse}
 
  The girl tenderly pressed the open book to her bosom and looked up soulfully,
her wopsical 
\footnote{ \textdbend commonly quoted from T.S. Eliot "In a coat of fur / And a
porpentine cat / And a wopsical hat: / How unpleasant to meet Mr. Eliot! /
(Whether his mouth be open or shut)."
}
eyes moist and glowing. Darconville thanked her.

  "Very good. Now, the poet here, as Miss Uprightly earlier pointed out, aspires
to the star's steadfastness but, you see, not at the cost of hermit-like
loneliness. He is in a state of suspicion about what must be sacrificed in the
pursuit of the Ideal. It is the problem, perhaps," said Darconville, "of the
paradoxical man who wants to move, say, but not migrate. Fanny Brawne, a
coquette 
\mynote{coquette}{(n.): A woman (more or less young), who uses arts to gain the admiration
and affection of men, merely for the gratification of vanity or from a desire of
conquest, and without any intention of responding to the feelings aroused; a
woman who habitually trifles with the affections of men; a flirt
}
of sorts, was the woman Keats loved, and while she was not specifically
the subject of the poem---it could be anyone---the poet wanted to live forever 
with whomever it was, such was his passion, or else swoon 
\mynote{swoon}{(v.): To fall into a fainting-fit; to faint. }
to death."

  "Thwoon in the thenthe of thuccumb?"

  "Precisely, Miss---"

  "Thnipeth."

  "The Romantic, you see," said Darconville, "is a man of extremes."

  "It'th tho thimple," said Millette Snipes, "a thad poem, but tho very very
thimple."

  "The greatest lines in English poetry," replied Darconville, "are always the
simplest." He lifted a lectern onto his desk and leaned forward on it. "I have
something for you---call it, if you will, `the Principle of Trim.' The poet, you
see, is rather like the seagull who has a perfect process for desalting water,
although no one can explain how it works: he baffles out of a given line, if
he's the enchanted metaphorsician we hope he is, that which otherwise would
directly explain it."

  Darconville wanted to make this clear.

  "The simple line seeks to outwit, not merely resist, the complexity of thought
it noncommittally grows out of and, by definition, filters out the ideas it must
nevertheless, to be great, always raise, do you see?"

  As he felt the presence of that one girl near him, he listened distractedly to
the voice that didn't seem his. Who are you? thought Darconville. Who are you?
He tried to concentrate his whole nature in one terrific effort to summon up,
that he might dismiss it for once and all, the formidable magnetic mystery on
his right that drew him so relentlessly off the subject.

  "And therefore the greatest lines always imply the longest essays, discourses,
metaphrases which the poet quite happily leaves to the agitation of critics,
schoolmen, all those academic Morlocks 
\footnote{ A member of a troglodyte cannibalistic race described in H. G. Wells'
  novel \textit{The Time Machine}.}
who study the brain of a line after its
face has come off in their hands." Darconville looked around. "Do you want me to
repeat that?" (cries of "yes, yes")

  Darconville thought for a moment. "The ring of a mighty bell," he explained,
"implies, logically, a technical if less melodic success in the shift and whirr
\mynote{whirr}{(n.): A continuous vibratory sound, such as that made by the rapid
fluttering of a bird's or insect's wings, by a wheel turning swiftly, or by a
body rushing through the air.
}
of its gudgeons, 
\mynote{gudgeon}{(n.): A pivot or spindle on which a bell or other object swings or rotates.}
a secondary sound that can be muffled 
\mynote{muffle}{(v.): OED 4. To envelop or wrap up (oars, a drum, bell, etc.) so
  as to deaden the sound. }
only in proportion to the genius of the campanologist."
\mynote{campanology}{(n.): The subject of bells; detailed examination
  of the principles of bell-founding, bell-ringing, etc. }

  Glycera Pentlock, throwing an exasperated 
\mynote{exasperate}{(v.): To irritate (a person); to provoke to anger;
  to enrage, incense.}
glance at her neighbor, Martha Van Ramm, recklessly resumed scribbling away in 
her notebook. But Ailsa Cragg, chumbling 
\mynote{chumble}{(v.): To gnaw, nibble, peck. }
the tip of her pen, gasped and blurted 
\mynote{blurt}{(v.): To utter abruptly, and as if by a sudden impulse;
to ejaculate impulsively; to burst out with. }
out, "What is a campanologist?"

  "Use," said Darconville, "your definitionary. It's one of the last few
pleasures left in life."

  Poteet Wilson looked aggravated. 
\mynote{aggravated}{(adj.): Exasperated, incensed, irritated, provoked. }
Wroberta Carter spun 
\footnote{ p.p. of spin}
her pencil. And Barbara
Celarent whispered a slang word which, though unbeknownst to her, was once held
in common usage by the daughters of joy in fifth-century Babylon.

  "Language," continued Darconville, "often disguises thought. The successful
poetic line, never discursive, is always thought extensively profound---and
surely, upon the examination that is our burden, will be---but at its very 
essence it is the furthest thing from it. Beauty only implies truth, as
\textit{fuit} implies \textit{est}. The poet doesn't say so, however; the 
factophile 
\footnote{ \textdbend presumbly a lover of facts}
does. Poetry is the more
imperfect when the less simple. The nature of great poetry---a perfect art
attempted by imperfect people--- prevents the direct concern with truth and can
only suggest by singing what it secretly shields by showing. The greater the
simplicity stated the greater the complexity implied, you see?" Several girls
put up their hands. "And putting aside for now the implications this has in the
human personality, in affairs of the heart, know only that the reader is left to
assume whether or not what the poet stated in Beauty he knew as Truth."

  A yawn was heard, followed by the jingling of some Scandinavian nail-jewelry.

  "Bear with me? God alone is pure act, never in potency, whereas finite things
are metaphysically composed of act and potentiality. But, in poetry, Beauty is
act, you might say, Truth the potency involved by inference in that act, and of
course there follow many attendant questions that can be raised: will Beauty
wane 
\mynote{wane}{(v.): To decrease, opposite to wax.}
as Truth waxes? 
\mynote{wax}{(v.): to grow, increase, opposite to wane.}
Is one the synonym, or the antonym, of the other? And what
of the old idea that Beauty, meeting Truth, must always corrupt it---the Beauty
that turns Truth, as Hamlet said, into a bawd? 
\mynote{bawd}{(n.):  pimp, procurer, pander }
What are the implications---the
essential question must be asked---of the Ideal?" A heavy inert apathy settled
over the class. "We must consider the final lines of another poem of Keats, 
`Ode on a Grecian Urn.' "

  The students, exhaling sighs, actually seemed to decompress. It didn't matter.
Darconville, perhaps not even aware of it himself, was busy sorting out matters
in the depths of his own mind.

  "`Beauty is Truth, Truth Beauty': I call this Keats's Comfort---he is saying
his prayers and, embarrassingly, is overheard. It is a correct epistemological
\mynote{epistemological}{(adj.):  pertaining to epistemology, the theory or science of the method or
  grounds of knowledge.}
utterance, you might say," continued Darconville, "made aesthetically incorrect
at the very moment it is introduced into the poem. Beauty is Truth, Keats
forgot, only if and when the didactician 
\mynote{didactician}{(n.):  one who follows a didactic method, a didactic writer; one who writes
  with the aim of instructing.}
is there to say so by whatever means he
can. Arguably, Beauty is Truth precisely when it does not say so, just as, turn
about, a holy man disproves his holiness as soon as he asserts it. Ironically
enough, then, this, the most famous line of Keats, while starkly simple---indeed,
it is generally accepted as the classic example of such---denies exactly what a
simple line must be valued for, asserting what should be implied, attesting to
what it can't. The line expresses what it does not embody---or, better, embodies
what it has no right to express: a beautiful messenger appears delivering ugly
news. It's a paradox in the same way that---"

  Darconville paused, turned to the blackboard, and chalked out a wide
rectangle. He wrote something and backed away for perspective.
 
%        --------------------------------
%        | This boxed sentence is false |
%        --------------------------------
\begin{center}
\fbox{This boxed sentence is false}
\end{center}
 
  "The paradox here is clear," asserted Darconville, "is it not? If the sentence
is true, then it is false; if false, then it is true." Various eyes roved 
\mynote{rove}{(v.): to look in various directions; to wander.}
in troubled scrutiny 
\mynote{scrutiny}{(n.):  OED 3. The action of looking searchingly at something; a searching
  gaze.}
over the board. "Do you see? A liar says that he lies: thus he
lies and does not lie at the same time. The \textit{jeu de mots}
\footnote{ French: play on words}
is fun grammatically,"
warned Darconville with a smile, "but imagine how vulnerable would be your
philosophical calm if this translated into the behavior of a human being?
Careful," said Darconville, "with your boyfriends."

  The girls tittered.
\mynote{titter}{(v.): to giggle}

  It was an academic matter, of course, and yet, incredibly, Darconville found
himself thus preoccupied, searching to apprehend 
\mynote{apprehend}{(v.): OED 2. To seize (a person) in name of law, to arrest. }
a figure in the distance, an
interval so wide, for a thousand reasons, it dulled the edge of consequence but
a distance nevertheless that also warned, by its very remoteness, that no real
exchange of feelings was possible, a morbid 
\mynote{morbid}{(adj.): of the nature of or indicative of disease,  
  ad. L. morbid-us, f. morb-us disease, f. root of morī to die. Cf. F. morbide, 
  It., Pg. morbido, Sp. mórbido. }
underbreath, as if aloofness itself, whispering, told him to desist 
\mynote{desist}{(v.):  to cease (from some action or procedure); to stop, leave off, 
  give over, forbear. From OF. desister (1358 in Littré; mod.F. dé-), ad. L. dēsist-ĕre, f. de- + 
  sistĕre to stop, stand still. }
from that which he must inevitably be excluded.
And yet how often, even from childhood, had this been the case in his life! And
so he listened to what he thought, serving to clarify a matter at the heart of
things and verify a falsehood. It formulated his final remarks.

  "The artist," concluded Darconville, going to the window and staring out at a
group of hardy catalpas, 
\footnote{ a genus of flowering plants in the family Bignoniaceae, native to warm 
temperate and subtropical regions of North America, the Caribbean, and East Asia. }
"when a logical paradox to himself becomes, I suppose,
the most unpoetical of men." A cardinal went snip-snip-snip in the trees. "I
don't know, perhaps Keats had one glass of claret 
\mynote{claret}{(n.): red wine }
too many, loitered 
\mynote{loiter}{(v.): to linger idly about a place; to waste time
   when engaged in some particular task.}
palely, and was overcome with the necessity of writing a line that could 
acknowledge two simultaneous but incompatible forces within him: poet and 
priest. He becomes a pomologist"
\mynote{pomologist}{(n.): someone versed in pomology or someone who cultivates 
  fruit trees }
---Ailsa Cragg looked frostily at Darconville and sucked a
tooth---"trying to get fruit back to its paradisaical best. `Beauty is Truth,
Truth Beauty'? It's a \textit{concio ad clerum}: 
a sermon to the clergy, wherein the poet
is attempting to cheer himself up---in a fit of un-steadfastness, say?---and is at
the same time begging us to accept not only what we should but also that which,
poor poet, he more than sufficiently allowed us without special pleading. The
poet's own voice has interrupted him. He is awakened by his own snore. The line,
which I disgressed to explain, is only a false bottom, collapsing an otherwise
beautiful poem---and one which provides, perhaps, the best single commentary we
have on the sonnet, `Bright Star.'" Darconville was finished. "Are there any
questions?"

  There was silence.

  "We're clearer then," asked Darconville, "on the meaning of the poem?"

  "It hath to do with the thufferingth and thorrowth of life."

  Everybody groaned.

  "I believe that young boy," came a sudden voluptuous 
\mynote{voluptuous}{(adj.): OED 3.a. Imparting a sense of delicious pleasure;
		suggestive of sensuous pleasures, esp. of a refined or luxurious kind. 
		From OF. (also mod.F.) voluptueux, -euse (= Sp. and Pg.
		voluptuoso, It. voluttuoso), or L. voluptuōsus
		(Pliny, etc.), f. voluptas pleasure, VOLUPTY.}
drawl 
\mynote{drawl}{(n.): The action of drawling; a slow, indolent utterance.} 
from near the back
of the room, a low magical level of loveliness which absorbed with a kind of
absoluteness any rival utterance that might have been offered against it, "was
just sayin' the most obvious thing in the whole wide world."

  It was the kind of female voice that Juvenal somewhere naughtily describes as
having fingers.

  "He was talkin'"---she touched her tongue to her upper lip--- "about love."

  Most of the girls, in icy silence, stared straight ahead.

  "I'm sorry," said Darconville who, although not taken in---he'd seen her back
there at the beginning of class and almost with amusement had been waiting for
her to speak---nevertheless feigned 
\mynote{feign}{(v.): OED II.8. trans. To make a show of, put on an appearance of, put on,
    pretend, simulate, sham; to pretend to utter (words). }
surprise, "but I can't seem to find you here
in my roll-book. You are a senior, aren't you, Miss Poore?"

  It was too much. Every head in the entire class---excepting one student's, whose
head was lowered---abruptly shifted from the teacher to her, the only student's
name he had known! And there she sat, confident and cool, wearing a nubby
\mynote{nubby}{(adj.): of textiles; having a rough surface}
placket-front 
\mynote{placket}{(n.): In modern usage, the term placket often refers to the
    double layers of fabric that hold the buttons and buttonholes in a shirt.}
shirt trimmed 
\mynote{trim}{(v.): OED 8. spec. To decorate (a hat, garment, etc.) with
  ribbons, laces, feathers, flowers, braids, embroideries, or the like, so as to
  give it a finished appearance.} 
in beige 
\mynote{beige}{(n.): OED A.2. A shade of colour like that of undyed and
        unbleached wool; yellowish-grey.}
to match a black velvet dirndl.
\mynote{dirndl}{(n.): OED, a style of woman's dress imitating Alpine peasant costume
    with bodice and full skirt.}

  Then, very slowly, Hypsipyle Poore lifted off her sunglasses and, with one
lingering provocative glance, revealed eyes as limpid
\mynote{limpid}{(adj.): OED a. Chiefly of fluids: Free from turbidity or
    suspended matter; pellucid, clear.}
as the pools in Heshbon by the gate of Bathrabbim
\footnote{KJV, Song of Solomon 7:4 Thy neck is as a tower of ivory; thine eyes
  like the fishpools in Heshbon, by the gate of Bathrabbim: thy nose is as the
  tower of Lebanon which looketh toward Damascus.}
--- and everyone there could have sworn she winked as with a
cryptic smile she breathed softly, "Well, should I see you after class?"

  "Love has been mentioned," said Darconville, who saw he had a minute or two.
"And suffering, as well. The history of romantic disappointment, I don't doubt,
often does nothing more than document the schism 
\mynote{schism}{(n.): OED 3.a. gen. In early use, a state of disunion,
  dissension, or mutual hostility. Now with more restricted meaning (influenced by
  sense 2 and the Gr. etymology), a division into mutually opposing parties of a
  body of persons that have previously acted in concert. Also, in recent use, a
  severance of unity, a discord, breach (between persons or things). }
between Beauty and Truth or, better, proves that Beauty, when it becomes 
an end in itself, often yields no
Truth. The simple line, in such cases, had no complexity within it---there can, of
course, never be too much. A knowledge of many things is possible, it's been
said, but one can never know everything about one thing, though, sadly, one
perhaps tries. The relationship with a boy or girl you spontaneously took for
perfection-in-beauty but didn't sequentially know by examination-in-truth can
result in disaster. The implications of the Ideal? Who can really know? It is
the chance one takes when one falls in love---the discussion of which," smiled
Darconville, "is perhaps beyond the province of the classroom here, agreed?"

  The girls laughed---and prepared to leave. But Darconville, alert to the chance,
quickly took up his class roll 
\mynote{roll}{(n.): OED 4. A list of names used to ascertain whether each one
  of a set of persons is present; esp. Mil. (= muster-roll) or in scholastic use.}
and began to call attendance.

  "Muriel Ambler"

  "Here"

  "Melody Blume"

  "Here"

  The students, by turn, acknowledged their names. But for the suspense it might
have been interminable, the list seemed so long. Here. Here. Here. But which was
hers? Was it there?

  "Sheila Mangelwurzel"

  "Here"

  "Christie McCarkle"

  A carrot-top put up her hand.

  "Glycera Pentlock"

  "Over here"

  "Isabel Rawsthorne"

  A thunderclap!---then a flashing light across his book. Dazed, almost
unconsciously, Darconville called out the remaining non-names, reflecting on
this alone: only love makes the pain of lifting a shy head a grace. It equally
unlocks a silent throat with the knowledge in the plaintive 
\mynote{plaintive}{(adj.): OED 3. Having the character of complaint; expressive
   of sorrow; mournful, sad.}
soul of what must be done, and so was, in the tender-taken breath he suddenly 
heard, soft as a swoon,
\mynote{swoon}{(n.): OED 1.b. a fainting-fit.}
say: "Present."

  And sculptured Aphrodite, in loose-flowing alb,
\mynote{alb}{(n.): A tunic or vestment of white cloth reaching to the feet, and
  enveloping the entire person; a variety of the surplice, but with close sleeves;
  worn by clerics in religious ceremonies, and by some consecrated kings. }
stepping slow and fragile under the loops of dripping ring willows, 
puts a finger to her lips, shakes out tearshaped petals 
\mynote{petal}{(n.): one of the modified often brightly colored leaves of the
    corolla of a flower}
from her flasket, 
\mynote{flasket}{(n.): OED 1. a long shallow basket.}
and then holds high her holy lights which silver 
\mynote{silver}{(v.): OED 2. To invest or suffuse with a silvery hue or lustre.} 
and bewitch the common wood where two, 
\footnote{\textdbend}
long asleep, sweetly wake under the baldric 
\mynote{baldric}{(n.) A belt or girdle, usually of leather and richly
    ornamented, worn pendent from one shoulder across the breast and under the
    opposite arm, and used to support the wearer's sword, bugle, etc.} 
of a new heaven to blush at the silence, pause, and faintly hear a
goddess whisper in a voice lower than leaf-fall: "Tomorrow shall be love for the
loveless, and for the lover tomorrow shall be love."

  In short, Isabel Rawsthorne looked up and smiled.
 
