\chapter{The Beginning}

\begin{quotation}
Delirium 
\mynote{delirium}{(n.) (dɪˈlɪrɪəm): a temporary mental state with a sudden onset, usually reversible,
including symptoms of confusion, inability to concentrate, disorientation,
anxiety, and sometimes hallucinations. Causes can include dehydration, drug
intoxication, and severe infection. 
}
is the disease of the night. \\
\sourceatright{--- St. PONTEFRACT}
\end{quotation} 
\vspace{0.2cm}

DARCONVILLE, the schoolmaster, always wore black. The single tree, however, that
shanked 
\mynote{shank}{\textdbend (v.): to fall off, as a leaf, flower, or capsule, on 
account of disease affecting the supporting footstalk;}
out of the front yard he now crossed in long strides 
\mynote{stride}{(n.): long step}
showed even more distinct a darkness, a simulacrum 
\mynote{simulacrum}{(n.): a likeness, a semblance, a mock appearance}
of the dread probationary 
\mynote{probationary}{(adj.): OED A.1., of, pertaining or relating to, or serving 
for probation; made, performed, or observed in the way of probation; belonging 
to the testing or trial of character or qualifications. \textit{probation} 
comes from Latin \textit{probatio}, French \textit{probare} to try, examine, prove.
}
tree --- trapfall 
\mynote{trapfall}{(n.): OED, from trap+fall, a trap consisting of a trap-door 
or covering over a pit or cellar arranged so as to give way beneath the feet. 
Also fig.
}
of all lost love --- for coming upon it, gibbet-high 
\mynote{gibbet}{(n.) (ˈdʒɪbɪt): a structure used to execute criminals by hanging.}
and half leafless in the moonlight, was to feel somehow disposed 
\mynote{disposed}{(adj.): inclined, minded. \textit{dispose} as a verb has several
meanings: (1) (vi. used with of) to eliminate or get rid of something; (2) to
distribute and put in place; (3) to deal out, to assign to a use; (4) to
incline.  }
to the general truth that it is a dangerous and pagan notion that beauty palliates 
\mynote{palliate}{(v.): to cover with excuses; to conceal the enormity of, by 
excuses and apologies. From Late Latin \textit{palliatus}, past participle of 
\textit{palliare} to cloak, conceal, from Latin \textit{pallium} cloak.
}
evil.

  He was alone. It had always seemed axiomatic for him that he be alone: a vow,
the linchpin 
\mynote{linchpin}{(n.): Originally means a pin used to prevent the wheel of a 
vehicle from sliding off the axletree. Now figuratively means a person or 
thing that holds something together: the most important part of a complex 
situation or system.
}
of his art, his praxis.
\mynote{praxis}{(n.): exercise or practice of an art, science, or skill.
From Greek πρᾶξις doing, acting, action, practice, noun of action from 
πράττειν to do; whence medieval Latin \textit{praxis} (Albertus Magnus 
\textit{Metaphys.} v. v. ii, c 1255).
}

  The imperscrutable 
\mynote{imperscrutable}{(adj.): not capable of being searched out; inscrutable.}
winds of autumn, blowing leaves across the porch, had
almost stripped the tree, leaving it nearly naked and essential against the moon
that shone down on the quiet little town in Virginia. It was late as he let
himself into the house and walked up the creaking stairs to his rooms where,
pulling a chair to the window, he sat meditatively in that dark chamber like a
nomadic gulsar
\mynote{gulsar}{(n.): \textdbend guslar? i.e. one who plays the gusla (a 
Bulgarian stringed instrument).}
--- his black coat still unbuttoned --- and was left alone with those odd
retrospective prophecies borne in on one at the start of that random moment we,
for some reason, choose to call the beginning of a new life.

  The night, solemn and beautiful, seemed fashioned to force those who would
observe it to look within themselves. He watched awhile and then grew weary. He
took a late mixt 
\mynote{mixt}{(n.): a product of mixture, a compound.}
of some rolls and a bottle of ale and soon dropped asleep on
his bed, dreaming out of fallen reason the rhymes received with joy he shaped
accordingly. It was only early the following morning that he found on the
bedside table next to his pen and unscrewed cap --- a huge Moore's Non-Leakable
--- the open commonplace book in which, having arisen in the middle of the 
night to do so, he had written a single question: ``Who is she?''
