\chapter{Hypsipyle Poore}

\begin{quotation}
\begin{verse}
  \hspace{-1.8em} Do what thou wilt, thou shalt not so, \\
  Dark Angel! triumph over me:  \\
  Lonely, unto the Lone I go;   \\
  Divine, to the Divinity.
\end{verse}
\sourceatright{--- LIONEL JOHNSON, "The Dark Angel"}
\end{quotation}
\vspace{0.2cm}
 
 
  "I NEED IT!"

  "That's a lot of pudding, miss."

  "I'm signing up, anyway."

  "The course, I told you yesterday, is not open to seniors," exclaimed Mrs.
McAwaddle, the registrar, her mouth cemented shut against the possibility of
further discourse.

  "My daddy," the girl drawled, 
\mynote{drawl}{(v.): to prolong or lengthen out the sounds of speech in 
an indolent or affected manner; to speak slowly, by affectedly prolonging the 
words. 
}
charging through in interruption and waving a
slip of paper, "had the dean on the telephone last night. Now, y'all want to
read this?"

  The student, obviously used to exacting compliance, was an arresting young
beauty in sunglasses with a soft pink sweater, raven-black hair cut to
perfection, and a pout 
\mynote{pout}{(n.): a protrusion of the lips, expressive of pique or annoyance.}
of wet lipstick that made her mouth look like a piece of
candy. She stepped back, unvanquished, 
\mynote{unvanquished}{(adj.):  Not vanquished or overcome; unsubdued.}
and seemed satisfied to wait, speaking to 
no one but admiring from a corner the indisposition 
\mynote{indisposition}{(n.): disordered bodily condition; ill health, illness, ailment; esp. of a
  slight or passing character. }
of the other girls there who
were trying to arrange their schedules during the first discouraging days of
registration. They shuffled about in determined little squads with drop/add
cards, course syllabi, and countless papers to have signed and stamped.

  The room was warm, sticky, from the crowding bodies, but the girls, somehow,
all smelled of fresh soap and mint-flavored gum. They were Southern girls, after
all, and unlike their counterparts in other sections of the country whose
morning beauty, from the normal wear-and-tear of a day, too easily faded only to
be carelessly ignored, they tried to keep powder-room perfect and as presentable
as possible. And, after all, this time there was a man present in the room.

  "Some people can just whistle and wait," snapped 
\mynote{snap}{(v.): to utter sharp, tart, or cutting words or remarks;
  to speak or reply irritably or abruptly.}
Mrs. McAwaddle, going jimmy-jawed. 
\footnote{ having a projecting lower jaw.}
A mite 
\mynote{mite}{(n.):  a very small object; often, a very small living creature.}
of a thing, resembling the perky 
\mynote{perky}{(adj.):  Self-assertive, forward, somewhat obtrusive or assuming;
  self-conceited, ‘cocky’; jaunty; smart.}
little owl commonly depicted resting on the hand of Minerva, she was wearing a 
dress covered with hearts. She spindled 
\mynote{spindle}{(v.): to impale on a device for holding paper documents.}
a card angrily and looked up at the man who'd been standing
there for some time. And now. His name? His business?

  "Darconville," he said, smiling.

  The girl with the black hair, waiting behind him, took a crystal vial 
\mynote{vial}{(n.): a vessel of a small or moderate size used for holding liquids; in
  later use spec., a small glass bottle, a phial.}
out of her handbag (tooled in studs: "H.P.") and, closing her eyes, sprayed a 
musky lavender fragrance around her perfect, prematurely formed silhouette, 
\mynote{silhouette}{(n.): a dark outline,a shadow in profile, thrown up against a lighter
  background.}
waxen 
\mynote{waxen}{(adj.):  grown up, full-grown, adult}
as a delicate shell.

  "Ah, of course," beamed the owl of Minerva, "the new professor." They laughed
together. "And I do believe I detect a Yankee accent?"

  Darconville asked about his courses. He had met his freshmen already but
hadn't yet been given the list.

  "Your freshman class list? O dear," muttered Mrs. McAwaddle. Apparently, it
had been forced into more revisions than a Dixiecrat 
\footnote{any of the Southern Democrats who seceded from the party in 1948 in
  opposition to its policy of extending civil rights.}
caucus 
\mynote{caucus}{(n.): In U.S. A private meeting of the leaders or representatives of a
political party, previous to an election or to a general meeting of the party,
to select candidates for office, or to concert other measures for the furthering
of party interests; opprobriously, a meeting of ‘wire-pullers’. 
}
and so laid aside. She searched a tray, lifted up a snow water-ball of glass, 
and then shot open an acidgreen file-cabinet, finally rummaging 
\mynote{rummage}{(v.):  to bring out by searching.}
up the sheets for English 100. He read the first few names on the list:
 
\begin{verse}
        \hspace{6em} Muriel Ambler  \\
        \hspace{6em} Melody Blume   \\
        \hspace{6em} Ava Caelano    \\
        \hspace{6em} Wroberta Carter    \\
        \hspace{6em} Barbara Celarent   \\
        \hspace{6em} Analecta Cisterciana
\end{verse}
 
  "Would you let me," smiled Mrs. McAwaddle, her little head at a prayerful
angle, "admire your coat?" Several girls nearby exchanged glances and winked.
Darconville's coat people loved. Cut to princely lines, it was an English
chesterfield 
\mynote{chesterfield}{(n.):  a kind of overcoat}
as black as the black swan of Juvenal. "I do declare, if this isn't
the most dashing
\mynote{dashing}{(adj.): transf. Of things: Fashionably showy; stylish, `swell'. }
---but here," she added, hunching up to the desk, "there's an
ever-so-small tear here, by that button." She patted his hand. "Now you have
your wife mend 
\mynote{mend}{(v.): to remove or atone for defects}
that and---"

  Darconville leaned forward and, like Wotan 
\footnote{ Odin, a deity of the old Germanic religion, and of modern German
reconstructions of this religion.}
consulting a weaving norn, 
\mynote{norn}{(n.):  One of the female Fates recognized in Scandinavian mythology. }
whispered with a close smile, "I am not married."

  Mrs. McAwaddle stared a moment---and then, with a conspiratorial 
wink, motioned him into a side-room off the registrar's office where under a 
portrait of Jefferson Davis a purple-stained mimeograph 
\mynote{mimeograph}{(n.): A duplicating machine for producing copies from a stencil;}
machine went bwam-bwam-bwam, spitting out single sheets of copy. Conventicle 
\mynote{conventicle}{(n.): meeting, assembly}
gave way to conclave. 
\mynote{conclave}{(n.): a private meeting or secret assembly}
"Always remember," she said, gripping his wrist, "to a handsome boy like you
---how old are you?"

  "Twenty-nine."

  She made the sound of a pip. 
\mynote{pip}{(n.): Something remarkably good; an excellent person or thing. Freq. 
  \textit{a pip}. Cf. \textit{pippin}, slang (orig. and chiefly U.S.). }
"---to a handsome boy like you something wonderful
will happen. You'll find a girl to love you as sure as wax candles have wicks."
\mynote{wick}{(n.): the bundle of fibre, now usually loosely twisted or woven cotton
(formerly rushes, tow, flax, etc.) in a lamp, candle, or taper (formerly also in
a torch), immersed or inclosed except at one end in the oil or grease, which it
absorbs and draws up on being kindled at the free end, so as to maintain the
flame. 
}
She had lost her own husband, she told him, eight years ago (tainted knockwurst,
\footnote{ =knackwurst, a type of German sausage.}
Jaycee picnic), and now it was just hell dipped in misery. To live alone? When
the good Lord created someone for everybody? Ridiculous! "But now," she added,
folding his hands in her own, "you be careful: these girls at Quinsy College can
work the insides out of a boy without him having a clue and, simple yokums
\footnote{ Slang., a rube, bumpkin, or otherwise green or stupid person.}
though they may seem, can be the untellinest 
\footnote{ \textdbend From \textit{untelling}, innumerable, countless.}
little commodities on earth. You saw that child out front fussing at me? I could 
have spit nails. They have nerve to burn."

  For the advice of this little hole-in-corner sibyl Darconville was grateful,
though he couldn't help but attribute her suspicions to the exigencies 
\mynote{exigency}{(n.): That which is needed or required; demands,
  needs, requirements.}
of her job, and, excusing himself, he worked his way out of the office, looking 
up only once: but then to catch, above a blur of fluffy 
\mynote{fluffy}{(adj.): Consisting of or resembling fluff; of soft, downy texture. }
pink, a perfect face unblindfolding slowly from a pair of sunglasses the 
fire-flash of two beautiful but dangerous eyes in which, he thought, he 
detected a sly, 
\mynote{sly}{(adj.): Playfully mischievous or malicious; roguish; waggish. }
premeditated smile. He paused in the outer corridor and once again looked at 
the names entered alphabetically on the list; it read like a spice chart:
 
\begin{verse}
        \hspace{6em} Ailsa Cragg   \\
        \hspace{6em} Childrey Fawcett  \\
        \hspace{6em} Galveston Foster  \\
        \hspace{6em} Scarlet Foxwell   \\
        \hspace{6em} Opal Garten       \\
        \hspace{6em} Marsha Goforth    \\
        \hspace{6em} LeHigh Hialeah    \\
        \hspace{6em} Elsie Magoun      \\
        \hspace{6em} Sheila Mangelwurzel
\end{verse}
 
  A single name there meant something more than it was: a symbolum 
\footnote{ Latin, symbol.}
---both a "sign" and a "confession." But which one was it? Which one was hers? 
It was curious, his preoccupation, for he'd seen her only once. Inexplicably, 
however, it mattered, if only for the hardly momentous irony that by knowing 
it he could then immediately dismiss it and put an end to it all. Her look 
had injured a silence in his life. The known name might somehow injure the 
look, and with the look gone the silence could continue, allowing him in 
consequence and inducing, for diversion, the equanimity 
\mynote{equanimity}{(n.): Fairness of judgement, impartiality, equity. }
to create out of the dormitive 
\mynote{dormitive}{(adj.): causing sleep; soporific.}
world the something
out of nothing we call art. There was actually scant attention being paid to
this unnamed girl in the upper part of his mind, but in the lower reaches she
several times appeared, a thing, rather like the libration of the moon,
\footnote{\textit{libration of the moon}: an apparent irregularity of the moon's
motion which makes it appear to oscillate in such a manner that the parts near
the edge of the disk are alternately visible and invisible. (There are three
kinds, called libration in latitude, libration in longitude, and diurnal
libration or parallactic libration.) 
}
alternately visible and invisible.

  He flipped a sheet and read more names.
 
\begin{verse}
        \hspace{6em} Christie McCarkle   \\
        \hspace{6em} Trinley Moss        \\
        \hspace{6em} Glycera Pentlock    \\
        \hspace{6em} Hallowe'ena Rampling  \\
        \hspace{6em} Isabel Rawsthorne   \\
        \hspace{6em} Cecilia Sketchley
\end{verse}
 
  Darconville couldn't help but smile. The names seemed absurd, but one didn't
really have to spend very much time down South to realize the regional
compulsion for this particular extravagance, daily coming upon such weird
examples as: Cylvia, Olgalene, Marcelette, Scharlott, Coquetilla, Mavis,
Latrina, Weeda and Needa, Mariedythe, Romiette, Coita, Vannelda, Moonean, Rhey,
Flouzelle, Balpha, Erdix, Colice, Icel, and Juella, all desperate parental
attempts to try to work some kind of sympathetic magic upon their daughters from
the very start. And yet how was it that upon hearing them one saw only
majorettes, 
\mynote{majorette}{(n.):  =drum-majorette, A female drum-major (drum-major 1 c); a girl who
  leads or takes part in a parade or the like, twirling a baton, etc.}
waitresses, and roller-derby queens?

  Darconville passed by the refectory 
\mynote{refectory}{(n.): A room for refreshment; esp. in religious houses and colleges, the
  hall or chamber in which the meals take place. }
(and the odor of mercilessly boiled brussels sprouts)
\footnote{ \textit{brussels sprout}: the bud-bearing Cabbage (Brassica oleracea 
gemmifera), a variety producing buds like small cabbages in the axils of its 
leaves. 
}
and sat down in a circular room where in the center stood a
sculpture of Chapu's Joan of Arc, the college patron. This was known as the
Rotunda. The main building at Quinsy College, its egg-shaped interior was a
respectable cream-and-green color, open, as it took one's attention higher and
higher past two circular balustrades, 
\mynote{balustrade}{(n.):  A row of balusters, surmounted by a rail or coping, forming an 
  ornamental parapet or barrier along the edge of a terrace, balcony, etc. }
to a voluminous inner dome covered with fake but sumptuous, 
\mynote{sumptuous}{(adj.): Of buildings, apparel, repasts, and the like: Made or produced at
great cost; costly and (hence) magnificent in workmanship, construction,
decoration, etc. 
}
over-elaborate, neo-biblical murals in rose and gold. On
several walls at ground level, a series of past college presidents, bald and
severe, glowered 
\mynote{glower}{(v.): To stare with wide-open eyes; to gaze intently
  or with an air of surprise. }
out of their frames. He was still reviewing the list, empty
pier-glasses, and pronouncing names, all but hers hostile to him because not
hers, but yet none hostile because to him any might be.
 
\begin{verse}
        \hspace{6em} Butone Slocum   \\
        \hspace{6em} Millette Snipes  \\
        \hspace{6em} April Springlove \\
        \hspace{6em} Lately Thompson  \\
        \hspace{6em} DeDonda Umpton
\end{verse}
 
  The memory had persisted. On an otherwise unexceptional day, for the first
time, he'd met that class of freshmen, silent little elves bunched-up 
\mynote{bunch}{(v.): To crowd together in a body.}
and sitting terrible-eyed as they contemplated the four years of college to 
come. No one had spoken or said so much as a syllable, but all took down the 
assignment and then the name he'd chalked on the board by way of introduction 
as if they were borrowing it for some felonious purpose. And then, turning, he 
had seen the girl, a face out of Domenichino 
\footnote{ Domenico Zampieri known as Domenichino for his shortness (October 21,
1581 – April 6, 1641), was an Italian Baroque painter of the Bolognese or
Carracci School of painters.
}
declaiming 
\mynote{declaim}{(v.): To speak or utter aloud with studied rhetorical
  expression; to repeat or recite rhetorically. }
itself with the supremacy of a mere look that rose like an oriental sun not 
announced by dawn and setting left no twilight---only the persistent memory 
of two brown eyes, soft and fraught 
\mynote{fraught}{(adj.): Stored, supplied, furnished, filled, equipped with. }
with soul, imparting 
\mynote{impart}{(v.): To communicate as knowledge or information; to
  make known, tell, relate.}
a strange kind of consecration. 
\mynote{consecration}{(n.): The action of consecrating; a setting apart as dedicated to the
  Deity; dedication with religious rites to a sacred purpose.}
Darconville, looking through the mist of his reverie, 
\mynote{reverie}{(n.): A fit of abstracted musing; a `brown study' or day-dream. }
then turned from his own idle thoughts and read the last names on the list:
 
\begin{verse}
        \hspace{6em} Shelby Uprightly   \\
        \hspace{6em} Martha Van Ramm    \\
        \hspace{6em} Poteet Wilson      \\
        \hspace{6em} Rachel Windt       \\
        \hspace{6em} Laurie Lee Zenker
\end{verse}
 
  "Yoo-hoo!" Halfway down the front stairs, Darconville turned back to the
voice. It was Mrs. McAwaddle, scooting after him on her tiny slue 
\footnote{ \textit{slew-foot} or \textit{slue-foot}: U.S. slang, A person who
  walks with his feet turned out; a clumsy person. }
feet. She was relieved she'd caught up with him, she said, puffing, her hand 
pressed to her heart. "I'm doing my level best to keep your classes down to a 
minimum, especially from those"---she handed him a piece of paper---"who have 
no business being there. The dean has decided to leave the matter up to you."

  The particular piece of paper, the formal request of a senior to take his
freshman course, was signed by the dean and countersigned in an affected paraph
\mynote{paraph}{(n.): A flourish made after a signature, originally as a kind of precaution
  against forgery. }
of lavender ink with the name: Hypsipyle Poore.

  "You remember that child out front fussing at me?" Mrs. McAwaddle shook her
head. "They have nerve to burn." And she squeezed Darconville's hand, turned,
and trundled 
\mynote{trundle}{(v.): To move along on a surface by revolving; to
  roll. Also figuratively.}
away toward her office, one shoulder lower than the other.
"Yoo-hoo!" Darconville looked up again to see Mrs. McAwaddle standing on the
landing. "Be careful."

  Dear Mrs. McAwaddle, wise Mrs. McAwaddle, widowed Mrs. McAwaddle, owlish Mrs.
McAwaddle, compassionate Mrs. McAwaddle, Mrs. McAwaddle in her dress of hearts!
But how could she know, poor soul, that it was entirely someone else who was on
his mind and to whom that stricture better applied: be careful. But of what? Of
whom? For still, of the many names she could have, she had none.

  Darconville, however, consigned 
\mynote{consign}{(v.): To hand over to another for custody; to entrust
  or commit to another's charge or care. }
her to the obscure and folding the class list
into his pocket walked out into the lovely afternoon, the rarefactions 
\mynote{rarefaction}{(n.): The action of rarefying, or process of being rarefied; diminution of
  density.}
in the air opposing, however pleasantly, his general conviction that the state 
of art should be in constant panic. The artistic nature, he knew, had an inborn
proneness to side with the beauty that breaks hearts, to single out the
aristocratic contours of what in human glory quickens the impulses of life to
mystic proportions. He found himself, again, absent-mindedly thinking of the
effect of that look in which everything that was most obscure in the relation
between two people rose to the surface, and yet he could find no possible
expression of it in words. But curiosity, he thought---the weakest form of
solicitude, even if it was the beginning of it---was not love.

  And crossing the lawn he only hoped that he'd gain somehow in veracity 
\mynote{veracity}{(n.): The quality or character in persons of speaking or stating the truth;
  habitual observance of the truth; truthfulness, veraciousness. }
what he
lost in mystery---a compromise, it also occurred to him, he wasn't always ready to
make. But what was he ready for? He didn't know. And so laughing he headed home,
walking without so much as a touch of regret up the street to his house, his
book, and the supra-mundane.
\mynote{supra-mundane}{(adj.):  =\textit{supermundane}, Elevated in nature or character above what
  pertains to the earth or world; belonging to a region above the world.}
