\chapter{Quinsyburg, Va.}
 
\begin{quotation}
  Death hath not only particular stars in Heaven, but malevolent Places on
Earth, which single out our Infirmities 
\mynote{infirmity}{(n.): Weakness or want of strength; lack of power to do something;
inability. Also with pl. an instance or case of this.
}
and strike at our weaker Parts.
\sourceatright{--- Sir THOMAS BROWNE, \textit{A Letter to a Friend}}
\end{quotation}
\vspace{0.2cm}

RULE A LINE from Charlottesville, directly through Scottsville and its lazy river, 
and draw it down --- a straight 180° --- into the southside, fixing it to a 
terminal in the 
heart of Prince Edward County, Va. Follow scale to measure the low point. Now, 
drive that sixty miles of narrow godforsaken road past old huts and shacks, scrub 
pines and blasted 
\mynote{blasted}{(adj.): cursed, damned.}
forests into a desolation 
\mynote{desolation}{(n.): The condition of a place which by hostile ravaging or by natural
character is unfit for habitation; waste or ruined state; dreary barrenness.
}
the crossed boundaries of which, though 
not silent to your eyes, one feels more in the depth of imagination, the kind of 
anxiety, a foreboding, 
\mynote{foreboding}{(n.): The action of the vb. forebode; hence, a prediction, 
presage. (Now only of evil.)
}
of a guilt within not traceable to a fact without; turn then and 
trail slowly on a wind across a tableland of sallow 
\mynote{sallow}{(adj.): having a sickly yellow or brownish yellow color.}
weeds and sunken 
\mynote{sunken}{(adj.): that has sunk below the usual or general level; subsided. }
dingles 
\mynote{dingle}{(n.): a deep dell or hollow; now usually applied (app. after Milton) to one
that is closely wooded or shaded with trees; but, according to Ray and in mod.
Yorkshire dialect, the name of a deep narrow cleft between hills. 
}
into flat tobacco country where the absence of perspective seems as if offered 
in awful proof of what suddenly, crouching in a perfect and primitive 
isolation, becomes a town. 
Stop your car. Your hesitations are real. You can hear yourself breathing. You can 
hear your hands move. You are in Quinsyburg.

  It is immediately a terrible letdown, a dislocation, solitary in the framework 
of its rigid and iconoclastic 
\mynote{iconoclastic}{(adj.): of or pertaining to iconoclasts or iconoclasm, 
i.e. image breakers.}
literalness, which yet sits in the exact center of the Commonwealth of Virginia, a 
state commemoratively 
\mynote{commemoratively}{(adv.): Having the attribute of commemorating. 
\textit{commemorate} as a verb means to call to the remembrance of hearers or readers.
}
named --- if we may charitably disregard for a moment her biological interludes 
\mynote{interlude}{(n.): an interval in the course of some action or event; an intervening time
or space of a different character or sort.}
with everyone from the 
royal dancing master to the beetle-faced Duke of Anjou
\footnote{Author's Note: Greetings here might also be extended to Alençon; Charles Blount; Hatton; 
De Vere; Heneage; Sir Walter Raleigh; Admiral Thomas Seymour; Robert Dudley, 
Earl of Leicester; Eric, King of Sweden; Robert Devereux, Earl of Essex; 
Archduke Charles of Austria; and any number of others.}  
--- after the twenty-third British sovereign, Elizabeth I, she of the judas wig
: bastard, usurper, 
\mynote{usurper}{(n.): One who usurps a crown or throne; one who seizes or arrogates supreme
power or authority without right or just cause.}
excommunicate, baldpate, heretic, murderess, schismatic, 
\mynote{schismatic}{(n.): one who promotes or countenances schism or breach of external unity in
the Church; one who is guilty of the sin of schism; a member or adherent of a
schismatical body. }
and willing copulatrix. 
\footnote{Latin for she who couples, connects.}
The sharp and instinctive disappointment you feel, that this must be the capital 
city of all failure, wrongheadedness, and provinciality, does not subside --- it 
increases, intensifies, heightens. The approach that announces with sadder and 
sadder emphasis its sterility leads only to a confirmation of its deeper afflictions: 
\mynote{affliction}{(n.): a pain, calamity, grief, distress.}
for it amounts to, infringes 
\mynote{infringe}{(vi.): to break in or encroach on or upon.}
on, nothing, shares nothing with the prospect of the sky but, deathshot 
with monotony, lies like a shroud wrapped around itself as if, so 
determined, it refuses to be inhabited by even so much as the relative 
humanity of a corpse. It is infinitely liker hell than earth, the proper 
place to feel the first hint of the decay of the fall. It appears to have 
extracted from beauty the piety given to it 
and, keeping that, dismissed the rest as ignominious 
\mynote{ignominious}{(adj.): full of ignominy; involving shame, disgrace, or 
obloquy; shameful, disgraceful, discreditable.
}
accident to build a town. A sign tells you where.
 
  The place --- nothing surrounded by nowhere --- is rigorously confined within its 
own settled limits, huddled, as if on its knees searching the corners of its rural 
conscience for some sin of omission or commission whereby, to ratify 
\mynote{ratify}{(v.): to give formal consent to}
the truth of natural depravity, every pleasure, every recreation, every trifle 
scratched out of the 
dust might then be magnified into a great offense, and less for its severe white 
churches than a general mood of dissent do you feel that the deepest solicitudes 
\mynote{solicitude}{(n.): The state of being solicitous or uneasy in mind; 
disquietude, anxiety; care, concern.
} 
of its inhabitants must have nothing to feed on except by what either outraged 
godliness or gave the devil his due. Crete had no owls, Thebes no swallows, Ithaca 
no hares, Pontus no asses, Scythia no swine. Quinsyburg had no hope. It is 
overpowering to realize and worse suddenly to accept, you fight it, and only the 
muteness of apprehension stifles 
\mynote{stifle}{(v.): to prevent or constrain}
an immediate impulse to cry out in despair, "Of 
all the loveless, lifeless things that quail 
\mynote{quail}{(v.): to give way, yield to or before.}
beneath the wrath of God, commend 
\mynote{commend}{(v.): to entrust someone or something to, cf. recommend}
Quinsyburg to it!"
 
 
 
